---
title: "01f_USEmploymentDemographics"
output: html_document
---

In the wake of the Great Recession of 2009, there has been a good deal of focus on employment statistics, 1 of the most important metrics policymakers use to gauge the overall strength of the economy. In the US, the government measures unemployment using the Current Population Survey (CPS), which collects demographic and employment info from a wide range of Americans each month. 

Observations in the dataset represent people surveyed in the September 2013 CPS *who actually completed a survey*. While the full dataset has 385 variables, we will use a more compact version of the dataset, `CPSData.csv`, which has the following variables: `PeopleInHousehold`: The number of people in interviewee's household, `Region`: census region where interviewee lives, `State`: state where interviewee lives, `MetroAreaCode`: A code that ID's the metropolitan area in which interviewee lives (missing if interviewee does not live in a metropolitan area, and the mapping from codes to names of metropolitan areas is provided in the `MetroAreaCodes.csv`), `Age`: age, in years, of interviewee (80 = people aged 80-84, 85 = people aged 85+), `Married`: marriage status of interviewee, `Sex`: sex of interviewee, `Education`: maximum level of education obtained by interviewee, `Race`: race of the interviewee, `Hispanic`: Whether interviewee is of Hispanic ethnicity, `CountryOfBirthCode`: A code ID'ing the country of birth of interviewee (mapping from codes to names of countries is provided in `CountryCodes.csv`, `Citizenship`: United States citizenship status of the interviewee, `EmploymentStatus`: status of employment of interviewee, `Industry: industry of employment of interviewee (only available if employed).

Codes like `MetroAreaCode` and `CountryOfBirthCode` are a compact way to encode factor variables wtih text as possible values, and are therefore quite common in survey datasets. In fact, all but 1 variable in this dataset was actually stored by a numeric code in the original CPS datafile. When analyzing a variable stored by a numeric code, we will often want to convert it into the values the codes represent. 
```{r}
library(tidyverse)

cps <- read.csv("data/CPSData.csv")
metro <- read.csv("data/MetroAreaCodes.csv")
country <- read.csv("data/CountryCodes.csv")

str(metro)
str(country)
# merge(cps,MetroAreaMap,by.x='MetroAreaCode',by.y='Code',all.x=TRUE) 

cps %<>% left_join(metro, by = c("MetroAreaCode" = "Code")) %>%
  left_join(country, by = c("CountryOfBirthCode" = "Code"))

rm(country,metro)
glimpse(cps)
```
Among interviewees with a value reported for Industry, the most common industry of employment is
```{r}
sort(table(cps$Industry), decreasing = T)[1]
```
Educational and health services 

Which states had the smallest and largest numbers of interviewees?
```{r}
sort(table(cps$State), decreasing = T)[1:3]
sort(table(cps$State), decreasing = F)[1:3]
```

What proportion of interviewees are citizens of the United States?
```{r}
prop.table(table(cps$Citizenship))
prop.table(table(cps$Citizenship))[1] + prop.table(table(cps$Citizenship))[2]
```

The CPS differentiates between race (American Indian, Asian, Black, Pacific Islander, White, Multiracial) nf ethnicity. A  of interviewees are of Hispanic ethnicity, as captured by the `Hispanic` variable. For which races are there at least 250 interviewees in the dataset of Hispanic ethnicity? 
```{r}
cps %>% filter(Hispanic == 1) %>%
  group_by(Race) %>%
  summarise(count = n()) %>%
  filter(count >= 250)
```

Which variables have at least 1 interviewee with a missing (NA) value?
```{r}
summary(cps) # MetroAreaCode, Married, Education, EmploymentStatus, Industry
```

Often when evaluating a new dataset, we try to ID if there is a pattern in the NA's in the dataset. We will try to determine if there is a pattern in the NA's of `Married`. `is.na()` returns a vector of TRUE/FALSE values for whether a variable is missing. Check the NA's of Married, split by Region
```{r}
table(cps$Region,is.na(cps$Married))
#lapply(tapply(cps$Married, cps$Region, is.na), sum)
```

For each possible value of `Region`, `Sex`, and `Citizenship`, there're both interviewees w/ missing *and* non-missing `Married` values. However, `Married` is missing for all interviewees Aged 0-14 and present for all interviewees 15+. This is because the CPS does not ask about marriage status for interviewees 14 and younger.

`MetroAreaCode` is missing if an interviewee does not live in a metropolitan area. Look at how many states had all interviewees living in a metropolitan area and all interviewees living in a non-metropolitan area (i.e. has all missing `MetroAreaCode` values, and for this, treat DC as a state, even though it is not technically a state)
```{r}
#table(cps$State,is.na(cps$MetroAreaCode))
#prop.table(table(cps$Region,is.na(cps$MetroAreaCode)))
cps %>%
  group_by(State) %>%
  summarise(prop_metro = mean(is.na(MetroAreaCode) > 0, na.rm = F)) %>%
  filter(prop_metro == 0)

cps %>%
  group_by(State) %>%
  summarise(prop_metro = mean(is.na(MetroAreaCode) > 0, na.rm = F)) %>%
  filter(prop_metro == 1)
```

Which region of the US has the largest proportion of interviewees living in a non-metropolitan area?
```{r}
#prop.table(table(cps$Region,is.na(cps$MetroAreaCode)))
cps %>%
  group_by(Region) %>%
  summarise(prop_metro = mean(is.na(MetroAreaCode) > 0, na.rm = F))
```

While able to use `table()` to compute the proportion of interviewees from each region not living in a metro area, it was tedious (manually computing proportion for each region) and isn't something you'd want to do if there were a larger number of options. There is a less tedious way to compute the proportion of values that are `TRUE`: use `mean()`, which will treat `TRUE` = 1 and `FALSE` = 0, meaning it returns the proportion of values that are true. 
Which state has a proportion of interviewees living in a non-metropolitan area closest to 30%?
```{r}
cps %>%
  group_by(State) %>%
  summarise(prop_metro = mean(is.na(MetroAreaCode) > 0, na.rm = F)) %>%
  filter(prop_metro > .25 & prop_metro < .35)
```

Which state has the largest proportion of non-metropolitan interviewees, ignoring states where all interviewees were non-metropolitan?
```{r}
cps %>%
  group_by(State) %>%
  summarise(prop_metro = mean(is.na(MetroAreaCode) > 0, na.rm = F)) %>%
  filter(prop_metro > .75 & prop_metro != 1)
```

Which of the following metropolitan areas has the largest number of interviewees?
```{r}
cps %>%
  group_by(MetroArea) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  filter(is.na(MetroArea) == 0)
```
summary(cps$MetroArea)

Which metro area has the highest proportion of Hispanic interviewees 
```{R}
cps %>%
  group_by(MetroArea) %>%
  summarise(prop_hispanic = mean(Hispanic > 0, na.rm = T)) %>%
  arrange(desc(prop_hispanic)) 
# sort(tapply(cps$Hispanic,cps$MetroArea,mean)) --> Laredo, TX
```

Determine the number of metro areas in the US from which at least 20% of interviewees are Asian.
```{r}
cps %>%
  group_by(MetroArea) %>%
  summarise(prop_asian = mean(if_else(Race == "Asian",1,0) > 0, na.rm = T)) %>%
  filter(prop_asian >= .2) 
```

Normally, we'd look at the sorted proportion of interviewees from each metro without a high school diploma, however, we know no interviewees <= the age of 14 have an education value reported, so mean = `NA` for each metro, so remove `NA`'s.
```{R}
cps %>%
  group_by(MetroArea) %>%
  summarise(prop_hs = mean(if_else(Education == "No high school diploma",1,0) > 0, 
                           na.rm = T)) %>%
  arrange(prop_hs) %>%
  head()
```

Among all interviewees born outside of N. America, which birth country was the most common? 
```{r}
cps %>%
  group_by(Country) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head()
```

What proportion of interviewees from "New York-Northern New Jersey-Long Island, NY-NJ-PA" metro have a country of birth that is NOT the US? 
```{R}
cps %>% filter(MetroArea == "New York-Northern New Jersey-Long Island, NY-NJ-PA") %>%
  summarise(prop_non_us = mean(if_else(Country == "United States",0,1) > 0, 
                               na.rm = T)) %>%
  arrange(prop_non_us) %>%
  head()
```

Which metro has the largest number of interviewees with country of birth = India, then Brazil, then Somalia? 
```{r}
cps2 <- cps %>% filter(Country %in% c("India", "Brazil", "Somalia")) %>%
  group_by(MetroArea, Country) %>%
  summarise(count = n())

# sort(tapply(cps$Country=="India",cps$MetroArea,sum,na.rm=TRUE)) 
cps2 %>%  arrange(desc(count)) %>%
  filter(Country == "India") %>%
  head(1)

cps2 %>%  arrange(desc(count)) %>%
  filter(Country == "Brazil",
         is.na(MetroArea) == F) %>%
  head(1)

cps2 %>%  arrange(desc(count)) %>%
  filter(Country == "Somalia") %>%
  head(1)
```