---
title: "Wk3 - Mult. Linear Reg."
author: "Steve Newns"
output: html_document
---

Predict book weight by vol + cover type
```{r}
library(DAAG)
data(allbacks)

model1 <- lm(weight ~ volume + cover, data = allbacks)
summary(model1)
```

Note only 1 level of `cover` is in the output = the **NON-reference level** --> hardcover books must be the **reference level**

Multiple R2 = 92.75% --> 92.75% of variability of book weights is explained by volume + cover type

\[
  \hat{y} = 197.96284 + 0.71795 \times volume - 184.04727 \times cover
\]

Remember PB =  non-reference level, meaning HC books = reference level 
<ul>
<li> For reference levels, always plug in 0 in our linear model. </li>
<li> To simplify this linear model to see what it'd look like only for HC books, plug in a 0 for cover. </li>
<li> For a PB, plug in 1 for cover. </li>
</ul>

```{r}
reference_model <- function(x1) {
  197.96284 + .71795*x1
} 

non_reference_model <- function(x1) {
  197.96284 + .71795*x1 - 184.04727
} 

# Predict weight of a PB book 600 cm^3 in volume. 
non_reference_model(600)
```