---
title: "Multiple linear regression"
author: "Steve Newns"
output: html_document
---

## Grading the professor

Many college courses conclude by giving students the opportunity to evaluate the course + the instructor anonymously. However, the use of student evals as an indicator of course quality + teaching effectiveness is often criticized b/c these measures may reflect the influence of non-teaching related characteristics, such as physical appearance of the instructor. The article titled, "Beauty in the classroom: instructors' pulchritude and putative pedagogical productivity" [@Hamermesh2005] found that instructors who are viewed to be better looking receive higher instructional ratings. 

### Load packages

```{r load-packages, message=FALSE}
library(statsr)
library(tidyverse)
library(ggplot2)
library(GGally) ## ggpairs()
```

### The data

The data were gathered from end of semester student evaluations for a large sample of professors from the University of Texas at Austin. In addition, 6 students rated the professors' physical appearance. (This is a slightly modified version of the original data set that was released as part of the replication 
data for *Data Analysis Using Regression and Multilevel/Hierarchical Models* [@Gelman2007].) The result is a data frame where each row contains a different course and columns represent variables about the courses + professors.

Load the data:

```{r load-data, message=FALSE}
data(evals)
glimpse(evals)
```

variable         | description
---------------- | -----------
`score`          | average professor evaluation score: (1) very unsatisfactory - (5) excellent.
`rank`           | rank of professor: teaching, tenure track, tenured.
`ethnicity`      | ethnicity of professor: not minority, minority.
`gender`         | gender of professor: female, male.
`language`       | language of school where professor received education: english or non-english.
`age`            | age of professor.
`cls_perc_eval`  | percent of students in class who completed evaluation.
`cls_did_eval`   | number of students in class who completed evaluation.
`cls_students`   | total number of students in class.
`cls_level`      | class level: lower, upper.
`cls_profs`      | number of professors teaching sections in course in sample: single, multiple.
`cls_credits`    | number of credits of class: one credit (lab, PE, etc.), multi credit.
`bty_f1lower`    | beauty rating of professor from lower level female: (1) lowest - (10) highest.
`bty_f1upper`    | beauty rating of professor from upper level female: (1) lowest - (10) highest.
`bty_f2upper`    | beauty rating of professor from second upper level female: (1) lowest - (10) highest.
`bty_m1lower`    | beauty rating of professor from lower level male: (1) lowest - (10) highest.
`bty_m1upper`    | beauty rating of professor from upper level male: (1) lowest - (10) highest.
`bty_m2upper`    | beauty rating of professor from second upper level male: (1) lowest - (10) highest.
`bty_avg`        | average beauty rating of professor.
`pic_outfit`     | outfit of professor in picture: not formal, formal.
`pic_color`      | color of professor's picture: color, black & white.

1. Is this an observational study or an experiment?
<ol>
<li> **Observational study** </li>
<li> Experiment </li>
</ol>

There are no control + experimental groups, and since this is only an observational study, there cannot be causation between explanatory + response variables, there can only be a correlation.

2. The original research question posed in the paper is whether beauty leads directly to the differences in course evaluations.  Given the study design, should the question be rephrased? If so, how? 
<ol>
<li> No, the question is worded accurately. </li>
<li> **Yes, revise wording to "Is there an association between beauty and course evaluations?"** </li>
<li> Yes, revise wording to "Does beauty score increase the professor's course evaluations?"  </li>
<li> Yes, revise wording to "Does beauty score decrease the professor's course evaluations?" </li>
</ol>

## Exploring the data

3. Which of the following statements is **false** about the distribution of `score`? 
<ol>
<li> The median of the distribution is 4.3. </li>
<li> 25\% of the students gave their professors a score of over 4.6. </li> 
<li> 11 of students gave a professor a score below 3. </li> 
<li> **The left skewness of the data suggests students are less likely to rate professors highly.** </li>
</ol>
```{r score-dist}
hist(evals$score)
summary(evals$score)
```

<div id="exercise">
**Exercise:** Excluding `score`, select two other variables and describe their relationship using an appropriate visualization (scatterplot, side-by-side boxplots, or mosaic plot).
</div>
```{r two-vars-eda}
boxplot(score ~ language, evals)
boxplot(score ~ rank, evals)
```

There looks to be some relationship between eval score and the language of school where a professor received education but possibly a smaller one in professor's rank, with tenure professor's having the lowest ranking.

## Simple linear regression

The fundamental phenomenon suggested by the study is that better looking teachers are evaluated more favorably. Let's create a scatterplot to see if this appears to be the case:

```{r scatter-score-bty_avg}
ggplot(evals, aes(bty_avg, score)) +
  geom_point()
```

Before we draw conclusions about the trend, compare the number of observations in the data frame with the approximate number of points on the scatterplot. Is anything awry?

**There are less DP's in the scatterplot because the points are overlapping.**

<div id="exercise">
**Exercise:** Replot the scatterplot, but this time replace the `geom_point()` layer with a `geom_jitter()` layer.
</div>
```{r jitter}
ggplot(evals, aes(bty_avg, score)) +
  geom_jitter()
```

Let's see if the apparent trend in the plot is something more than natural variation.  Fit a linear model called `m_bty` to predict average professor score by average beauty rating and add the line to your plot using the following.

```{r scatter-score-bty_avg-line}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  geom_smooth(method = "lm")
```

The blue line is the model. The shaded gray area around the line tells us about the variability we might expect in our predictions. To turn that off, use `se = FALSE`.

```{r scatter-score-bty_avg-line-no-se}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE)
```

<div id="exercise">
**Exercise:** Print a summary of the linear model, write out the equation, and interpret the slope. 
</div>
```{r summary-model}
m_bty <- lm(score ~ bty_avg, evals)
summary(m_bty)
```
\[
\widehat{score} = 3.88 + 0.067 \times bty\_avg
\]

4. Average beauty score is a statistically significant predictor of evaluation score. 
<ol> 
<li> **True** </li>
<li> False </li>
</ol>

5. Use residual plots to evaluate whether the conditions of least squares regression are reasonable.  Which of the following statements is an incorrect analysis of the residual plots and conditions? 
<ol>
<li> Linear association: The residuals plot shows a random scatter. </li>
<li> Constant variance of residuals: No fan shape in residuals plot. </li>
<li> Nearly normal residuals: Residuals are right skewed, but the sample size is large, so this may not be an important violation of conditions. </li>
<li> Independent observations: Classes sampled randomly, no order effect. </li>
</ol>
```{r diag-score-bty_avg}
# linearity w/ residuals plot
plot(evals$bty_avg, m_bty$residuals)
abline(0,0)

# normal residuals
hist(m_bty$residuals)
qqnorm(m_bty$residuals)
qqline(m_bty$residuals)

# constant variability of residuals
plot(m_bty$fitted.values, m_bty$residuals)
abline(0,0)
plot(m_bty$fitted.values, abs(m_bty$residuals))

# independence of residuals (+ therfore observations)
plot(m_bty$residuals)
abline(0,0)
```
Looks like all conditions are met except the normality of the residuals.

## Multiple linear regression

The data set contains several variables on the beauty score of the professor: individual ratings from each of the six students who were asked to score the physical appearance of the professors and the average of these six scores. Let's take a look at the relationship between one of these scores and the average beauty score.

```{r bty-rel}
ggplot(evals, aes(bty_f1lower, bty_avg)) +
  geom_jitter()

evals %>% 
  summarise(cor(bty_avg, bty_f1lower))
```

As expected the relationship is quite strong - after all, the average score is calculated using the individual scores. We can actually take a look at the relationships between all beauty variables (columns 13 through 19)

```{r bty-rels, fig.width = 12, fig.height= 5}
# correlation plot/pairwise plot
ggpairs(evals, columns = 13:19)
```

These variables are collinear (correlated), + adding more than one of these variables to the model would not add much value to the model. In this application + with these highly-correlated predictors, it is reasonable to use the average beauty score as the single representative of these variables.

In order to see if beauty is still a significant predictor of professor score after we've accounted for the gender of the professor, we can add `gender` into the model.

```{r scatter-score-bty_avg_gender}
m_bty_gen <- lm(score ~ bty_avg + gender, evals)
summary(m_bty_gen)
```

6. P-values + parameter estimates should only be trusted if the conditions for the regression are reasonable. Using diagnostic plots, we can conclude that the conditions for this model are reasonable. 
<ol>
<li> True </li>
<li> False </li>
</ol>
</div>
```{r diag-mlr-model}
plot(evals$bty_avg, m_bty_gen$residuals)
abline(0,0)
plot(evals$gender, m_bty_gen$residuals)

# normal residuals
hist(m_bty_gen$residuals)
qqnorm(m_bty_gen$residuals)
qqline(m_bty_gen$residuals)

# constant variability of residuals
plot(m_bty_gen$fitted.values, m_bty_gen$residuals)
abline(0,0)
plot(m_bty_gen$fitted.values, abs(m_bty_gen$residuals))

# independence of residuals (+ therfore observations)
plot(m_bty_gen$residuals)
abline(0,0)
```

No, the residuals aren't normal.

<div id="exercise">
**Exercise:** Is `bty_avg`still a significant predictor of `score`?  Has the addition of `gender` to the model changed the parameter estimate for `bty_avg`?
</div>
```{r summary-mlr-model}
summary(m_bty)
summary(m_bty_gen)
```

Yes, it's still significant, and its estimate has increased a bit.

Note that the estimate for `gender` is now called `gendermale`. R recodes the categorical `gender` from having the values of `female` and `male` to being an indicator variable called `gendermale` that takes a value of $0$ for females and a value of $1$ for males. (**"dummy" variables**)

As a result, for females, the parameter estimate is multiplied by zero, leaving the intercept + slope form familiar from simple regression.

\[
  \begin{aligned}
\widehat{score} &= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg + \hat{\beta}_2 \times (0) \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg\end{aligned}
\]

7. For two professors (one male and one female) who received the same beauty rating, the male professor is predicted to have the higher course evaluation score than the female.
<ol>
<li> **True** </li>
<li> False </li>
</ol>

The decision to call the indicator variable `gendermale` instead of`genderfemale` has no deeper meaning. R simply codes the category that comes first alphabetically as a $0$. (can change the reference level of a categorical variable (level coded as a 0) using `relevel()`

<div id="exercise">
**Exercise:** Create a new model `m_bty_rank` with `gender` removed + `rank` added in.  How does R appear to handle categorical variables that have more than 2 levels? 
</div>
```{r new-mlr-model}
m_bty_rank <- lm(score ~ bty_avg + rank, evals)
summary(m_bty_rank)
```

See 1 row for each non-reference level of the categorical `rank`

8. Which of the following is the correct order of the 3 levels of rank if we were to order them from lowest predicted course evaluation score to highest predicted course evaluation score? 
<ol>
<li> Teaching, Tenure Track, Tenured </li>
<li> Tenure track, Tenured </li>
<li> Tenure Track, Tenured, Teaching </li>
<li> **Teaching, Tenured, Tenure Track** </li>
</ol>

The interpretation of the coefficients in multiple regression is slightly different from that of simple regression. The estimate for `bty_avg` reflects how much higher a group of professors is expected to score if they have a beauty rating that 1 point higher *while holding all other variables constant*. In
this case, that translates into considering only professors of the same rankwith `bty_avg` scores that are 1 point apart.

## Prediction

Suppose we want to use `m_bty_gen` to predict evaluation score for a male tenure track professor with an average beauty of 3.

Calculate the predicted value in R.

First, create a new data frame for this professor.
```{r new-prof}
newprof <- data.frame(gender = "male", bty_avg = 3)
```

Then, do prediction using `predict(model, new data)` 
```{r new-prof-predict}
predict(m_bty_gen, newprof)
```

We can also construct a **prediction interval** (wider than CI) around this prediction, which provides a measure of uncertainty around the prediction.
```{r new-prof-predict-interval}
predict(m_bty_gen, newprof, interval = "prediction", level = 0.95)
```

Hence, the model predicts, with 95% confidence, that a male professor with an average beauty score of 3 is expected to have an evaluation score between 3.1 and 5.18. 

## The search for the best model

We will start with a full model that predicts professor score based on rank, ethnicity, gender, language of the university where they got their degree, age, proportion of students that filled out evaluations, class size, course level, number of professors, number of credits, average beauty rating, outfit, and 
picture color.

<div id="exercise">
Which variable would you expect to have the highest p-value in this model? Why? *Hint:* Think about which variable would you expect to not have any association with the professor score.
</div>

Let's run the model...

```{r m_full, tidy = FALSE}
m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```

9. Which of the following is the correct intrepetation of the coefficient associated with `ethnicity`  <br> Non-minority professors are expected on average to score ... 
<ol>
<li> 0.12 points lower than minority professors, all else held constant. </li>
<li> **0.12 points higher than minority professors, all else held constant.** </li> 
<li> 0.02 points lower than minority professors, all else held constant. </li>
<li> 0.02 points higher than minority professors, all else held constant. </li>
</ol>

<div id="exercise">
**Exercise:** Drop the variable with the highest p-value and re-fit the model. Did the coefficients and significance of the other explanatory variables change? 
</div>
```{r p-val-select}
m_drop_p <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_drop_p)
```

1 of the things that makes multiple regression interesting is coefficient estimates depend on the other variables included in the model. What does the above say about whether or not the dropped variable was collinear with the other explanatory variables?

**It was not, as significance of all other variables did not change, even though coefficients did.**


Now we try a different model selection method: adjusted $R^2$. Create a new model, `m1`, where you remove `rank` from the list of explanatory variables. Check out the adjusted $R^2$ of this new model and compare it to the adjusted $R^2$ of the full model.

```{r eval=FALSE}
m1 <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg, data = evals)

summary(m1)$adj.r.squared
summary(m_full)$adj.r.squared
```

Then, try dropping the next variable from the full model (`ethnicity`):

```{r eval=FALSE}
m2 = lm(score ~ rank + gender + language + age + cls_perc_eval + 
    cls_students + cls_level + cls_profs + cls_credits + bty_avg, data = evals)

summary(m2)$adj.r.squared
summary(m1)$adj.r.squared
summary(m_full)$adj.r.squared
```

<div id="exercise">
**Exercise:** Repeat this process until you have tried removing each variable from the full model at a time, and determine removal of which variable yields the highest improvement in the adjusted $R^2$.
</div>
```{r adj-rsq-select}
m3 <- lm(score ~ ethnicity + rank + language + age + cls_perc_eval +
           cls_students + cls_level + cls_profs + cls_credits + bty_avg, data = evals)
m4 <- lm(score ~ ethnicity + rank + gender + age + cls_perc_eval +
           cls_students + cls_level + cls_profs + cls_credits + bty_avg, data = evals)
m5 <- lm(score ~ ethnicity + rank + gender + language + cls_perc_eval + 
           cls_students + cls_level + cls_profs + cls_credits + bty_avg, data = evals)
m6 <- lm(score ~ ethnicity + rank + gender + language + age + cls_students + 
           cls_level + cls_profs + cls_credits + bty_avg, data = evals)
m7 <- lm(score ~ ethnicity + rank + gender + language + age + cls_perc_eval + 
           cls_level + cls_profs + cls_credits + bty_avg, data = evals)
m8 <- lm(score ~ ethnicity + rank + gender + language + age + cls_perc_eval + 
           cls_students + cls_profs + cls_credits + bty_avg, data = evals)
m9 <- lm(score ~ ethnicity + rank + gender + language + age + cls_perc_eval +
           cls_students + cls_level + cls_credits + bty_avg, data = evals)
m10 <- lm(score ~ ethnicity + rank + gender + language + age + cls_perc_eval + 
            cls_students + cls_level + cls_profs + bty_avg, data = evals)
m11 <- lm(score ~ ethnicity + rank + language + age + cls_perc_eval +
           cls_students + cls_level + cls_profs + cls_credits, data = evals)

summary(m_full)$adj.r.squared
summary(m1)$adj.r.squared
summary(m2)$adj.r.squared
summary(m3)$adj.r.squared
summary(m4)$adj.r.squared
summary(m5)$adj.r.squared
summary(m6)$adj.r.squared
summary(m7)$adj.r.squared
summary(m8)$adj.r.squared
summary(m9)$adj.r.squared
summary(m10)$adj.r.squared
summary(m11)$adj.r.squared
```

10. Elimination of which variable from the full model yielded the highest adjusted R-squared? 
<ol>
<li> `bty_avg` </li>
<li> **`cls_profs`** </li> 
<li> `cls_students` </li>
<li> `rank` </li>
</ol>

To complete the model selection we would continue removing variables one at a time until removal of another variable did not increase adjusted $R^2$.

```{r adj-rsq-select}
m9 <- lm(score ~ ethnicity + rank + gender + language + age + cls_perc_eval +
           cls_students + cls_level + cls_credits + bty_avg, data = evals)
m12 <- lm(score ~ rank + gender + language + age + cls_perc_eval +
           cls_students + cls_level + cls_credits + bty_avg, data = evals)
m13 <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval +
           cls_students + cls_level + cls_credits + bty_avg, data = evals)
m14 <- lm(score ~ ethnicity + rank + language + age + cls_perc_eval +
           cls_students + cls_level + cls_credits + bty_avg, data = evals)
m15 <- lm(score ~ ethnicity + rank + gender + age + cls_perc_eval +
           cls_students + cls_level + cls_credits + bty_avg, data = evals)
m16 <- lm(score ~ ethnicity + rank + gender + language + cls_perc_eval +
           cls_students + cls_level + cls_credits + bty_avg, data = evals)
m17 <- lm(score ~ ethnicity + rank + gender + language + age +
           cls_students + cls_level + cls_credits + bty_avg, data = evals)
m18 <- lm(score ~ ethnicity + rank + gender + language + age + cls_perc_eval +
          cls_level + cls_credits + bty_avg, data = evals)
m19 <- lm(score ~ ethnicity + rank + gender + language + age + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
m20 <- lm(score ~ ethnicity + rank + gender + language + age + cls_perc_eval +
           cls_students + cls_level + cls_credits, data = evals)

summary(m9)$adj.r.squared
summary(m12)$adj.r.squared
summary(m13)$adj.r.squared
summary(m14)$adj.r.squared
summary(m15)$adj.r.squared
summary(m16)$adj.r.squared
summary(m17)$adj.r.squared
summary(m18)$adj.r.squared
summary(m19)$adj.r.squared
summary(m20)$adj.r.squared
```

```{r}
m19 <- lm(score ~ ethnicity + rank + gender + language + age + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
m21 <- lm(score ~ rank + gender + language + age + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
m22 <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
m23 <- lm(score ~ ethnicity + rank + language + age + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
m24 <- lm(score ~ ethnicity + rank + gender + age + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
m25 <- lm(score ~ ethnicity + rank + gender + language + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
m26 <- lm(score ~ ethnicity + rank + gender + language + age +
           cls_students + cls_credits + bty_avg, data = evals)
m27 <- lm(score ~ ethnicity + rank + gender + language + age + cls_perc_eval +
           cls_credits + bty_avg, data = evals)
m28 <- lm(score ~ ethnicity + rank + gender + language + age + cls_perc_eval +
           cls_students + bty_avg, data = evals)
m29 <- lm(score ~ ethnicity + rank + gender + language + age + cls_perc_eval +
           cls_students + cls_credits, data = evals)

summary(m19)$adj.r.squared
summary(m21)$adj.r.squared
summary(m22)$adj.r.squared
summary(m23)$adj.r.squared
summary(m24)$adj.r.squared
summary(m25)$adj.r.squared
summary(m26)$adj.r.squared
summary(m27)$adj.r.squared
summary(m28)$adj.r.squared
summary(m29)$adj.r.squared
```

```{r}
m22 <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
m30 <- lm(score ~ gender + language + age + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
m31 <- lm(score ~ ethnicity + language + age + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
m32 <- lm(score ~ ethnicity + gender + age + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
m33 <- lm(score ~ ethnicity + gender + language + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
m34 <- lm(score ~ ethnicity + gender + language + age + 
           cls_students + cls_credits + bty_avg, data = evals)
m35 <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval +
           cls_credits + bty_avg, data = evals)
m36 <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval +
           cls_students + bty_avg, data = evals)
m37 <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval +
           cls_students + cls_credits, data = evals)

summary(m22)$adj.r.squared
summary(m30)$adj.r.squared
summary(m31)$adj.r.squared
summary(m32)$adj.r.squared
summary(m33)$adj.r.squared
summary(m34)$adj.r.squared
summary(m35)$adj.r.squared
summary(m36)$adj.r.squared
summary(m37)$adj.r.squared
```
```{r}
best_model <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval +
           cls_students + cls_credits + bty_avg, data = evals)
summary(best_model)
```

<div id="exercise">
**Exercise:** The original paper describes how these data were gathered by taking a sample of professors from the University of Texas at Austin and including all courses that they have taught. Considering that each row represents a course, could this new information have an impact on any of the conditions of linear regression?
</div>

**Yes, since they could be non-independent observations**

<div id="exercise">
**Exercise:** Based on your final model, describe the characteristics of a professor and course at University of Texas at Austin that would be associated with a high evaluation score.
</div>

** A younger, non-minority (white), male professor with a degree from an english-speaking school who taught larger, one-credit classes with more participation in the survey who was more attractive.** s

<div id="exercise">
**Exercise:** Would you be comfortable generalizing your conclusions to apply to professors generally (at any university)?  Why or why not?
</div>

**No, because this is an observational study**