---
title: "Wk1 - Bayes' Rule"
author: "Steve Newns"
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2)
library(reshape2)

tbl <- matrix(c(60,86,58,21,255,426,450,382), 2, 4, byrow=TRUE)
colnames(tbl) <- c('18-29', '30-49', '50-64','65+')
rownames(tbl) <- c('Used online dating', 'No online dating')
tbl
```

Probability of using online dating sites given 30-49 years old.
```{r}
p_a <- sum(tbl[1]) / sum(tbl) # probability of using online dating
p_b <- sum(tbl[,2]) / sum(tbl) # probability of being 30-49
p_ab <- tbl[1,2] / sum(tbl) # probability of being 30-49 and using online dating

(p_AGivenB <- p_ab / p_b)
```

# Military
```{r}
prob_posGivenHIV_elisa <- .93 # TP rate = sensitivity of 1st ELISA
prob_negGivenNoHIV_elisa <- .99 # TN rate = specificity of 1st ELISA
prob_posGivenHIV_wb <- .999
prob_negGivenNoHIV_wb <- .991
prob_HIV <- 1.48 / 1000 # prevalence
prob_noHIV <- 1 - prob_HIV
```

Probability a recruit has HIV given positive on the 1st ELISA

```{r}
prob_negGivenHIV_elisa <- 1 - prob_posGivenHIV_elisa 
prob_posGivenNoHIV_elisa <- 1 - prob_negGivenNoHIV_elisa
 
prob_posAndHIV_elisa <- prob_HIV * prob_posGivenHIV_elisa
prob_posAndNoHIV_elisa <- prob_noHIV * prob_posGivenNoHIV_elisa

prob_pos_elisa1 <- prob_posAndHIV_elisa + prob_posAndNoHIV_elisa

(prob_HIVGivenPos_elisa1 <- prob_posAndHIV_elisa / prob_pos_elisa1)
```

# Bayes Updating
    
What is probability of having HIV if the 2nd ELISA also yields a positive result? 

Update our prior probability w/ a posterior
```{r}
post_probHIV <- prob_HIVGivenPos_elisa1
post_probNoHIV <- 1 - post_probHIV

prob_posAndHIV_elisa2 <- post_probHIV * prob_posGivenHIV_elisa
prob_posAndNoHIV_elisa2 <- post_probNoHIV * prob_posGivenNoHIV_elisa

prob_pos_elisa2 <- prob_posAndHIV_elisa2 + prob_posAndNoHIV_elisa2

(prob_HIVGivenPos_elis2 <- prob_posAndHIV_elisa2 / prob_pos_elisa2)
```

Do a third time
```{r}
post_probHIV <- prob_HIVGivenPos_elis2
post_probNoHIV <- 1 - post_probHIV

prob_posAndHIV_elisa3 <- post_probHIV * prob_posGivenHIV_elisa
prob_posAndNoHIV_elisa3 <- post_probNoHIV * prob_posGivenNoHIV_elisa

prob_pos_elisa3 <- prob_posAndHIV_elisa3 + prob_posAndNoHIV_elisa3

(prob_HIVGivenPos_elis3 <- prob_posAndHIV_elisa3 / prob_pos_elisa3)
```

# frequentist inference for a proportion
```{r}
k <- 4
n <- 20
sum(dbinom(0:k,n,.5))
```

# bayesian inference for proportion
```{r}
p_values <- seq(.1,.9,.1)
priors <- c(rep(.06,4), .52, rep(.6,4))
(likelihood <- dbinom(4, 20, prob = p_values))
```
```{r}
(posteriors <- (priors * likelihood) / sum((priors * likelihood)))
sum(posteriors)
```


# M&M's

```{r}
k <- 1
n <- 5
p_values <- seq(.1,.2,.1)
priors <- c(rep(.5,2))
(likelihood <- dbinom(k, n, prob = p_values))
```
```{r}
(posteriors <- (priors * likelihood) / sum((priors * likelihood)))
sum(posteriors) # check they sum to 1
```

# larger n
```{r}
k <- 2
n <- 10
p_values <- seq(.1,.2,.1)
priors <- c(rep(.5,2))
(likelihood <- dbinom(k, n, prob = p_values))
(posteriors <- (priors * likelihood) / sum((priors * likelihood)))
sum(posteriors) # check they sum to 1
```

```{r}
k <- 3
n <- 15
p_values <- seq(.1,.2,.1)
priors <- c(rep(.5,2))
(likelihood <- dbinom(k, n, prob = p_values))
(posteriors <- (priors * likelihood) / sum((priors * likelihood)))
sum(posteriors) # check they sum to 1
```


```{r}
k <- 4
n <- 20
p_values <- seq(.1,.2,.1)
priors <- c(rep(.5,2))
(likelihood <- dbinom(k, n, prob = p_values))
(posteriors <- (priors * likelihood) / sum((priors * likelihood)))
sum(posteriors) # check they sum to 1
```

# Quiz

```{r}
p_rain <- 50/365
p_no_rain <- 1 - p_rain

p_correct_given_rain <- .8
p_correct_given_no_rain <- .3

p_rain_and_correct <- p_rain*p_correct_given_rain
p_no_rain_and_correct <- p_no_rain*p_correct_given_no_rain
p_correct <-p_rain_and_correct + p_no_rain_and_correct


(p_rain_giv_pred <- p_rain_and_correct_rain / p_correct)
```

```{r}
k <- 3
n <- 3
p_values <- c(.2,.4,.5,.6,.8)
priors <- c(rep(.5/4,2),.5,rep(.5/4,2))
(likelihood <- dbinom(k, n, prob = p_values))
(posteriors <- (priors * likelihood) / sum((priors * likelihood)))
sum(posteriors) # check they sum to 1
sum(posteriors[4:5])
```


```{r}
binom.test(x=9942, n=10000, p=.993, alternative="greater")
```

```{r}
k <- 10
n <- 16
priors_did_not_make_playoff <- c(20/32,6/32,4/32,2/32)
(likelihood <- dbinom(k, n, prob = .5)) # prob of winning coint toss
(posteriors <- (priors_did_not_make_playoff * likelihood) / sum((priors_did_not_make_playoff * likelihood)))
sum(posteriors)
```

```{r}
k <- 0
n <- 10
p_values <- c(1/1000,1/1000000)
priors <- c(.5,.5)
(likelihood <- dbinom(k, n, prob = p_values))
(posteriors <- (priors * likelihood) / sum((priors * likelihood)))
sum(posteriors)
```

```{r}
k <- 999
n <- 6000
p_values <- c(1/6,.175)
priors <- c(.8,.2)
#(likelihood <- dbinom(k, n, prob = p_values))
#(posteriors <- (priors * likelihood) / sum((priors * likelihood)))
#sum(posteriors)
1 - pbinom(q = 999-1, size = 999, .2)
#1-pbinom(999-1,n,priors[1])

```