---
title: "Exploring the BRFSS data"
output: 
  html_document: 
  fig_height: 4
highlight: pygments
theme: spacelab
---
  
## Setup
  
### Load packages
  
```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(maps)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```



* * *
  
  ## Part 1: Data
  
  
  * * *
  
  ## Part 2: Research questions
  
  **Research quesion 1: What are the US states with highest proportion of smoking who smoke every day relative to all those respondents who smoke?**
  
  **Research quesion 2: What percentage of non-retired, employed college graduates and non-retired non-college graduates make more than $50,000 per year?**
  
  **Research quesion 3: What are the most popular types of exercise for individuals who have no bad mental health days in the past 30 days and those who have more than 15 bad mental health days in the past 30 days?**
  
  
  * * *
  
  ## Part 3: Exploratory data analysis
  
  NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**
  
1st, let's look at the number of records in each state.
  
```{r state counts} 
# get count observations from each states
brfss2013 %>%
  group_by(X_state) %>%
  summarise(count = n())
```

So there's a state "0"? That can't be right, but it might just be some kind of header. I'll come back to that later.

Now to look at the frequency table of the **smokday2** variable, which is a description of the frequency of days a respondent is now smoking, ranging from "not at all"" to "every day".

```{r smoke frequency}
summary(brfss2013$smokday2)
```

It looks like the majority of people do not smoke, which is good. But there's a lot of NA values in this dataset for this variable. Let's get only those who answered the smoking question into a subset titled **smoking**.

```{r smoking answered, message = FALSE}
# get only those who answered frequency of smoking question
smoking <- brfss2013 %>%
  filter(!is.na(smokday2))
```

Now I need to get the proportion of those who smoke every day by taking the proportion of the "Every day" response and filtering down to just those respondents. Then, I can look at those states with the highest proportion of those who smoke every day.

```{r smoke proportion}
# get proportion of each level of smoking by state
smoking %>%
  group_by(X_state, smokday2) %>%
  summarise(count = n()) %>%
  mutate(prop = count / sum(count)) %>%
  filter(smokday2 == "Every day") %>%
  arrange(desc(prop))
```

It seems Guam has the highest proportion of people who say the smoke every day. According to the codebook for this dataset, Guam and Puerto Rico are includedbecause they are US territories. Now I want to check if there are any other options for X_state other than US states and territories.

```{r check states}
levels(smoking$X_state)
```

Okay, so it looks like I need to remove Guam and Puerto Rico, the US territories. I also need to remove the "0" and the "80". I'll keep the District of Columbia even though it's not really a *state*.

I can create a function to do the opposite of **%in%** to remove those records with the above values for X_state, then get the proportions of those who smoke every day relative to every respondent to the question and then arranging the results by descending proportion of those who smoke every day.

```{r remove states}
# function to be opposite of %IN%
`%not in%` <- function (x, table) is.na(match(x, table, nomatch=NA_integer_))

# keep only us official states
(top_every_day_smoker_states <- smoking %>% 
    filter(X_state %not in% c("0", "80", "Guam", "Puerto Rico")) %>%
    group_by(X_state, smokday2) %>%
    summarise(count = n()) %>%
    mutate(prop = count / sum(count)) %>%
    filter(smokday2 == "Every day") %>%
    arrange(desc(prop)))
```

I now want to create a heatmap to see which states have the highest proportion of those who smoke every day.

```{r heat map}
top_every_day_smoker_states$region <- tolower(top_every_day_smoker_states$X_state)

states <- map_data("state")
map.df <- merge(states,top_every_day_smoker_states, by = "region", all.x = T)
map.df <- map.df[order(map.df$order),]
ggplot(map.df, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill=prop)) +
  geom_path() + 
  scale_fill_gradientn(colours = rev(heat.colors(10)), na.value="grey90") +
  coord_map() + 
  theme(rect = element_blank(), axis.text.x = element_blank(), 
        axis.text.y = element_blank(), axis.ticks = element_blank()) + # remove grid background and ticks and tick labels
  ggtitle("Heatmap of the Proportion of Everyday Smokers in the United States") + 
  labs(caption="* Data reflects statistics for those who have responses for the variable smokday2: 54516 respondents", 
       x = element_blank(), y = element_blank())
  
```


Now I want to visualize the proportions of the top 10 states with the highest proportion of everyday smokers.

```{r}
#top 15 states
top_every_day_smoker_states %>%
  head(10) %>%
  ggplot() + 
  geom_bar(aes(reorder(X_state, -prop), prop, fill = "Red"), stat = "identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  xlab('State') + 
  ylab('Proportion') + 
  ggtitle('Top 10 State Proportions of Every Day Smoking') + 
  guides(fill=FALSE)
```

Let's also look at the these proportions along those who don't smoke at all and those who smoke some days for these same states.

```{r}
#get those top 15 states
top_every_day_states <- top_every_day_smoker_states %>%
  head(10) %>% 
  select(X_state) %>%
  as.data.frame() #%>%

# array of character elements of the state names
top_every_day_states <- top_every_day_states[,1]

smoking %>%
  group_by(X_state, smokday2) %>%
  summarise(count = n()) %>%
  mutate(prop = count / sum(count)) %>%
  arrange(desc(prop)) %>%
  filter(X_state %in% top_every_day_states) %>% # get only top 10 states
  ggplot() + 
  geom_bar(aes(X_state, prop, fill = smokday2), stat = "identity") + # stacked bar chart
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  xlab("State") + 
  ylab("Proportion") + 
  ggtitle("Proportion of Smoker Frequency") + 
  guides(fill=guide_legend(title="Frequency"))
```



**Research quesion 2:**

Now I want to see if there is a discrepency in the proportions of the salaries of both college graduates and non-college graduates compared to the median household salary in the US. First, I need to filter out the US territories like above and also filter down to just college gradutes to start with.
  
```{r}
united.states <- brfss2013 %>%
  filter(X_state %not in% c("0", "80", "Guam", "Puerto Rico"))

united.states %>%
  filter(educa == "College 4 years or more (College graduate)") %>%
  count
```

We end up with about 168,000 observations. I also want to filter down to those that are actually working, which would only include those respondents that are employed for wages or are self-employed.
```{r working grads}
us_college_grads_not_retired <- brfss2013 %>%
  filter(educa == "College 4 years or more (College graduate)",
         employ1 %in% c("Employed for wages","Self-employed"))
```

I also know the income level has been put into buckets that have levels that are character strings, such as "Less than $20,000". I want to recode that into numerics so I can rank them and then look at a histogram of the distribution of the incomes of non-retired, working college graduates in the United States.

```{r recode}
# rename factors to numerics
us_college_grads_not_retired$incomeRank <- recode(us_college_grads_not_retired$income2, "Less than $10,000" = 1, "Less than $15,000" = 2,
                                                                                "Less than $20,000" = 3, "Less than $25,000" = 4,
                                                                                "Less than $35,000" = 5, "Less than $50,000" = 6,
                                                                                "Less than $75,000" = 7,"$75,000 or more" = 8)
ggplot(us_college_grads_not_retired) + 
  geom_histogram(aes(as.numeric(incomeRank)), binwidth = 1)
```

# percent of people making more than 50k
college_grads_not_retired %>%
  summarize(prop = mean(as.numeric(incomeRank) > 6, na.rm = T))
#0.7699314 = 77%




non_grads_not_retired <- brfss2013 %>%
  filter(educa %not in% c("College 4 years or more (College graduate)","Refused"),
         employ1 %in% c("Employed for wages","Self-employed"))
table(non_grads_not_retired$educa)

# rename factors to numerics
non_grads_not_retired$incomeRank <- recode(non_grads_not_retired$income2, "Less than $10,000" = 1, "Less than $15,000" = 2,
                                               "Less than $20,000" = 3, "Less than $25,000" = 4,
                                               "Less than $35,000" = 5, "Less than $50,000" = 6,
                                               "Less than $75,000" = 7,"$75,000 or more" = 8)
ggplot(non_grads_not_retired) + 
  geom_histogram(aes(as.numeric(incomeRank)), binwidth = 1)

# percent of people making more than 50k
non_grads_not_retired %>%
  summarize(prop = mean(as.numeric(incomeRank) > 6, na.rm = T))
#0.4310216 = 43%
```



**Research quesion 3:**
  
  ```{r}

```
