---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(maps)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data


* * *

## Part 2: Research questions

**Research quesion 1: What are the US states with highest proportion of smokers who smoke every day relative to all those respondents who smoke?**

**Research quesion 2: What percentage of non-retired, employed college graduates and non-retired non-college graduates make more than $50,000 per year?**

**Research quesion 3: What are the most popular types of exercise for individuals who have no bad mental health days and those who have bad mental health days?**


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

```{r}

```



**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```
