---
  title: "Exploring the BRFSS data"
output: 
  html_document: 
  fig_height: 4
highlight: pygments
theme: spacelab
---
  
## Setup
  
### Load packages
  
```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(maps)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```



* * *
  
  ## Part 1: Data
  
  
  * * *
  
  ## Part 2: Research questions
  
  **Research quesion 1: What are the US states with highest proportion of smoking who smoke every day relative to all those respondents who smoke?**
  
  **Research quesion 2: What percentage of non-retired, employed college graduates and non-retired non-college graduates make more than $50,000 per year?**
  
  **Research quesion 3: What are the most popular types of exercise for individuals who have no bad mental health days in the past 30 days and those who have more than 15 bad mental health days in the past 30 days?**
  
  
  * * *
  
  ## Part 3: Exploratory data analysis
  
  NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**
  
  ```{r state counts} 
# get count observations from each states
brfss2013 %>%
  group_by(X_state) %>%
  summarise(count = n())
```

So there's a state "0"? That can't be right, but it might just be some kind of header.

Now to look at the frequency table of the **smokday2** variable, which is a description of the frequency of days a respondent is now smoking, ranging from not at all to every day.

```{r smoke frequency}
summary(brfss2013$smokday2)
```

It looks like the majority of people do not smoke, which is good. But there's a lot of NA values in this dataset for this variable. Let's get only those who answered the smoking question into a subset titled **smoking**.

```{r smoking answered}
# get only those who answered frequency of smoking question
smoking <- brfss2013 %>%
  filter(!is.na(smokday2))
```

Now I need to get the proportion of those who smoke every day by taking the proportion of the "Every day" response and filtering down to just those respondents. Then, I can look at those states with the highest proportion of those who smoke every day.

```{r smoke proportion}
# get proportion of each level of smoking by state
smoking %>%
  group_by(X_state, smokday2) %>%
  summarise(count = n()) %>%
  mutate(prop = count / sum(count)) %>%
  filter(smokday2 == "Every day") %>%
  arrange(desc(prop))
```

It seems Guam is the highest proportion. This is because it is a US territory. Now I want to check if there are any other options for X_state other than US states.

```{r check states}
levels(smoking$X_state)
```

Okay, so it looks like I need to remove Guam and Puerto Rico, the US territories. I also need to remove the "0" and the "80".

I can create a function to do the opposite of **%in%** to remove those records with the above values for X_state, then get the proportions of those who smoke every day relative to every respondent to the question and then arranging the results by descending proportion of those who smoke every day.

```{r remove states}
# function to be opposite of %IN%
`%not in%` <- function (x, table) is.na(match(x, table, nomatch=NA_integer_))

# keep only us official states
(top_every_day_smoker_states <- smoking %>% 
    filter(X_state %not in% c("0", "80", "Guam", "Puerto Rico")) %>%
    group_by(X_state, smokday2) %>%
    summarise(count = n()) %>%
    mutate(prop = count / sum(count)) %>%
    filter(smokday2 == "Every day") %>%
    arrange(desc(prop)))
```

I now want to create a heatmap to see which states have the highest proportion of those who smoke every day.
```{r heat map}
top_every_day_smoker_states$region <- tolower(top_every_day_smoker_states$X_state)

states <- map_data("state")
map.df <- merge(states,top_every_day_smoker_states, by="region", all.x=T)
map.df <- map.df[order(map.df$order),]
ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=prop))+
  geom_path()+ 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+
  coord_map()
```


Now I want to see the proportions of the top 10 states with the highest proportion of everyday smokers
```{r}
#top 15 states
top_every_day_smoker_states %>%
  head(10) %>%
  ggplot() + 
  geom_bar(aes(reorder(X_state, -prop), prop, fill = "Red"), stat = "identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  xlab('State') + 
  ylab('Proportion') + 
  ggtitle('Proportion of Every Day Smoking') + 
  guides(fill=FALSE)
```

Let's also look at the the proportions of those who don't smoke at all and those who smoke some days for these states.

```{r}
#get those top 15 states
top_every_day_states <- top_every_day_smoker_states %>%
  head(10) %>% 
  select(X_state) %>%
  as.data.frame() #%>%

top_every_day_states <- top_every_day_states[,1]

#'West Virginia' %in% top_every_day_states

smoking %>%
  group_by(X_state, smokday2) %>%
  summarise(count = n()) %>%
  mutate(prop = count / sum(count)) %>%
  arrange(desc(prop)) %>%
  #head(10) %>%
  filter(X_state %in% top_every_day_states) %>%
  ggplot() + 
  geom_bar(aes(X_state, prop, fill = smokday2), stat = "identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  xlab("State") + 
  ylab("Proportion") + 
  ggtitle("Proportion of Different Smokers")
```



**Research quesion 2:**
  
  ```{r}

```



**Research quesion 3:**
  
  ```{r}

```
