---
title: "K-means Clustering"
author: "Steve Newns"
output: html_document
---

# Kernel SVM

```{r, warning=F,message=F}
library(tidyverse)
library(ggplot2)
library(cluster) # cluster viz
```

Mall has data on clients who have a membership card with info on age, gender, and annual income. It also has purchase history of the clients to calculate their **spending score** based on several criteria such as income, # of times the shop at the mall, and $ spent in a year. (closer score is to 100 = more client spends). You were hired to segment clients into different groups based on annual income and spending score.

We will only use age and salary to predict this
```{r}
## import data 
mall <- read.csv("../data/Mall_Customers.csv")

## Cut down dataset
x <- mall %>%
  select(Annual.Income..k..,Spending.Score..1.100.)

# find optimal method of clusters K with elbow method
set.seed(123)

# vector to hold wcss for different # of clusters K
wcss <- vector()
for (i in 1:10) {
  # compute wcss for each number of clusters from 1 to 10 on our data, x
  wcss[i] <- sum(kmeans(x, i)$withinss)
}
# plot BOTH lines + points of WCSS for each cluster choice
plot(1:10,wcss, type = "b", main = "Clusters of clients", xlab = "# of clusters", ylab = "WCSS")
```

Optimal # of clusters = 5
```{r}
#Fit kmeans to mall dataset w/ optimal # of clusters for 10 random sets and 300 iterations of algorithm
kmeans_mall <- kmeans(x, 5, iter.max = 300, nstart = 10)
```

# Vizualize Clusters
```{r, warning=F,message=F}
# plot clusters w/ no distance lines
# shade + color ellipses w/ respect to their density
# label points in ellipses + keep same characters for all clusters
# each cluster is represented by the ellipse with smallest area containing all its points (span = T)
clusplot(x = x, clus = kmeans_mall$cluster, lines = 0, shade = T, color = T, labels = 2, plotchar = F,
         span = T, main = "Clusters of Clients", xlab = "Annual Income", ylab = "Spending Score")
```

Cluster 4 has high income, low score (careful/frugal), cluster 2 has high income + higher score (big rollers/main potential **target** of marketing campaigns = find out what they buy ), cluster 3 = average income + spending (typical/standard), cluster 1 = low income + low score (sensible), cluster 5 = low income, high score (careless)