---
title: "demystifyingR2"
author: "Steve Newns"
date: "September 3, 2017"
output: html_document
---
Demystifying R Part 2
========================================================

1. Run the following command to see what it does.
```{r}
summary(mtcars)
```

2. In 2013, the average mpg for a car was 23 mpg. The car models in mtcars come from 1973-1974.
Subset the data so that you create a new data frame that contains cars get 23+ mpg
```{r}
efficient <- subset(mtcars, mpg > 23)
summary(efficient)
```

3. How many cars get more than 23 mpg? 
```{r}
dim(efficient)
```

4. We can also use logical operators to find out which car(s) get > 30 mpg and have more than 100 raw horsepower.
```{r}
subset(mtcars, mpg > 30 & hp > 100)
```

There's only one car that gets more than 30 mpg and 100 hp.

5. What do you think this code does? --> subsets for cars that have < 14 mpg OR > 390 displacement
```{r}
subset(mtcars, mpg < 14 | disp > 390)
```

6. Print cars that have a 1/4 mile time (qsec) <= 16.90 seconds to the console.
```{r}
subset(mtcars, qsec <= 16.9)
```

7. Save subset of cars that weigh under 2k lbs (measured in lb/1000) to a variable called lightCars. 
Print the numbers of cars and the subset to the console.
```{r}
lightCars <- subset(mtcars, wt < 2)
nrow(lightCars)
```

8. You can also create new variables in a data frame. Create a variable mtcars$year + assume all of models were from 1974.
Run the code below.
```{r}
library(tidyverse)
mtcars <- mtcars %>%
  mutate(year = 1974)
head(mtcars)
```

Notice the # of variables changed in the work space.

To drop a variable, subset the data frame and select the variable you
want to drop with a negative sign in front of it.
```{r}
mtcars <- subset(mtcars, select = -year)
head(mtcars)
```

Or use tidyverse
```{r}
mtcars <- mtcars %>%
  mutate(year = 1974)
head(mtcars)
```

```{r}
mtcars <- mtcars %>%
  mutate(year = NULL)
head(mtcars)
```

Notice, we are back to 11 variables in the data frame.

9. What do you think this code does? --> adds 2 years to each record, duplicating the record so it has both year values
```{r}
mtcars$year <- c(1973, 1974)
```

Open the table of values to see what values year takes on.
```{r}
head(mtcars)
```
Drop the year variable from the data set.
```{r}
mtcars <- mtcars %>%
  mutate(year = NULL)
head(mtcars)
```


10. Preview of ifelse()
```{r}
#all weights
mtcars$wt

# logical vector of T/F for if a car is> 3000 pounds
cond <- mtcars$wt < 3
cond

# for each elemen in new weight class variable, if condition = True, make it light, if not make it average
mtcars$weight_class <- ifelse(cond, 'light', 'average')
mtcars$weight_class

# make condition a higher weight
cond <- mtcars$wt > 3.5
# if an average car > 3500 lbs, make it heavy, if not leave it
mtcars$weight_class <- ifelse(cond, 'heavy', mtcars$weight_class)
mtcars$weight_class
```

Be careful that you don't bring in too much data into R at once since R holds all data in working memory.

```{r}
rm(cond)
rm(efficient)
```


Which car(s) have an mpg >= 30 OR horsepower < 60
```{r}
data("mtcars")
subset(mtcars, mpg >= 30 | hp < 60)
```