---
title: "lab3"
author: "Steve Newns"
date: "May 21, 2018"
output: html_document
---
##Logistic regression and NFL kickers

In this lab, we’ll gain experience implementing logistic regression to estimate the probability of successful NFL field goals given play and game specific conditions. Logistic regression will be a tool we use throughout the semester, and it is useful for a variety of sports/statistics questions. While field goal kickers are not the most exciting players in the game, their analysis provides an insight into our topic on expected points and decision-making, which we’ll cover next week.

First, there’s no known R-package with NFL data, so we’ll have to use one on the internet. I uploaded a .csv file with lots of kicker data. You can view that data by clicking [here](https://raw.githubusercontent.com/statsbylopez/StatsSports/master/Kickers.csv)

In R, there are a few ways to get data from the Internet, such as the `RCurl` package

```{r}
library(RCurl)
library(mosaic)
library(tidyverse)
library(ggplot2)

# load data in 
url <- getURL("https://raw.githubusercontent.com/statsbylopez/StatsSports/master/Kickers.csv")
nfl_kick <- read.csv(text = url)
head(nfl_kick)
```

### Exploratory data analysis

Let’s start with some basic data analysis, which should be the first thing we think about when looking at a new data set.

```{r}
bwplot(Distance ~ as.factor(Year), nfl_kick)
tally(Success ~ Grass, data = nfl_kick, format = "proportion")
tally(Success ~ GameMinute >= 57, data = nfl_kick, format = "proportion")
```
**1. Describe the distributions of field goal distance by year. Do there appear to be any changes over time? Note In the R-command, we turn `Year` into a factor for the boxplot. `Year` can be treated as either categorical or continuous.**

* There appears to be no real different in the overall IQR (spread) nor median (center) of FG distance by year, with Q1 being basically the same for all years, and Q3 fluctuating but not too widly
* There is a but more variance in 2011 (taller box), and 2008 is the only year with some long outliers
```{r}
nfl_kick %>% filter(Year==2008) %>% arrange(desc(Distance)) %>% head()
```

**2. Describe the link between success rates (`Success`) and `Grass` and the `GameMinute` variables.**

* Kickers appear to make more kicks if the field is indeed grass and that the kick does not take place within the final 3 minutes.

**3. Propose one alternative reason for field goals in the game’s final minutes showing lower success rates.**

* This could be due to pressure, as the game is coming to and end and the kick could be for the tie/the win.

**4. Use exploratory data analysis tools to** 

* **(i) identify if kickers on grass surfaces attempt shorter kicks than those on non-grass surfaces and**
* **(ii) identify if there is any link between the distance of the kick and the game minute of the kick.**

* 

### Logistic regression

At this point, we are aware that, overall, there seems to be a link between surface, distance, game minute, and field goal success. However, there’s also so correlation between a few of these predictors; kickers take longer field goals in the final few minutes of each half, for example. This makes it difficult to discern if the link between GameMinute and lower Success rates is real, or if its just simply because the kickers were kicking longer field goals.

We are going to walk through several logistic regression models of field goal kickers, with a focus on 

* (i) identifying the most important parts of the output
* (ii) interpreting the results, and, eventually 
* (iii) linking back to the probability of a successful kick.

First, let’s fit the model that we used in class on Tuesday $log \frac {P(Success = 1)} {1 - P(Success=1)} = \beta_0 + \beta_1Distance$
```{r}
fit_1 <- glm(Success ~ Distance, nfl_kick, family = "binomial")
msummary(fit.1)
```
We can write the estimated regression equation as:
log(
P (Success=1)
1−P (Success=1) ) = 5.7246 + (−0.1026) ∗ Distance
Estimated probabilities
Recall, probability estimates can be made by transforming the above equation (written on the log-odds scale)
into probabilities.
P(Success = 1) = e
5.7246+(−0.1026)∗Distance
1+e
5.7246+(−0.1026)∗Distance
R will allow us to make such calculations quickly, and there are several ways to do this.
First, we can use the exp() function to compute the exponential function. Let’s say we have a 50-yard field
goal:
phat.50 <- exp(5.7246 + (-0.1026)*50) / (1 + exp(5.7246 + (-0.1026)*50))
phat.50
Our model suggests that the probability of a successful 50-yard field goal is 64.4%.
We can also use the predict() function in R to get the predicted probabilities for an assortment of distances.
In the code below, the type = "response" ensures that the predictions are done on the probability scale (as
opposed to, say, the log-odds scale).
df.predict <- data.frame(Distance = c(20, 30, 40, 50, 60))
pred <- predict(fit.1, df.predict, type = "response")
pred
Slope estimates from logistic regression
Recall that the estimated slopes from a logistic regression model are given on the log-odds scale. We can
exponentiate to get estimated odds ratio.
2
exp(-0.1026)
exp(fit.1$coeff)
exp(confint(fit.1))
The first command calculated the estimated odds ratio by hand, while the second used the fact that R stores
regression coefficients as coeff. In the third command, we use the confint command to get estimated
confidence intervals for each parameter.
In the logistic regression output, we also see an AIC value (more on that later), as well as a z-test statistic
for each parameter. In this case, the estimated β1 is significantly different from 0. There is strong evidence
that distance is linked to field goal success (shocker!).
Multiple logistic regression
Ultimately, we are interested in how several of our variables link to field goal success. Let’s create one model,
fit.2, below.
options(scipen=99)
fit.2 <- glm(Success ~ Distance + Grass + Year + GameMinute,
data = nfl.kick, family = "binomial")
msummary(fit.2)
There’s a bunch to get to in this model. Let’s start by getting our estimated odds ratios:
odds.ratios <- exp(fit.2$coeff)
round(odds.ratios, 3)
odds.intervals <- confint(fit.2)
round(odds.intervals, 3)
First, notice that the estimated effecet of Distance did not change much from fit.1. This suggests that,
when including the other predictors, the link between Distance and Success remains the same.
Next, Grass is a categorical variable. We interpret this as follows:
Given ‘Year‘, ‘GameMinute‘, and ‘Distance‘, the odds of a successsful kick on grass surfaces are about 0.856
times that of one kicked on non-grass surfaces.
Given ‘Year‘, ‘GameMinute‘, and ‘Distance‘, the odds of a successful kick on grass surfaces are about 14.4%
less than that of one kicked on non-grass surfaces.
Either interpretation above is fine; I tend to think of the second one (relative percent odds) as more intuitive.
In this model, Year and GameMinute are treated as continuous, and their interpretations would be similar to
those for Distance.
3
On your own
5. Interpret the coefficient on Year in fit.2 (using the odds-ratio scale).
6. Using fit.2, estimate the probability of a successful 40-yard kick in 2010, made on a grass surface in
the 10th minute of the game.
7. Do the same as in (5), only now using a non-grass surface.
8. Turn your probility estimates from (5) and (6) into odds, and then take their ratio. Where else do you
see this number?
9. In the summary of fit.2, the coefficient for Grass is larger than the coefficient for Distance. Does
this mean that surface is a more important predictor of field goal success than distance? If not, give a
guess as to why the results are the way they are.
10. There are several variables in this data set. Using the AIC criteria, identify the model that is the best
fit for our Success outcome.
11. One variable that is difficult to include in a regression model is the kicker himself. Using the idea of
Expected points, think of ways that we can estimate the net benefit or harm of each kicker, at least
based on their field goal attempts.
12. Given your readings, are there any other variables that you would want to account for when measuring
field goal success that aren’t in the current data set?
4