---
title: "Inference for categorical data"
author: "Steve Newns"
date: "October 25, 2017"
output: html_document
---

In August of 2012, news outlets ranging from the [Washington Post](http://www.washingtonpost.com/national/on-faith/poll-shows-atheism-on-the-rise-in-the-us/2012/08/13/90020fd6-e57d-11e1-9739-eef99c5fb285_story.html) to the [Huffington Post](http://www.huffingtonpost.com/2012/08/14/atheism-rise-religiosity-decline-in-america_n_1777031.html) ran a story about the rise of atheism in America. The source for the story was a poll that asked people, "Irrespective of whether you attend a place of worship or not, would you say you are a religious person, not a religious person or a convinced atheist?" This type of question, which asks people to classify themselves in 1 way or another, is common in polling + generates categorical data.


```{r load-packages, message=FALSE}
library(statsr) # Duke package
library(tidyverse)
library(ggplot2)
```

### The survey

The press release for the poll, conducted by WIN-Gallup International, can be accessed [here](http://www.wingia.com/web/files/richeditor/filemanager/Global_INDEX_of_Religiosity_and_Atheism_PR__6.pdf).

1. How many people were interviewed for this survey? 
<ol>
<li> A poll conducted by WIN-Gallup International surveyed 51,000 people from 57 countries. </li>
<li> A poll conducted by WIN-Gallup International surveyed 52,000 people from 57 countries. </li>
<li> A poll conducted by WIN-Gallup International surveyed 51,917 people from 57 countries. </li>
<li> **A poll conducted by WIN-Gallup International surveyed 51,927 people from 57 countries.** </li>
</ol>

2. Which of the following methods were used to gather information? 
<ol>
<li> Face to face </li> 
<li> Telephone </li>
<li> Internet </li> 
<li> **All of the above** </li>
</ol>

3. T/F: In the 1st paragraph, several key findings are reported. These percentages appear to be **sample statistics**. 
<ol>
<li> **True** </li> 
<li> False </li>
</ol>

4. T/F: The title of the report = "Global Index of Religiosity and Atheism". To generalize the report's findings to the global human population, we must assume the sample was a random sample from the entire population in order to be able to generalize the results to the global human population. This seems to be a reasonable assumption.
<ol>
<li> True </li> 
<li> **False** </li>
</ol>

### The data

Turn your attention to Table 6 (pages 15 and 16), which reports sample size + response %'s for all 57 countries. While this is a useful format to summarize data, we will base our analysis on the original data set of individual responses to the survey.

```{r load-data}
data(atheism)
```

5. What does each row of Table 6 correspond to? 
<ol>
<li> **Countries** </li> 
<li> Individual Persons </li> 
<li> Religions </li>
</ol>

```{r}
head(atheism)
```

6. What does each row of `atheism` correspond to?
<ol>
<li> Countries </li> 
<li> **Individual Persons** </li> 
<li> Religions </li>
</ol>

To investigate the link between these 2 ways of organizing this data, take a look at the estimated proportion of atheists in the US. Towards the bottom of Table 6, we see that this is 5%. We should be able to come to the same number using the `atheism` data.

```{r us-atheism}
us12 <- atheism %>%
  filter(nationality == "United States" , atheism$year == "2012")# %>%
  #group_by(response) %>%
  #summarise(prop = mean(response))
prop.table(table(us12$response))
```


7. What is the proportion of atheist responses in the US in 2012? Does this % agree w/ the % in Table 6?
<ol>
<li> **True** </li> 
<li> False </li>
</ol>

## Inference on proportions

As hinted earlier, Table 6 provides **sample statistics** (calculations made from the sample of 51,927 people). We'd like insight into the **population parameters**. You answer "What proportion of people in your sample reported being atheists?" with a *statistic*; while "What proportion of people on earth would report being atheists" is answered with an *estimate of the parameter*.

Inferential tools for estimating population proportion are analogous to those used for means: the CI and the hypothesis test.

<div id="exercise">
**Exercise**: Write out the conditions for inference to construct a 95% confidence interval for the proportion of atheists in the United States in 2012. Are you confident all conditions are met?
</div>

<ul>
<li> **Independence**: Assume observations are independent via random sampling or assignment, and if sampling w/out replacement, < 10% of the population must be sampled </li>
<li> **Sample Size/Skew**: Check success-failure condition </li>
</ul>

If the conditions for inference are reasonable, we can either calculate the SE + construct the interval by hand, or allow `inference` to do it for us.

```{r us-atheism-ci}
inference(y = response, data = us12, statistic = "proportion", type = "ci", method = "theoretical", success = "atheist")
```

Here, a **success** = a response of `atheist`.

Although formal CI's and hypothesis tests don't show up in the report, suggestions of inference appear at the bottom of page 7: "In general, the error margin for surveys of this kind is $\pm$ 3-5% at 95% confidence."

<div id="exercise">
**Exercise**: Imagine that, after reading a front page story about the latest public opinion poll, a family member asks you, "What is a margin of error?" In 1 sentence, and ignoring the mechanics behind the calculation, how would you respond in a way that conveys the general concept?
</div>

**A margin of error is a range of values that the *actual* value may vary by, related to the reported statistic**

```{r}
.0499 - .0364
```

8. Based on the R output, what is the margin of error for the estimate of the proportion of the proportion of atheists in US in 2012? 
<ol>
<li> The margin of error for the estimate of the proportion of atheists in the US in 2012 is 0.05. </li> 
<li> The margin of error for the estimate of the proportion of atheists in the US in 2012 is 0.025. </li> 
<li> **The margin of error for the estimate of the proportion of atheists in the US in 2012 is 0.0135.** </li>
</ol>

<div id="exercise">
**Exercise**: Using `inference`, calculate CI's for proportion of atheists in 2012 in 2 other countries of your choice + report the associated margins of error. Be sure to note whether the conditions for inference are met.
</div>
```{r me-perc-atheist-other-countries}
ire12 <- atheism %>%
  filter(nationality == "Ireland" , atheism$year == "2012")
prop.table(table(ire12$response))

inference(y = response, data = ire12, statistic = "proportion", type = "ci", method = "theoretical", success = "atheist")

pol12 <- atheism %>%
  filter(nationality == "Poland" , atheism$year == "2012")
prop.table(table(pol12$response))

inference(y = response, data = pol12, statistic = "proportion", type = "ci", method = "theoretical", success = "atheist")
```

```{r}
(ire.moe <- .0990099 - .0806)
(pol.moe <- .04952381 - 0.031)
```

**The margin of error for Ireland is 1.84% and the margin of error for Poland is 1.85%**

## How does the proportion affect the margin of error?

Imagine you've set out to survey 1000 people on 2 questions: are you female? and are you left-handed? Since both of these sample proportions were calculated from the same sample size, they should have the same margin of error, right? Wrong! While the margin of error *does* change w/ *sample size*, it is *also* affected by the *proportion*.

Think back to the formula for SE: $SE = \sqrt{p(1-p)/n}$, which is then used in the formula for margin of error for a 95% CI: $ME = 1.96\times SE = 1.96\times\sqrt{p(1-p)/n}$. Since the population proportion $p$ is in this $ME$ formula, it should make sense that the margin of error is in some way dependent on the population proportion. We can visualize this relationship by creating a plot of $ME$ vs. $p$.

1st, make a vector `p` = a sequence from $0$ to $1$ w/ each number separated by $0.01$, then create a vector of the margin of error (`me`) associated w/ each of these values of `p` using the approximate formula ($ME = 1.96 \times SE$), + lastly, plot the 2 vectors against each other to reveal the relationship

```{r me-plot}
d <- data.frame(p <- seq(0, 1, 0.01))
n <- 1000

d <- d %>%
  mutate(me = 1.96*sqrt(p*(1 - p)/n))
d %>% 
  ggplot(aes(p, me)) +
    geom_line()
```

9. Which of the following is false about the relationship between $p$ and $ME$. 
<ol>
<li> The $ME$ reaches a minimum at $p = 0$. </li> 
<li> The $ME$ reaches a minimum at $p = 1$. </li> 
<li> The $ME$ is maximized when $p = 0.5$. </li> 
<li> **The most conservative estimate when calculating a confidence interval occurs when $p$ is set to 1**. </li>
</ol>

The question of atheism was asked by WIN-Gallup International in a similar survey conducted in 2005. We assume here that sample sizes have remained the same. Table 4 on page 13 of the report summarizes survey results from 2005-2012 for 39 countries.

Answer the following 2 questions using `inference` + write out the hypotheses for any tests you conduct + outline the status of the conditions for inference.

10. T/F: There is convincing evidence Spain has seen a change in its atheism index between 2005 and 2012. <br><br> *Hint:* Create a new data set for respondents from Spain. Then use their responses as the first input on the `inference`, and use `year` as the grouping variable.

<ul>
<li> H0: p05 = p12 </li>
<li> H0: p05 != p12 </li>
</ul>
```{r}
esp <- atheism %>%
  filter(nationality == "Spain")

inference(esp$response, factor(esp$year), est = "proportion", type = "ht", method = "theoretical",
          success = "atheist", null = 0, alternative = "twosided")
```

<ol>
<li> True </li> 
<li> **False** </li>
</ol>

11. T/F: There is convincing evidence the US has seen a change in its atheism index between 2005 and 2012.

```{r}
us <- atheism %>%
  filter(nationality == "United States")

inference(us$response, factor(us$year), est = "proportion", type = "ht", method = "theoretical",
          success = "atheist", null = 0, alternative = "twosided")
```

<ol>
<li> **True** </li> 
<li> False </li>
</ol>

12. If in fact there has been no change in the atheism index in the countries listed in Table 4, in how many of those countries would you expect to detect a change (at a significance level of 0.05) simply by chance? <br><br> *Hint:* Type 1 error.

```{r}
# sig level * # of countries
39 * .05
```

<ol>
<li> 0 </li>
<li> 1 </li> 
<li> **1.95** </li> 
<li> 5 </li>
</ol>

13. Suppose you're hired by the local government to estimate the proportion of residents that attend a religious service on a weekly basis. According to the guidelines, the estimate must have a margin of error no greater than 1% with 95% confidence. You have no idea what to expect for $p$. How many people would you have to sample to ensure that you are within the guidelines? <br><br> *Hint:* Refer to your plot of the relationship between $p$ and margin of error. Do not use the data set to answer this question.
<ol>
<li> 2401 people </li>
<li> At least 2401 people </li>
<li> 9604 people </li> 
<li> **At least 9604 people** </li>
</ol>

```{r}
tail(d)
```

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.
</div>