---
title: "DataVizInR"
author: "Steve Newns"
date: "November 29, 2017"
output: html_document
---

```{r}
library(MASS)
library(car)
library(wordcloud)
library(ggplot2)
```
`matplot()` can be used to easily generate a plot w/ several scatterplots on the same set of axes. By default, the points in these scatterplots are represented by the numbers 1 through n, where n = the total number of scatterplots included, but most of options available w/ `plot()` are also possible by specifying the appropriate arguments.

Set up a plot array w/ 4 of these multiple scatterplot displays, each including 1 more scatterplot than the previous one + judge for yourself how many scatterplots is too many.

    Use the par() function to set up the mfrow parameter to generate a two-by-two plot array.
    Use the matplot() function to construct a two-scatterplot display of
    protein and fat versus calories. Give this plot the title "Two scatterplots" using the title() function. Label the x-axis "calories" and the y-axis "" to get rid of the default y-axis label.
    protein, fat and fibre versus calories. Give this plot the title "Three scatterplots". Similarly, label the x-axis "calories" and the y-axis "".
    protein, fat, fibre and carbo versus calories. Give this plot the title "Four scatterplots". Again, label the x-axis "calories" and the y-axis "".
    protein, fat, fibre, carbo, and sugars versus calories. Give this plot the title "Five scatterplots". Be sure to label the x-axis "calories" and the y-axis "" again.


```{r}
# Construct the vector keep_vars +  extract the desired subset of UScereal
keep_vars <- c("calories", "protein", "fat",
               "fibre", "carbo", "sugars")
df <- UScereal[, keep_vars]

# Set up a two-by-two plot array
par(mfrow = c(2,2))

# Use matplot() to generate an array of 2, 3, 4, then 5 scatterplots
matplot(df$calories, df[,c("protein","fat")], xlab = "calories", ylab = "")
title("Two scatterplots")

matplot(df$calories, df[,c("protein","fat","fibre")], xlab = "calories", ylab = "")
title("Three scatterplots")

matplot(df$calories, df[,c("protein","fat","fibre","carbo")], xlab = "calories", ylab = "")
title("Four scatterplots")

matplot(df$calories, df[,c("protein","fat","fibre","carbo","sugars")], xlab = "calories", ylab = "")
title("Five scatterplots")
```

# How many words is too many?

Scatterplot arrays lose their utility if allowed to become too complex, either including too many plots, or attempting to include too many scatterplots on 1 set of axes. More generally, any data viz loses its utility if it becomes too complex.

**Wordclouds** presents words in varying sizes depending on frequency via `wordcloud()` called w/ a character vector of words + a 2nd numerical vector giving the # of times each word appears in the collection used to generate the wordcloud.

2 other useful arguments = `scale` = a 2-component numerical vector giving relative size of the largest + smallest words in the display + `min.freq` = only display words that occur at least this # of times in the collection w/ a default value = 3.

The wordcloud package is already loaded in your workspace.

    Use the table() function to create the variable mfr_table that tabulates the number of times each level of the Manufacturer variable appears in the Cars93 data frame.
    Use the wordcloud() function to display this manufacturer data, setting the scale argument to c(2, 0.25) to make the wordcloud fit in the display window.
    By default, the wordcloud() function shows only those words that appear 3 or more times. Use the min.freq argument to obtain a display with all Manufacturer values in the wordcloud.
    Use the table() function to create the variable model_table that tabulates the number of times each level of the Model variable appears in the Cars93 data frame.
    Use the wordcloud() function with the scale argument set to c(0.75, 0.25) and the min.freq argument set to display all Model values. Use the zoom feature in the 'Plots' window to see this wordcloud more clearly. Does it convey useful information?

```{r}
# Create table of manufacturer frequencies
mfr_table <- table(Cars93$Manufacturer)

# create word cloud to show all manufacturers
wordcloud(words = names(mfr_table), 
          freq = as.numeric(mfr_table), 
          scale = c(2,.25), 
          min.freq = 1)

# Create model_table of model frequencies
model_table <- table(Cars93$Model)

# Create the wordcloud of all model names with smaller scaling
wordcloud(words = names(model_table), 
          freq = as.numeric(model_table), 
          scale = c(0.75, 0.25), 
          min.freq = 1)
```

# The Anscombe quartet

**Anscombe quartet** = collection of 4 datasets that appear to be essentially identical on the basis of simple summary statistics (means + SDs) but the behavior of the 4 datasets is quite different which becomes immediately apparent when we plot them.

    Use the par() function to set up a two-by-two plot array.
    Using plot(), create 4 separate plots from the anscombe data frame:
        y1 vs. x1
        y2 vs. x2
        y3 vs. x3
        y4 vs. x4
```{r}
# Set up a two-by-two plot array
par(mfrow = c(2,2))

# Plot alls y's vs x's
plot(anscombe$x1, anscombe$y1)
plot(anscombe$x2, anscombe$y2)
plot(anscombe$x3, anscombe$y3)
plot(anscombe$x4, anscombe$y4)
```

The point of this exercise is to illustrate how much more clearly we can see differences in datasets if we plot all of them w/ the *same* x and y ranges + also illustrates the utility of improving the x- and y-axis labels and of adding informative plot titles.


    Examine the range of x and y values from the previous four plots to determine a common range that covers the ranges of all four of these plots. Use integer values to keep things simple.
    Set up a two-by-two plot array using the par() function.
    Plot the y1 variable against the x1 variable using these common ranges, with x-axis label "x value" and y-axis label "y value", and use the main argument to add the title "First dataset".
    Repeat for the other three Anscombe data pairs, adding the appropriate titles.

```{r}
# Define common x and y limits for the four plots
xmin <- 4
xmax <- 19
ymin <- 3
ymax <- 13

# Set up a two-by-two plot array
par(mfrow = c(2,2))

# Plot y1 vs. x1 with common x and y limits, labels & title
plot(anscombe$x1, anscombe$y1,
     xlim = c(xmin, xmax),
     ylim = c(ymin, ymax),
     xlab = "x value", ylab = "y value",
     main = "First dataset")

# Do the same for the y2 vs. x2 plot
plot(anscombe$x2, anscombe$y2,
     xlim = c(xmin, xmax),
     ylim = c(ymin, ymax),
     xlab = "x value", ylab = "y value",
     main = "Second dataset")

# Do the same for the y3 vs. x3 plot
plot(anscombe$x3, anscombe$y3,
     xlim = c(xmin, xmax),
     ylim = c(ymin, ymax),
     xlab = "x value", ylab = "y value",
     main = "Third dataset")

# Do the same for the y4 vs. x4 plot
plot(anscombe$x4, anscombe$y4,
     xlim = c(xmin, xmax),
     ylim = c(ymin, ymax),
     xlab = "x value", ylab = "y value",
     main = "Fourth dataset")
```

# Using multiple plots to give multiple views of a dataset

Another useful application of multiple plot arrays besides comparison is presenting multiple related views of the same dataset.

Give 4 views of the same dataset: raw data values, a histogram of these data values, a density plot, and a normal QQ-plot.

Note that the MASS and car packages have been pre-loaded, making the geyser data and the truehist() and qqPlot() fuctions available for your use.
Use the par() function to set the mfrow parameter for a two-by-two plot array.
In the upper left, use the plot() function to show the values of the duration variable from the geyser dataset, using the main argument to specify the plot title as "Raw data".
In the upper right, use the truehist() function to generate a histogram of the duration data, using the main argument to give this plot the title "Histogram".
In the lower left, use the plot() and density() functions to display the 

```{r}
# Set up a two-by-two plot array
par(mfrow = c(2,2))

# Plot the raw duration data
plot(geyser$duration, main = "Raw data")

# Plot the normalized histogram of the duration data
truehist(geyser$duration, main = "Histogram")

# Plot the density of the duration data
plot(density(geyser$duration), main = "Density")

# Construct the normal QQ-plot of the duration data
qqPlot(geyser$duration, main = "QQ-plot")
```

# Constructing and displaying layout matrices

Can set up a **layout matrix** to be used w/ `layout()` in creating a plot array. Think of the layout matrix as the plot pane, where 0 = empty space + other #'s represent the plot #, which is determined by the sequence of visualization function calls. For example, a 1 in the layout matrix = the viz that was called 1st, a 2 = plot of 2nd viz call, etc. Create your own 3 x 2 layout matrix w/ `c()` to concatenate numbers into vectors that will form the rows of the matrix.

Then use `matrix()` to convert these rows into a matrix + apply `layout()` to set up the desired plot array. The **convenience function** `layout.show()` can then be used to verify the plot array has the shape you want.


    Using the matrix() function, create a matrix called layoutMatrix with three rows and two columns:
        the first row designates an empty plot to the left and plot 1 to the right.
        the second row designates plot 2 to the left and an empty plot to the right.
        the third row designates an empty plot to the left and plot 3 to the right.
    Use the layout() function to set up the desired plot array.
    Use the layout.show() function to show the arrangement of all three plots.

```{r}
# Use the matrix function to create a matrix with three rows and two columns
layoutMatrix <- matrix(
  c(0, 1,
    2, 0,
    0, 3), byrow = T, nrow = 3)

# Call the layout() function to set up the plot array
layout(layoutMatrix)

# Show where the three plots will go 
layout.show(n=3)
```

`layout()` is a little harder than setting up plot arrays w/ `mfrow` parameter, but it is a lot more flexible.

# Creating a triangular array of plots

Use layoutMatrix to give 3 different views of `whiteside` w/ 1st plot on upper right of plot array = the relationship of `Gas` + `Temp` using data where `Insul = "Before"`, 2nd plot in center left of plot array = relationship of same 2 variables using all data, + 3rd plot on lower left of plot array = relationship using data where `Insul = "After"`.

The primary motivation for this exercise is that it is not possible to construct a plot array in this format using the mfrow parameter, since the array is not rectangular.

The layout matrix you set up in the previous is available in your workspace as layoutMatrix. The whiteside data frame is already available in your workspace as well.

    Call the layout() function on layoutMatrix to set up the plot array.
    Construct a vector indexB that points to only those records with the Insul value "Before" and a vector indexA that points to only those records with the Insul value "After".
    Plot the Gas versus Temp values from the indexB data in the upper right plot in your array, using the y-axis limits c(0, 8). Give this plot the title "Before data only".
    Plot the Gas versus Temp values from the complete dataset in the center left plot, using the same y-axis limits as the first plot. Give this plot the title "Complete dataset".
    Plot the Gas versus Temp values from the indexA data in the lower right plot, again using the same y-axis limits. Give this plot the title "After data only".

```{r}
# Set up the plot array via layout(layoutMatrix) call to set up the plot array.
layout(layoutMatrix)

# Construct vectors indexB and indexA
indexB <- which(whiteside$Insul == "Before")
indexA <- which(whiteside$Insul == "After")

# Create plot 1 and add title
plot(whiteside$Temp[indexB], whiteside$Gas[indexB],
     ylim = c(0,8))
title("Before data only")

# Create plot 2 and add title
plot(whiteside$Temp, whiteside$Gas,
     ylim = c(0,8))
title("Complete dataset")

# Create plot 3 and add title
plot(whiteside$Temp[indexA], whiteside$Gas[indexA],
     ylim = c(0,8))
title("After data only")
```

# Creating arrays with different sized plots

Besides creating non-rectangular arrays, `layout()` can be used to create plot arrays w/ different sized component plots, something else not possible via `par()`'s `mfrow` parameter.

Create standard scatterplot of `zn` vs. `rad` from `Boston` as a smaller plot in the upper left, w/ a larger sunflower plot of the same data in the lower right.

The Boston data frame in the MASS package is already available in your workspace.

    Using the c() function:
    Create the three-element vector row1 with values (1, 0, 0).
    Create the three-element vector row2 with values (0, 2, 2).
    Combine the first vector with two copies of the second vector into layoutVector, a vector of length 9.
    Using the matrix() function, convert layoutVector into the 3-by-3 matrix layoutMatrix whose first row is row1 and whose second and third rows are row2.
    Use the layout() function with layoutMatrix to configure a two-element plot array.
    In the first (smaller) plot, create a standard scatterplot of the zn variable versus the rad variable from the Boston data frame.
    In the second (larger) plot, create a sunflower plot using the sunflowerplot() function and the same variables.

```{r}
# Create row1, row2, and layoutVector
row1 <- c(1,0,0)
row2 <- c(0,2,2)
layoutVector <- c(row1,row2,row2)

# Convert layoutVector into layoutMatrix
layoutMatrix <- matrix(layoutVector, byrow = T, nrow = 3)

# Set up the plot array
layout(layoutMatrix)

# Plot scatterplot
plot(Boston$rad, Boston$zn)

# Plot sunflower plot
sunflowerplot(Boston$rad, Boston$zn)
```

# Some plot functions also return useful information

calling `barplot()` function has the side effect of creating the plot we want but also a numerical vector w/ the center positions of each bar in the plot, + the vector is returned invisibly so we don't normally see it, but we can capture it w/ an assignment statement.

These return values can be especially useful to overlay text on bars of a horizontal barplot. Then, we capture the return values + use them as the `y` parameter in a subsequent call to `text()` to place the text at whatever `x` position we want but overlaid in the middle of each horizontal bar. 

    Use the table() function to generate the object tbl summarizing the number of records listing each value of the Cylinders variable from the Cars93 data frame.
    Use the barplot() function with the horiz argument set to TRUE to construct a horizontal barplot of this record summary, specifying the color of the bars to be "transparent" and returning the vector mids giving the vertical positions of the centers of each bar in the plot. Specify the names.arg argument as "" to suppress the y-axis legend on this plot.
    Use the text() function to label the Cylinders value for each bar in the barplot at the horizontal position 20. The names() function may prove useful here.
    Use the text() function to list the counts for each cylinders value at the horizontal position 35. Here, the as.numeric() function may prove useful.

```{r}
# Create a table of Cylinders frequencies
tbl <- table(Cars93$Cylinders)

# Generate a horizontal barplot of these frequencies with transparent color w/ suppressed y-labels
mids <- barplot(tbl, horiz = T, col = "transparent", names.arg = "")

# Add names labels at horiz. position 20 + count labels at horiz. position 35
text(20, mids, labels = names(tbl))
text(35, mids, labels = as.vector(tbl))
```


# Using symbols() to display relations between more than 2 variables

Scatterplots allow us to see how 1 numerical variable changes w/ values of a 2nd numerical variable. `symbols()` allows us to extend scatterplots to show the influence of other variables when called w/ variables x + y that define a scatterplot, along w/ another argument that specifies 1 of several possible shapes. 

Use the `circles` arg to create a bubbleplot where each DP is represented by a circle whose radius depends on the 3rd variable specified by the value of this argument.


    Use the symbols() function with its default settings and the appropriate arguments to create a bubble plot of MPG.city versus Horsepower from the Cars93 data frame, with the bubble area by the Price variable. Note that this means the bubble radius should be proportional to the square root of Price.
    Re-create the first plot, but with the optional argument inches set to 0.1.

```{r}
# Call symbols() to create the default bubbleplot
symbols(Cars93$Horsepower, Cars93$MPG.city, circles = sqrt(Cars93$Price))

# Repeat, with the inches argument specified
symbols(Cars93$Horsepower, Cars93$MPG.city, circles = sqrt(Cars93$Price), inches = .1)
```

Note how much cheaper the low-horsepower cars in the upper left are than the high-horsepower cars in the lower right. 

# Saving plot results as files

Generating a collection of different plots + often using results to help decide how to proceed w/ our analysis is common + is particularly true in early phases of an EDA, but once we have generated a plot to share w/ others, it is important to save it in an external file.

Can save graphical results in several different external file formats = jpeg, png, tiff, pdf. In addition, can incorporate graphical results into external documents, using tools like `Sweave()` or `knitr` package such as w/ R Markdown docs.

B/c PNG files can be easily shared + viewed as e-mail attachments + incorporated into many slide prep packages (e.g. Powerpoint), create a plot + save it as a PNG via `png()` w/ the name of the file to be generated by setting up a special environment that captures all graphical output until we exit this environment w/ `dev.off()`


    Use the png() function to direct all subsequent plot results to the external file bubbleplot.png.
    Run the code to re-create the second bubble plot from the previous exercise.
    Exit the png() environment to return graphics control to your session by calling dev.off().
    Use the list.files() code provided to verify that bubbleplot.png was created. To do this, you need to specify "png" as the value to the pattern argument to list.files().

```{r}
# Call png() w/ name of file to create to direct all subsequent plot results to the external file
#png("bubbleplot.png")

# Re-create plot from the last exercise
#symbols(Cars93$Horsepower, Cars93$MPG.city, circles = sqrt(Cars93$Price), inches = 0.1)

# Save + file + return to interactive session
#dev.off()

# Verify that we have created the file
list.files(pattern = "png")
```


# Iliinsky and Steele's 12 recommended colors

Create a horizontal barplot that shows Iliinsky + Steele's set of recommended 12 colors, in descending order of desirability w/ the 1st 6 "more preferred" colors displayed w/ longer bars to visually emphasize their preferred status over the other 6

    Notice how the barWidths vector of length 12 contains the length of each bar shown.
        The 6 longer bars have a value of 2.
        The 6 shorter ones have a value of 1.
    Recreate the horizontal barplot you see on the right, using the horiz, col, axes, names.arg, and las arguments to barplot(). Note that since the horizontal barplot is constructed from the bottom up, it is necessary to reverse the IScolors and barWidths vectors in the barplot() function call, using the rev() function.
```{r}
# Iliinsky and Steele color name vector
IScolors <- c("red", "green", "yellow", "blue","black", "white", "pink", "cyan", "gray", "orange", "brown", "purple")

# Make more preferred colors bars twice as long
barWidths <- c(rep(2, 6), rep(1, 6))

# Recreate the horizontal barplot with colored bars w/ no horizontal axis
barplot(rev(barWidths), horiz = T, col = rev(IScolors), axes = F, names.arg = rev(IScolors), las = 1)
```

# Using color to enhance a bubbleplot

`symbols()` could be used to generate a bubbleplot showing the relationship between 3 variables (like Horsepower, MPG.city, + Price from `Cars93`). There, the basic format was a scatterplot of MPG.city vs. Horsepower, w/ points represented as bubbles, sized by Price.

Create a variation on this plot, using the factor variable `Cylinders` to determine both size + color of the bubbles. `as.numeric(Cars93$Cylinders)` generates a sequence of numerical values from 1 to 6 that specify the 6 unique levels of the `Cylinders` factor.

In cases where color is an option, the clarity of this bubbleplot can be improved substantially through the use of color in addition to symbol size. Since `Cylinders` exhibits 6 unique values, 6 colors are required to make this enhancement, so the top 6 colors recommended by Iliinsky + Steele are used here.

    Create the vector cylinderLevels giving numerical labels to the unique values of the Cylinders variable.
    Using the symbols() function, recreate the bubbleplot shown on the right of MPG.city vs. Horsepower with the bubbles sized by the cylinderLevels variable.
        Specify a maximum circle radius of 0.2 inches.
        Use the bg parameter to color each bubble according to the numerical level of the Cylinders factor variable (i.e., the cylinderLevels values), using the first six recommended colors from Iliinsky and Steele (e.g. the first Cylinder level should be colored "red", the second level colored "green", etc.)

```{r}
# Iliinsky and Steele color name vector
IScolors <- c("red", "green", "yellow", "blue", "black", "white", "pink", "cyan", "gray", "orange", "brown", "purple")

# Create the `cylinderLevel` variable
cylinderLevel <- as.numeric(Cars93$Cylinders)

# Create the colored bubbleplot
symbols(Cars93$Horsepower, Cars93$MPG.city, circles = cylinderLevel, inches = .2, bg = IScolors[cylinderLevel])
```

See how the number of cylinders influences the relationship between horsepower and gas mileage

# Using color to enhance stacked barplots

The most common barplot consists = a collection of vertical bars, each representing some characteristic of 1 of a finite # of datasets or data subsets. We used a somewhat less common horizontal barplot, useful in part b/c it allows text to be displayed horizontally across the bars.

Another useful variant of the standard bar plot = the **stacked bar plot**, where each bar in the plot = partitioned into segments characterizing portions of the data characterized by the bar. Stacked bar plots can also be generated using barplot()` where each bar is specified by a *matrix* whose columns = the heights of the segments in each bar.

By default, `barplot()` generates stacked bar plots using different shades of gray for different segments of each bar in the plot. Color can be a more effective alternative.

    Create the table tbl giving the record counts for each Cylinders value at each Origin value (this table should have 6 rows and 2 columns).
    Using the barplot() function, create a stacked barplot that summarizes this information using shades of gray.
    Recreate this first plot, but using the first six Iliinsky and Steele colors for the six Cylinders levels.
```{r}
# Create a table of Cylinders by Origin
tbl <- table(Cars93$Cylinders,Cars93$Origin)

# Create the default stacked barplot
barplot(tbl)

# Enhance this plot with color
barplot(tbl, col = IScolors)
```

# The tabplot package and grid graphics

`tabplot` allows you to tap into the power of **grid graphics** without explicit knowledge of how the system works under the hood. The main function in `tabplot` = `tableplot()`, developed to visualize data distributions + relationships between variables in large datasets.

Specifically, `tableplot()` constructs a set of side-by-side horizontal barplots, 1 for each variable, + works best when viewing up to about 10 variables at a time, for datasets w/ arbitrarily many records. Apply this function to a dataset w/ just under 68k records + 11 variables.

`tableplot()` is called w/ a data frame +, if no optional arguments are specified, it selects the 1st data col as the reference variable (can be of any type, but the display is easiest to explain it's numeric).

    Load the insuranceData package.
    Use the data() function to load the dataCar data frame.
    Load the tabplot package the normal way, but surround your call to library() with the suppressPackageStartupMessages() function to avoid a bunch of unncessary output from printing to the console.
    Apply the tableplot() function to the dataCar data frame.

```{r}
# Load the insuranceData package + load the dataCar data frame
library(insuranceData)
data(dataCar)

# Load tabplot package
suppressPackageStartupMessages(library(tabplot))

# Generate the default tableplot() display
tableplot(dataCar)
```

Note how much color enhances the displays of categorical variables like `veh_body` and `area`. 

# A lattice graphics example

Another package built on grid graphics = `lattice`, which unlike `tabplot`, is a general-purpose graphics package that provides alternative implementations of many of the plotting functions available in `base` graphics. Specific examples = scatterplots w/ `xyplot()`, bar charts w/ `barchart()`, + boxplots w/ `bwplot()`.

1 important difference between lattice graphics + base graphics = similar functions available in both graphics systems often produce very different results when applied to the same data. `bwplot()` creates horizontal boxplots, while the default result of `boxplot()` = a vertical boxplot display.

Another more important difference between lattice + base graphics = lattice graphics supports conditional plots that show the separate relationships between variables w/in different groups = construct a plot showing the relationship between `calories` + `sugars` from `UScereal`, conditional on the value of `shelf`


    Load the lattice package to make the function xyplot() available.
    Using the conditional formula format "y ~ x | z" (scatterplot of y vs. x, conditioned on z), construct a conditional scatterplot of calories vs. sugars conditional on shelf from the UScereal data frame in the MASS package. Make sure to convert shelf to a factor.
```{r}
# Load the lattice package
library(lattice)

# Use xyplot() to construct the conditional scatterplot
xyplot(calories ~ sugars| as.factor(shelf), data = UScereal)
```

# A ggplot2 graphics example

Like `lattice`, `ggplot2` is also based on grid graphics + also represents a general purpose graphics package.

`ggplot2` = based on grammar of graphics, a systematic approach to building + modifying graphical displays, starting w/ a basic graphics element + refining it through the addition of successive modifiers.

1st create a simple scatterplot of `MPG.city` vs. `Horsepower` from `Cars93` data frame then add a simple modifier that adds color based on the value of `Cylinders` + convert this result into a colored bubble plot, w/ both bubble sizes + colors determined by `Cylinders`.

The character vector IScolors from the previous exercises is still available in your workspace.

    Load the ggplot2 package.
    Create basePlot as the ggplot object based on the Cars93 data frame from the MASS package and the scatterplot aesthetic with x as the Horsepower variable and y as the MPG.city variable. When passing the variable names to x and y, they should be unquoted (e.g. x = var_name). (Note that this definition does not render the scatterplot.)
    Make a simple rendering of basePlot with the geom_point() function.
    Make a second rendering using the colour parameter of geom_point() to specify the first six Iliinsky-Steele colors for the six Cylinders levels.
    Make a third rendering with the same coloring as the second one, but with the point sizes corresponding to the levels of the Cylinders factor variable.
```{r}
# Create the basic plot (not displayed): basePlot
basePlot <- ggplot(Cars93, aes(x = Horsepower, y = MPG.city))

# Display the basic scatterplot
basePlot + 
  geom_point()

# Color the points by Cylinders value
basePlot + 
  geom_point(col = IScolors[Cars93$Cylinders])

# Make the point sizes also vary with Cylinders value
basePlot + 
  geom_point(col = IScolors[Cars93$Cylinders], 
             size = as.numeric(Cars93$Cylinders))
```

