---
title: "ImportingDataInR_Pt2"
author: "Steve Newns"
date: "December 5, 2017"
output: html_document
---

```{r}
library(ggplot2)
```
# Establish a connection

1st step to import data from a SQL db = creating a connection to it via different packages depending on db 

`dbConnect()` creates a connection between an R session + a SQL db  w/ 1st arg = a **DBIdriver object** that specifies how connections are made + how data is mapped between R + the db. For MySQL db's, you can build such a driver w/ ``RMySQL::MySQL()``. If the MySQL db is a remote db hosted on a server, you'll also have to specify the following arguments: `dbname`, `host`, `port`, `user` + `password.` Most of these details have already been provided.


    Load the DBI library, which is already installed on DataCamp's servers.
    Edit the dbConnect() call to connect to the MySQL database. Change the port argument (3306) and user argument ("student").
```{r}
# Load the DBI package
library(DBI)

# Edit dbConnect() call to connect to tweater db
con <- dbConnect(RMySQL::MySQL(), 
                 dbname = "tweater", 
                 host = "courses.csrrinzqubik.us-east-1.rds.amazonaws.com", 
                 port = 3306,
                 user = "student",
                 password = "datacamp")
class(con)
```

# List the database tables

Next step = see what tables db contains w/ `dbListTables()` which requires the connection object as an input + outputs a character vector w/ table names.
```{r}
tables <- dbListTables(con)
str(tables)
```

# Import users

DB contains data on a more tasty version of Twitter, **Tweater** where users can post tweats w/ short recipes for delicious snacks + comment on other tweats. There are 3 tables that have relations among

Importing the data on the users into the R session w/ `dbReadTable()` by passing the connection objec + the table to import. The resulting object is a standard R data frame.

```{r}
users <- dbReadTable(con,"users")
users
```

# Import all tables

Separate `dbReadTable()` calls for each + every table in a db = a lot of code duplication. Use `lapply()` to load in all tables in the database.

```{r}
(table_names <- dbListTables(con))
# apply read table to each table name + specify connection
tables <- lapply(table_names,dbReadTable, conn = con)
tables
```

Now w/ an R version of all data contained in the db, you can dive a little deeper into relations between the different data frames.

# How do the tables relate?

`tweats` table contains a column `user_id` = users that have posted the tweat. Similarly, `comments` contain both a `user_id` + a `tweat_id` col = which user posted a comment on which tweat.

Who posted the tweat on which somebody commented "awesome! thanks!" (comment 1012)?
```{r}
comments <- tables[[1]]
tweats <- tables[[2]]
users <- tables[[3]]

users[tweats[which(tweats$id == comments[which(comments$message == "awesome! thanks!"),"tweat_id"]),"user_id"],"name"]
```

# Query tweater (1)

Often be working w/ huge db's w/ tables w/ millions of rows. To do some analyses on this data, it's possible you only need a fraction of this data = good idea to send SQL queries to your db + only import data you actually need into R.

`dbGetQuery()` = pass connection object + a SQL query in the form of a character string

    Use dbGetQuery() to create a data frame, elisabeth, that selects the tweat_id column from the comments table where elisabeth is the commenter, her user_id is 1
    Print out elisabeth so you can see if you queried the database correctly.
```{r}
elisabeth <- dbGetQuery(con,"SELECT tweat_ID FROM comments WHERE user_id = 1")
elisabeth

elisabeth2 <- dbGetQuery(con,"SELECT a.tweat_ID FROM comments a 
                         LEFT JOIN users b
                          ON a.user_id = b.id
                         WHERE b.name = 'elisabeth'")
elisabeth2
```



    Create a data frame, latest, that selects the post column from the tweats table observations where the date is higher than '2015-09-21'.
    Print out latest.
```{r}
latest <- dbGetQuery(con, "SELECT post FROM tweats WHERE date > '2015-09-21'")
latest
```


    Create an R data frame, specific, that selects the message column from the comments table where the tweat_id is 77 and the user_id is greater than 4.
    Print specific.
```{r}
specific <- dbGetQuery(con, "SELECT message FROM comments WHERE tweat_id = 77 AND user_id > 4")
specific
```


    Create a data frame, short, that selects the id and name columns from the users table where the number of characters in the name is strictly less than 5.
    Print short.
```{r}
short <- dbGetQuery(con, "SELECT id, name FROM users WHERE CHAR_LENGTH(name) < 5")
short
```

```{r}
dbGetQuery(con, "SELECT post, message
  FROM tweats INNER JOIN comments on tweats.id = tweat_id
    WHERE tweat_id = 77")
```

# Send - Fetch - Clear

`dbGetQuery()` = a virtual function from the `DBI` package, but is actually implemented by the `RMySQL` package. Behind the scenes, the following steps are performed:
<ul>
<li> Sending the specified query with `dbSendQuery()`; </li>
<li> Fetching the result of executing the query on the database with `dbFetch()`; </li>
<li> Clearing the result with `dbClearResult()`. </li>
</ul>

Implement the steps above, tedious but gives the ability to fetch the query's result in chunks rather than all at once. (by specifying the `n` argument inside `dbFetch()`.


    Inspect the dbSendQuery() call that has already been coded for you. It selects the comments for the users with an id above 4.
    Use dbFetch() twice. In the first call, import only two records of the query result by setting the n argument to 2. In the second call, import all remaining queries (don't specify n). In both calls, simply print the resulting data frames.
    Clear res with dbClearResult().
```{r}
# Send query to the database + get back a MySQLResult object
res <- dbSendQuery(con, "SELECT * FROM comments WHERE user_id > 4")

# Use dbFetch() twice
dbFetch(res, n = 2) # get 2 records
dbFetch(res) # get all records

# Clear res
dbClearResult(res)
```

In our toy example, chopping up the fetches doesn't make a lot of sense, but make sure to remember this technique when you're struggling with huge databases!

Every time you connect to a db using `dbConnect()`, you're creating a *new* connection to the db you're referencing. `RMySQL` automatically specifies a max # of open connections + closes some connections for you, but still: it's always polite to manually disconnect from the db afterwards via `dbDisconnect()`.

    Using the technique you prefer, build a data frame long_tweats. It selects the post and date columns from the observations in tweats where the character length of the post variable exceeds 40.
    Print long_tweats.
    Disconnect from the database by using dbDisconnect().
```{r}
(long_tweats <- dbGetQuery(con, "SELECT post, date FROM tweats WHERE CHAR_LENGTH(post) > 40"))
dbDisconnect(con)
```




