---
title: "ImportingDataInR_Pt2"
author: "Steve Newns"
date: "December 5, 2017"
output: html_document
---

```{r}
library(ggplot2)
```
# Establish a connection

1st step to import data from a SQL db = creating a connection to it via different packages depending on db 

`dbConnect()` creates a connection between an R session + a SQL db  w/ 1st arg = a **DBIdriver object** that specifies how connections are made + how data is mapped between R + the db. For MySQL db's, you can build such a driver w/ ``RMySQL::MySQL()``. If the MySQL db is a remote db hosted on a server, you'll also have to specify the following arguments: `dbname`, `host`, `port`, `user` + `password.` Most of these details have already been provided.


    Load the DBI library, which is already installed on DataCamp's servers.
    Edit the dbConnect() call to connect to the MySQL database. Change the port argument (3306) and user argument ("student").
```{r}
# Load the DBI package
library(DBI)

# Edit dbConnect() call to connect to tweater db
con <- dbConnect(RMySQL::MySQL(), 
                 dbname = "tweater", 
                 host = "courses.csrrinzqubik.us-east-1.rds.amazonaws.com", 
                 port = 3306,
                 user = "student",
                 password = "datacamp")
class(con)
```

# List the database tables

Next step = see what tables db contains w/ `dbListTables()` which requires the connection object as an input + outputs a character vector w/ table names.
```{r}
tables <- dbListTables(con)
str(tables)
```

# Import users

DB contains data on a more tasty version of Twitter, **Tweater** where users can post tweats w/ short recipes for delicious snacks + comment on other tweats. There are 3 tables that have relations among

Importing the data on the users into the R session w/ `dbReadTable()` by passing the connection objec + the table to import. The resulting object is a standard R data frame.

```{r}
users <- dbReadTable(con,"users")
users
```

# Import all tables

Separate `dbReadTable()` calls for each + every table in a db = a lot of code duplication. Use `lapply()` to load in all tables in the database.

```{r}
(table_names <- dbListTables(con))
# apply read table to each table name + specify connection
tables <- lapply(table_names,dbReadTable, conn = con)
tables
```

Now w/ an R version of all data contained in the db, you can dive a little deeper into relations between the different data frames.

# How do the tables relate?

`tweats` table contains a column `user_id` = users that have posted the tweat. Similarly, `comments` contain both a `user_id` + a `tweat_id` col = which user posted a comment on which tweat.

Who posted the tweat on which somebody commented "awesome! thanks!" (comment 1012)?
```{r}
comments <- tables[[1]]
tweats <- tables[[2]]
users <- tables[[3]]

users[tweats[which(tweats$id == comments[which(comments$message == "awesome! thanks!"),"tweat_id"]),"user_id"],"name"]
```

# Query tweater (1)

Often be working w/ huge db's w/ tables w/ millions of rows. To do some analyses on this data, it's possible you only need a fraction of this data = good idea to send SQL queries to your db + only import data you actually need into R.

`dbGetQuery()` = pass connection object + a SQL query in the form of a character string

    Use dbGetQuery() to create a data frame, elisabeth, that selects the tweat_id column from the comments table where elisabeth is the commenter, her user_id is 1
    Print out elisabeth so you can see if you queried the database correctly.
```{r}
elisabeth <- dbGetQuery(con,"SELECT tweat_ID FROM comments WHERE user_id = 1")
elisabeth

elisabeth2 <- dbGetQuery(con,"SELECT a.tweat_ID FROM comments a 
                         LEFT JOIN users b
                          ON a.user_id = b.id
                         WHERE b.name = 'elisabeth'")
elisabeth2
```



    Create a data frame, latest, that selects the post column from the tweats table observations where the date is higher than '2015-09-21'.
    Print out latest.
```{r}
latest <- dbGetQuery(con, "SELECT post FROM tweats WHERE date > '2015-09-21'")
latest
```


    Create an R data frame, specific, that selects the message column from the comments table where the tweat_id is 77 and the user_id is greater than 4.
    Print specific.
```{r}
specific <- dbGetQuery(con, "SELECT message FROM comments WHERE tweat_id = 77 AND user_id > 4")
specific
```


    Create a data frame, short, that selects the id and name columns from the users table where the number of characters in the name is strictly less than 5.
    Print short.
```{r}
short <- dbGetQuery(con, "SELECT id, name FROM users WHERE CHAR_LENGTH(name) < 5")
short
```

```{r}
dbGetQuery(con, "SELECT post, message
  FROM tweats INNER JOIN comments on tweats.id = tweat_id
    WHERE tweat_id = 77")
```

# Send - Fetch - Clear

`dbGetQuery()` = a virtual function from the `DBI` package, but is actually implemented by the `RMySQL` package. Behind the scenes, the following steps are performed:
<ul>
<li> Sending the specified query with `dbSendQuery()`; </li>
<li> Fetching the result of executing the query on the database with `dbFetch()`; </li>
<li> Clearing the result with `dbClearResult()`. </li>
</ul>

Implement the steps above, tedious but gives the ability to fetch the query's result in chunks rather than all at once. (by specifying the `n` argument inside `dbFetch()`.


    Inspect the dbSendQuery() call that has already been coded for you. It selects the comments for the users with an id above 4.
    Use dbFetch() twice. In the first call, import only two records of the query result by setting the n argument to 2. In the second call, import all remaining queries (don't specify n). In both calls, simply print the resulting data frames.
    Clear res with dbClearResult().
```{r}
# Send query to the database + get back a MySQLResult object
res <- dbSendQuery(con, "SELECT * FROM comments WHERE user_id > 4")

# Use dbFetch() twice
dbFetch(res, n = 2) # get 2 records
dbFetch(res) # get all records

# Clear res
dbClearResult(res)
```

In our toy example, chopping up the fetches doesn't make a lot of sense, but make sure to remember this technique when you're struggling with huge databases!

Every time you connect to a db using `dbConnect()`, you're creating a *new* connection to the db you're referencing. `RMySQL` automatically specifies a max # of open connections + closes some connections for you, but still: it's always polite to manually disconnect from the db afterwards via `dbDisconnect()`.

    Using the technique you prefer, build a data frame long_tweats. It selects the post and date columns from the observations in tweats where the character length of the post variable exceeds 40.
    Print long_tweats.
    Disconnect from the database by using dbDisconnect().
```{r}
(long_tweats <- dbGetQuery(con, "SELECT post, date FROM tweats WHERE CHAR_LENGTH(post) > 40"))
dbDisconnect(con)
```

# Import flat files from the web

`utils` functions import flat file data, such as `read.csv()` + `read.delim()`, + are capable of automatically importing from URLs that point to flat files on the web. `readr`

    Load the readr package. It's already installed on DataCamp's servers.
    Use url_csv to read in the .csv file it is pointing to. Use the read_csv() function. The .csv contains column names in the first row. Save the resulting data frame as pools.
    Similarly, use url_delim to read in the online .txt file. Use the read_tsv() function and store the result as potatoes.
    Print pools and potatoes. Looks correct?
```{r}
# Load the readr package
library(readr)

# Import the csv file: pools
url_csv <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/swimming_pools.csv"
pools <- read_csv(url_csv)

# Import the txt file: potatoes
url_delim <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/potatoes.txt"
potatoes <- read_tsv(url_delim)

# Print pools and potatoes
pools
potatoes
```

There is a safer alternative to HTTP, HTTPS (https://) connections, (HypterText Transfer Protocol Secure) that you can use w/ the standard importing functions (since R version 3.2.2).

    Take a look at the URL in url_csv. It uses a secure connection, https://.
    Use read.csv() to import the file at url_csv. The .csv file it is referring to contains column names in the first row. Call it pools1.
    Load the readr package. It's already installed on DataCamp's servers.
    Use read_csv() to read in the same .csv file in url_csv. Call it pools2.
    Print out the structure of pools1 and pools2. Looks like the importing went equally well as with a normal http connection!
```{r}
# https URL to the swimming_pools csv file.
url_csv <- "https://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/swimming_pools.csv"

# Import the file using read.csv(): pools1
pools1 <- read.csv(url_csv)

# Load the readr package
library(readr)

# Import the file using read_csv(): pools2
pools2 <- read_csv(url_csv)

# Print the structure of pools1 and pools2
str(pools1)
str(pools2)
```

# Downloading Files
## Import Excel files from the web

When you learned about `gdata`, it was already mentioned that `gdata` can handle `.xls` files on the internet. `readxl` can't, at least not yet.

    Load the readxl and gdata packages. They are already installed on DataCamp's servers.
    Import the .xls file located at the URL url_xls using read.xls() from gdata. Store the resulting data frame as excel_gdata.
    You can not use read_excel() directly with a URL. Complete the following instructions to work around this problem:
    Use download.file() to download the .xls file behind the URL and store it locally as "local_latitude.xls".
    Call read_excel() to import the local file, "local_latitude.xls". Name the resulting data frame excel_readxl.
```{r}
# Load the readxl and gdata package
library(readxl)
library(gdata)

# Specification of url: url_xls
url_xls <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/latitude.xls"

# 1) Import the .xls file DIRECLTY with gdata: excel_gdata
excel_gdata <- gdata::read.xls(url_xls)

# 2a)  file behind URL, name it local_latitude.xls
download.file(url_xls, destfile = 'local_latitude.xls')
# 2b) Import the local .xls file with readxl: excel_readxl
excel_readxl <- read_excel('local_latitude.xls')
```

# Downloading any file, secure or not

Can read excel files on the web using `read_excel` by 1st downloading the file w/ `download.file()`.

There's more: with `download.file()` you can download ANY kind of file from the web using HTTP + HTTPS: images, executable files, but also .RData files (very efficient format to store R data)

You can load data from an RData file using `load()`, but this function does not accept a URL string as an argument. 

    Take a look at the URL in url_rdata. It uses a secure connection, https://. This URL points to an RData file containing a data frame with some metrics on different kinds of wine.
    Download the file at url_rdata using download.file(). Call the file "wine_local.RData" in your working directory.
    Load the file you created, wine_local.RData, using the load() function. It takes one argument, the path to the file, which is just the filename in our case. After running this command, the variable wine will automatically be available in your workspace.
    Print out the summary() of the wine dataset.
```{r}
# https URL to wine RData file.
url_rdata <- "https://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/wine.RData"

# Download the wine file to your working directory
download.file(url_rdata, "wine_local.RData")

# Load the wine data into your workspace using load(path) --> wine variable is now available
load("wine_local.RData")

# Print out the summary of the wine data
summary(wine)
```

Another way to load remote RData files is to use `url()` inside `load()`. However, this will not save the RData file to a local file. 

# Reading a text file from the web

Which way of importing data is NOT possible?
<ul>
<li> Importing a .csv file residing on the web using the URL with read.csv(): </li>
<li> Downloading a remote excel and saving it to your working directory using download.file():</li>
<li> Importing a .txt file residing on the web using the URL with read_tsv():</li>
<li> **Using the load() function to load a remote RData file into the workspace with only the URL string:** -> need to be local </li> 
```{r}
read.csv("http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/swimming_pools.csv")
read_tsv("http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/potatoes.txt")
download.file("http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/latitude.xlsx", "lat.xlsx")
load("https://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/wine.RData")
```

# HTTP? httr!

Downloading a file from the Internet = sending a **GET request** + receiving the file you asked for. Internally, all previously-discussed functions use a GET request to download files.

`httr` provides `GET()` to execute this GET request w/ a result = a **response object** that provides easy access to the **status code, content-type** and, of course, the actual content.

You can extract content from the request using `content()` function. At the time of writing, there are 3 ways to retrieve this content: as a raw object, character vector, or an R object, such as a list. If you don't tell `content()` how to retrieve the content through `as` argument, it'll try its best to figure out which type is most appropriate based on the content-type.

    Load the httr package. It's already installed on DataCamp's servers.
    Use GET() to get the URL stored in url. Store the result of this GET() call as resp.
    Print the resp object. What information does it contain?
    Get the content of resp using content() and set the as argument to "raw". Assign the resulting vector to raw_content.
    Print the first values in raw_content with head().
```{r}
library(httr)

# Get the url, save response to resp
url <- "http://www.example.com/"
resp <- GET(url)

# Print resp
resp

# Get the raw content of resp: raw_content
raw_content <- content(resp, as = "raw")

# Print the head of raw_content
head(raw_content)
```

The raw content of the response doesn't make a lot of sense. Luckily, `content()` by default, if you don't specify `as` argument, figures out what type of data you're dealing with and parses it for you.

Web content does not limit itself to HTML pages + files stored on remote servers such as DataCamp's Amazon S3 instances. There are many other data formats out there. A very common one is **JSON**, a format very often used by so-called Web APIs, interfaces to web servers w/ which you as a client can communicate to get or store info in more complicated ways.


    Use GET() to get the url that has already been specified in the sample code. Store the response as resp.
    Print resp. What is the content-type?
    Use content() to get the content of resp. Set the as argument to "text". Simply print out the result. What do you see?
    Use content() to get the content of resp, but this time do not specify a second argument. R figures out automatically that you're dealing with a JSON, and converts the JSON to a named R list.
```{r}
# Get the url
url <- "http://www.omdbapi.com/?apikey=ff21610b&t=Annie+Hall&y=&plot=short&r=json"
resp <- GET(url)

# Print resp
resp

# Print content of resp as text
content(resp, as = "text")

# Print content of resp
content(resp)
```

The fact that httr converts the JSON response body automatically to an R list is very convenient. 

# From JSON to R