---
title: "Writing Functions in R"
author: "Steve Newns"
date: "December 19, 2017"
output: html_document
---
# Quick Refresher

## Writing a function

The function template is a useful way to start writing a function:

my_fun <- function(arg1, arg2) {
  # body
}

my_fun is the variable that you want to assign your function to, arg1 and arg2 are arguments to the function. The template has two arguments, but you can specify any number of arguments, each separated by a comma. You then replace # body with the R code that your function will execute, referring to the inputs by the argument names you specified.

Let's get to writing functions!
Instructions
100xp

    Using the function template, write a function ratio() that takes arguments x and y and returns their ratio, x / y.
    Check your function by calling it with the arguments 3 and 4.
```{R}

```
Beyond the first couple of arguments you should always use matching by name. It makes your code much easier for you and others to read. This is particularly important if the argument is optional, because it has a default. When overriding a default value, it's good practice to use the name.

Notice that when you call a function, you should place a space around = in function calls, and always put a space after a comma, not before (just like in regular English). Using whitespace makes it easier to skim the function for the important components.
Instructions
100xp

Here's a rather cryptic call to mean(). Rewrite the call to follow the best practices we just discussed. Use the natural ordering of the arguments, which you can find by typing ?mean in the console.
```{r}
# Rewrite the call to follow best practices
# mean(0.1,x=c(1:9, NA),TRUE)
mean(c(1:9, NA), trim = 0.1, na.rm = TRUE)
```

If y is not passed in as an argument to the function, R looks outside of the function environment. 
y <- 10
f <- function(x) {
  x + y
}
f(10) = 20

y <- 10
f <- function(x) {
  y <- 5
  x + y
}
f(10) = 15

he value of x is passed in as an argument to the function and the value of y is defined inside of the functio

f <- function(x) {
  y <- 5
  x + y
}
f(5)
ERROR = Even though y is set equal to 5 within the body of the function, the object does not exist in the global environment. 

Atomic types of vectors in R = logical
integer
double
character


## Subsetting lists

There are a few ways to subset a list. Throughout the course we'll mostly use double bracket ([[]]) subsetting by index and by name.

That is, my_list[[1]] extracts the first element of the list my_list, and my_list[["name"]] extracts the element in my_list that is called name. If the list is nested you can travel down the heirarchy by recursive subsetting. For example, mylist[[1]][["name"]] is the element called name inside the first element of my_list.

A data frame is just a special kind of list, so you can use double bracket subsetting on data frames too. my_df[[1]] will extract the first column of a data frame and my_df[["name"]] will extract the column named name from the data frame.

I've set up a list called tricky_list in your workspace. Use the function typeof() combined with double bracket subsetting to answer the following questions.
Instructions
100xp

What type of object is the...

    2nd element in tricky_list?
    Element called x in tricky_list?
    2nd element inside the element called x in tricky_list?
```{r}
# 2nd element in tricky_list
tricky_list[[2]]
typeof(tricky_list[[2]])
# Element called x in tricky_list
tricky_list[["x"]]
typeof(tricky_list[["x"]])
# 2nd element inside the element called x in tricky_list
tricky_list[["x"]][[2]]
typeof(tricky_list[["x"]][[2]])
```

## Exploring lists

Often you won't know exactly what is inside a list. But, you'll need to figure it out to get some useful piece of data. Extracting elements from the output of the names() and str() functions is a great way to explore the structure of a list.

Calling names() on a list will give you names at the top level of the list and str() will give you a full description of the entire list (which can sometimes be a little overwhelming).

tricky_list has a regression model stored in it. Let's see if we can drill down and pull out the slope estimate corresponding to the wt variable.
Instructions
100xp

    Use names() on tricky_list to guess where the regression model is stored.
    Use names() and str() on the model element of tricky_list.
    Extract the coefficients element of the model element of tricky_list.
    Extract the wt element of the coefficients element of the model element of tricky_list
```{r}
# Guess where the regression model is stored
names(tricky_list)

# Use names() and str() on the model element
names(tricky_list[["model"]])
str(tricky_list[["model"]])

# Subset the coefficients element
tricky_list[["model"]][["coefficients"]]

# Subset the wt element
tricky_list[["model"]][["coefficients"]][["wt"]]
```
 Another way is to use the coef() function on the model element of tricky_list. 
 
## A safer way to create the sequence

Let's take a look at the sequence component of our for loop:

i in 1:ncol(df)

Each time our for loop iterates, i takes the next value in 1:ncol(df). This is a pretty common model for a sequence: a sequence of consecutive integers designed to index over one dimension of our data.

What might surprise you is that this isn't the best way to generate such a sequence, especially when you are using for loops inside your own functions. Let's look at an example where df is an empty data frame:

df <- data.frame()
1:ncol(df)

for (i in 1:ncol(df)) {
  print(median(df[[i]]))
}

Our sequence is now the somewhat non-sensical: 1, 0. You might think you wouldn't be silly enough to use a for loop with an empty data frame, but once you start writing your own functions, there's no telling what the input will be.

A better method is to use the seq_along() function. This function generates a sequence along the index of the object passed to it, but handles the empty case much better.
Instructions
100xp

    Replace the 1:ncol(df) sequence in our for loop with seq_along(df).
    Change the value of df to an empty data frame.
    Repeat the for loop to verify there is no error.
```{r}
#for (i in 1:ncol(df)) {
  #print(median(df[[i]]))
#}
# Replace the 1:ncol(df) sequence
for (i in seq_along(df)) {
  print(median(df[[i]]))
}

# Change the value of df
df <- data.frame()

# Repeat for loop to verify there is no error
for (i in seq_along(df)) {
  print(median(df[[i]]))
}
```

##Keeping output

Our for loop does a good job displaying the column medians, but we might want to store these medians in a vector for future use.

Before you start the loop, you must always allocate sufficient space for the output, let's say an object called output. This is very important for efficiency: if you grow the for loop at each iteration (e.g. using c()), your for loop will be very slow.

A general way of creating an empty vector of given length is the vector() function. It has two arguments: the type of the vector ("logical", "integer", "double", "character", etc.) and the length of the vector.

Then, at each iteration of the loop you must store the output in the corresponding entry of the output vector, i.e. assign the result to output[[i]]. (You might ask why we are using double brackets here when output is a vector. It's primarily for generalizability: this subsetting will work whether output is a vector or a list.)

Let's edit our loop to store the medians, rather than printing them to the console.
Instructions
100xp

    Create a new "double" vector of length ncol(df) and name it output.
    Inside the body of the loop, store the result of each iteration result in the ith element of output.
    Print output to the console.
```{R}
# Create new double vector: output
output <- vector(mode = "double", length = ncol(df))

# Alter the loop
for (i in seq_along(df)) {
  # Change code to store result in output
  output[[i]] <- median(df[[i]])
}

# Print output
output
```

# When and how you should write a function

##Start with a snippet of code

We have a snippet of code that successfully rescales a column to be between 0 and 1:

(df$a - min(df$a, na.rm = TRUE)) /  
  (max(df$a, na.rm = TRUE) - min(df$a, na.rm = TRUE))

Our goal over the next few exercises is to turn this snippet, written to work on the a column in the data frame df, into a general purpose rescale01() function that we can apply to any vector.

The first step of turning a snippet into a function is to examine the snippet and decide how many inputs there are, then rewrite the snippet to refer to these inputs using temporary names. These inputs will become the arguments to our function, so choosing good names for them is important. (We'll talk more about naming arguments in a later exercise.)

In this snippet, there is one input: the numeric vector to be rescaled (currently df$a). What would be a good name for this input? It's quite common in R to refer to a vector of data simply as x (like in the mean function), so we will follow that convention here.
Instructions
100xp

    To help you test your code, create a vector x containing the numbers 1 through 10.
    Rewrite the code snippet to use the temporary variable x instead of referring to the data frame column df$a.
```{r}
# Define example vector x
x <- seq(20)

# Rewrite this snippet to refer to x
#(df$a - min(df$a, na.rm = TRUE)) /
  #(max(df$a, na.rm = TRUE) - min(df$a, na.rm = TRUE))
(x - min(x, na.rm = TRUE)) /
  (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
```

## Rewrite for clarity

Our next step is to examine our snippet and see if we can write it more clearly.

Take a close look at our rewritten snippet. Do you see any duplication?

(x - min(x, na.rm = TRUE)) /
  (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))

One obviously duplicated statement is min(x, na.rm = TRUE). It makes more sense for us just to calculate it once, store the result, and then refer to it when needed. In fact, since we also need the maximum value of x, it would be even better to calculate the range once, then refer to the first and second elements when they are needed.

What should we call this intermediate variable? You'll soon get the message that using good names is an important part of writing clear code! I suggest we call it rng (for "range").
Instructions
100xp

    Define the intermediate variable rng to contain the range of x using the function range(). Specify the na.rm() argument to automatically ignore any NAs in the vector.
    Rewrite the snippet to refer to this intermediate variable.
```{r}
# Define example vector x
x <- seq(10) #1:10

# Define rng
rng <- range(x, na.rm = T)

# Rewrite this snippet to refer to the elements of rng
#(x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
(x - rng[1]) / (rng[2] - rng[1])

```

##Finally turn it into a function!

What do you need to write a function? You need a name for the function, you need to know the arguments to the function, and you need code that forms the body of the function.

We now have all these pieces ready to put together. It's time to write the function!
Instructions
100xp

    Use the function template to create a function called rescale01():
    The function should take a single argument x.
    The body of the function should be our rewritten snippet from the previous exercise: rng <- range(x, na.rm = TRUE) (x - rng[1]) / (rng[2] - rng[1])
    Once you've written the function, test it out by calling it on the x we've already defined.
```{R}
# Define example vector x
x <- 1:10 

# Use the function template to create the rescale01 function
rescale01 <- function(x) {
   rng <- range(x, na.rm = TRUE) 
   return((x - rng[1]) / (rng[2] - rng[1]))
  
}

# Test your function, call rescale01 using the vector x as the argument
rescale01(x)
```

# How to write a function: Start with a simple problem

Let's tackle a new problem. We want to write a function, both_na() that counts at how many positions two vectors, x and y, both have a missing value.

How do we get started? Should we start writing our function?

both_na <- function(x, y) {
  # something goes here?
}

No! We should start by solving a simple example problem first. Let's define an x and y where we know what the answer both_na(x, y) should be.

Let's start with:

x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3,  4)

Then both_na(x, y) should return 1, since there is only one element that is missing in both x and y, the third element.

(Notice we introduced a couple of extra spaces to each vector. Adding spaces to x and y to make them match up makes it much easier to see what the correct result is. Code formatting is an important aspect of writing clear code.)

Your first task is to write a line of code that gets to that answer.
Instructions
100xp

Write a line of code that finds the number of positions where both x and y have missing values. You may find the is.na() and sum() functions useful, as well as the & operator.
```{r}
# Define example vectors x and y
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3,  4)

# Count how many elements are missing in both x and y
sum(is.na(x) & is.na(y))
```

## 2) Rewrite snippet as function

Great! You now have a snippet of code that works:

sum(is.na(x) & is.na(y))

You've already figured out it should have two inputs (the two vectors to compare) and we've even given them reasonable names: x and y. Our snippet is also so simple we can't write it any clearer.

So, let's go ahead and turn this snippet into a function!
Instructions
100xp

    This time, instead of using the function template, we've provided the body of the function. Wrap the body with function assignment and curly braces.
    Assign the function to the name both_na().
    both_na() should take two arguments, x and y.
```{r}
# Turn this snippet into a function: both_na()
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```

## Put our function to use

We have a function that works in at least one situation, but we should probably check it works in others.

Consider the following vectors:

x <-  c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

What would you expect both_na(x, y1) to return? What about both_na(x, y2)? Does your function return what you expected? Try it and see!

```{R}
# Define x, y1 and y2
x <-  c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

# Call both_na on x, y1
both_na(x,y1)

# Call both_na on x, y2
both_na(x,y2) # ERROR
```


## How can you write a good function?: 1) Good function names

Consider the following function, f2(), which has also been loaded into your workspace:

f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}

f2() isn't a very good function name! Which of the following would make a good name for this function? = remove_last() = a verb + descriptive

## Argument names

It's not just your function that needs a good name. Your arguments need good names too!

Take a look at this function, which calculates a confidence interval for a population mean:

mean_ci <- function(c, nums) {
  se <- sd(nums) / sqrt(length(nums))
  alpha <- 1 - c
  mean(nums) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

The argument nums is a sample of data and the argument c controls the level of the confidence interval. For example, if c = 0.95, we get a 95% confidence interval. Are c and nums good arguments names?

c is a particularly bad name. It is completely non-descriptive and it's the name of an existing function in R. What might be better? Maybe something like confidence, since it reveals the purpose of the argument: to control the level of confidence for the interval. Another option might be level, since it's the same name used for the confint function in base R and your users may already be familiar with that name for this parameter.

nums is not inherently bad, but since it's the placeholder for the vector of data, a name like x would be more recognizable to users.
Instructions
100xp

Rewrite the mean_ci() function to take arguments named level and x instead of c and nums, in that order for now.
```{r}
# Rewrite mean_ci to take arguments named level and x
mean_ci <- function(c, nums) {
  se <- sd(nums) / sqrt(length(nums))
  alpha <- 1 - c
  mean(nums) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

mean_ci2 <- function(level, x) {
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - level
  mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

mean_ci(.95,1:10)
```

## Argument order

Aside from giving your arguments good names, you should put some thought into what order your arguments are in and if they should have defaults or not.

Arguments are often one of two types:

    Data arguments supply the data to compute on.
    Detail arguments control the details of how the computation is done.

Generally, data arguments should come first. Detail arguments should go on the end, and usually should have default values.

Take another look at our function:

mean_ci <- function(level, x) {
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - level
  mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

x is the data to be used for the calculation, so it's a data argument and should come first.

level is a detail argument; it just controls the level of confidence for which the interval is constructed. It should come after the data arguments and we should give it a default value, say 0.95, since it is common to want a 95% confidence interval.
Instructions
100xp

    Move the data argument, x, to the front.
    Move the detail argument, level, to the end and give it the default value 0.95.
```{r}
# Alter the arguments to mean_ci
mean_ci <- function(level, x) {
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - level
  mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

# Alter the arguments to mean_ci
mean_ci2 <- function(x, level = .95) {
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - level
  mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}
```

## Return statements

One of your colleagues has noticed if you pass mean_ci() an empty vector it returns a confidence interval with missing values at both ends (try it: mean_ci(numeric(0))). In this case, they decided it would make more sense to produce a warning "x was empty" and return c(-Inf, Inf) and have edited the function to be:

mean_ci <- function(x, level = 0.95) {
  if (length(x) == 0) {
    warning("`x` was empty", call. = FALSE)
    interval <- c(-Inf, Inf)
  } else { 
    se <- sd(x) / sqrt(length(x))
    alpha <- 1 - level
    interval <- mean(x) + 
      se * qnorm(c(alpha / 2, 1 - alpha / 2))
  }
  interval
}

Notice how hard it is now to follow the logic of the function. If you want to know what happens in the empty x case, you need to read the entire function to check if anything happens to interval before the function returns. There isn't much to read in this case, but if this was a longer function you might be scrolling through pages of code.

This is a case where an early return() makes sense. If x is empty, the function should immediately return c(-Inf, Inf).
Instructions
100xp

Edit the mean_ci function using an if statement to check for the case when x is empty and if so, to produce the same warning as the code above then immediately return() c(-Inf, Inf).
```{r}
# Alter the mean_ci function
mean_ci <- function(x, level = 0.95) {
  if (length(x) == 0) {
    warning("'x' was empty", call. = F)
    interval <- c(-Inf,Inf)
  } else {
    se <- sd(x) / sqrt(length(x))
    alpha <- 1 - level
    interval <- mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
  }
  return(interval)  
}
```

## What does this function do?

Over the next few exercises, we'll practice everything you've learned so far. Here's a poorly written function, which is also available in your workspace:

f <- function(x, y) {
  x[is.na(x)] <- y
  cat(sum(is.na(x)), y, "\n")
  x
}

Your job is to turn it in to a nicely written function.

What does this function do? Let's try to figure it out by passing in some arguments.
Instructions
100xp

    Define a numeric vector x with the values: 1, 2, NA, 4 and 5.
    Call f() with the arguments x = x, and y = 3.
    Call f() with the arguments x = x, and y = 10.
```{r}
# Define a numeric vector x with the values 1, 2, NA, 4 and 5
x <- c(1, 2, NA, 4, 5)

f <- function(x, y) {
  x[is.na(x)] <- y # sub in argument value for NA values
  cat(sum(is.na(x)), y, "\n") # print out how many nulls we have after substitution, substitution value, new line, then final vector
  x
}

# Call f() with the arguments x = x and y = 3
f(x,3)

# Call f() with the arguments x = x and y = 10
f(x,10)
```

## Let's make it clear from its name

Did you figure out what f() does? f() takes a vector x and replaces any missing values in it with the value y.

Imagine you came across the line df$z <- f(df$z, 0) a little further on in the code. What does that line do? Now you know, it replaces any missing values in the column df$z with the value 0. But anyone else who comes across that line is going to have to go back and find the definition of f and see if they can reason it out.

Let's rename our function and arguments to make it obvious to everyone what we are trying to achieve.
Instructions


    Rename the function f() to replace_missings().
    Change the name of the y argument to replacement.
    Now replace the missing values of df$z with 0's using your new function. Make sure you assign the result back to df$z.
```{r}
# Rename the function f() to replace_missings()
replace_missings <- function(x, replacement) {
  x[is.na(x)] <- replacement
  cat(sum(is.na(x)), replacement, "\n")
  x
}

# Rewrite the call on df$z to match our new names
df$z <- replace_missings(df$z, 0)
```

Make the body more understandable

Great! Now when someone comes across

df$z <- replace_missings(df$z, replacement = 0)

in your code, it's pretty obvious what you are trying to achieve. The body of our replace_missings() function is still a little messy. There is certainly some unnecessary duplication.
Instructions
100xp

    Define is_miss, a logical that identifies the missing values in x.
    To reduce unncessary duplication, alter the rest of function to refer to is_miss instead of is.na(x).
```{r}
replace_missings <- function(x, replacement) {
  # Define is_miss
  is_miss <- is.na(x)
  
  # Rewrite rest of function to refer to is_miss
  x[is_miss] <- replacement
  cat(sum(is_miss), replacement, "\n")
  x
}
```

## Much better! But a few more tweaks...

Did you notice replace_missings() prints some output to the console? Try it:

replace_missings(df$z, replacement = 0)

That output isn't exactly self-explanatory. It would be much nicer to say "2 missing values replaced by the value 0".

It is also bad practice to use cat() for anything other than a print() method (a function designed just to display output). Having an important message just print to the screen makes it very hard for other people who might be programming with your function to capture the output and handle it appropriately.

The official R way to supply simple diagnostic information is the message() function. The unnamed arguments are pasted together with no separator (and no need for a newline at the end) and by default are printed to the screen.

Let's make our function nicer for users by using message() and making the output self-contained.
Instructions
100xp

    Rewrite to use message() and output "sum(is_miss) missings replaced by the value replacement."
    Check your new function by replacing all the missing values of df$z with 0's, as we've done before.
```{r}
replace_missings <- function(x, replacement) {
  is_miss <- is.na(x)
  x[is_miss] <- replacement
  
  # Rewrite to use message()
  message(sum(is_miss), " missings replaced by the value ", replacement)
  x
}

# Check your new function by running on df$z
replace_missings(df$z, 0)
```