---
title: "Lab1 - baseball stats"
---

[Moneyball](http://en.wikipedia.org/wiki/Moneyball_(film)) focuses on the "quest for the secret of success in baseball". It follows a low-budget team, the Oakland A's, who believed that underused statistics, such as a player's ability to get on base, better predict the ability to score runs, rather than typical statistics like HR's, RBI's, and BA. Obtaining players who excelled in these underused statistics turned out to be much more affordable for the team.

We'll be looking at data from all 30 MLB teams and examining the relationships between in-game statistics. 

```{r load-abrbuthnot-data, eval=TRUE}
# access the OpenIntro website and fetch the `mlb11` data set.
download.file("http://www.openintro.org/stat/data/mlb11.RData", destfile = "mlb11.RData")
load("mlb11.RData")
```

Let's load up the data for the 2011 season. In addition to `RS`, there are 7 traditionally-used variables in the data set: `{AB, H, HR, BA, SO, SB, W}`. There are also 3 newer variables: `OBP`, slugging % `SLG`, + on-base plus slugging `OBPS`. For the 1st portion of the analysis, we'll consider the 7 traditional variables. At the end of the lab, we'll work w/ the newer variables.

`mlb11` = a **data frame** w/ 30 observations of 12 variables. 

## The Data: 2011 baseball statistics

Each row = different team w/ an index (not actually part of dataset), team name, + remaining cols = team-specific metrics

```{r, eval=TRUE}
# get observations/rows and variables/cols
dim(mlb11)

# get col names
names(mlb11)
```

## Some Exploration

```{r, eval=FALSE}
# returns col as a vector
mlb11$runs
```

```{r}
head(mlb11)
tail(mlb11)
```

### mosaic

`mosaic` package streamlines all commands needed in course = **specifically designed by a team of NSF-funded educators to make R more accessible to statistics students** = doesn't provide new functionality, so much as make existing functionality more logical, consistent, all the while emphasizing importants concepts in statistics. 

```{r, eval=F,message=F,warning=F}
#install.packages("mosaic")
library(mosaic)
library(lattice)
library(tidyverse)
```

Centerpiece of the `mosaic` syntax = the use of the *modeling language*, which involves the use of a tilde (~), which can be read as *"is a function of"*. For example, to create a simple plot of runs each team scored vs. hits:

```{r, eval=FALSE, message=FALSE}
# simple plot of runs each team scored vs. H
xyplot(runs ~ hits, mlb11)
```

By default = scatterplot w/ each $(x,y)$ pair indicated by an open circle. 

**2.  Is there an association between runs scored and hits? How would you describe it?**

* *Yes, moderate-to-strongly positive*

Now plot only total # of runs.

```{r, eval=FALSE}
histogram( ~ runs, mlb11)
bwplot( ~ runs, mlb11)
```

**3.  Describe the distribution of runs scored among MLB teams in 2011.  What is the median?**

* *We have a right-skewed distribution, with a median value of just above 700. We know from the skew that the mean **should** be higher than the median, as it would be pulled towards the majority of the data points to the right of the distribution's center.*

```{r}
median(mlb11$runs)
mean(mlb11$runs)
```

### Additional R functionalities

Get the total # of HR and SO by the Texas Rangers (1st row in data), + to do so every row = add the vectors together + R computes all sums simultaneously.

```{r, eval=F,message=F}
# all teams
mlb11$homeruns + mlb11$strikeouts
# just rangers
mlb11[1,"homeruns"] + mlb11[1,"strikeouts"]
# tidyverse way
library(tidyverse)
mlb11 %>%
  filter(team == "Texas Rangers") %>%
  mutate(sumHR_SO = homeruns+strikeouts) %>%
  select(sumHR_SO)
```

**4.  Which team had the highest cumulative number of strikeouts and homeruns in 2011?**

```{r}
library(tidyverse)

mlb11[which.max(mlb11$homeruns + mlb11$strikeouts),"team"]

# tidyway
mlb11 %>%
  mutate(sumHR_SO = homeruns+strikeouts) %>%
  select(team,sumHR_SO) %>%
  arrange(desc(sumHR_SO)) %>%
  head()
```

**Which home run observations were greater than 200?**

```{r, eval=FALSE}
mlb11$homeruns > 200
```

5.  Which MLB teams had more than 200 home runs in 2011?

```{r}
subset(mlb11, homeruns > 200)

# tidyway
mlb11 %>%
  filter(homeruns>200) %>%
  select(team)
```

## Summaries and tables

***Good 1st step in any analysis = distill all the info from the data into a few summary statistics + graphics***
```{r, eval=FALSE}
summary(mlb11$runs)
```

To compute IQR for team runs, one could look at the output above + enter
```{r, eval=FALSE}
734 - 629
```

Built-in functions to compute summary stats one-by-one, like mean, median, and variance of `runs`, type 

```{r, eval=FALSE, message=FALSE}
mean(~runs, data=mlb11)
var(~runs, data=mlb11)
median(~runs, data=mlb11)
```

Command `mosaic::favstats()`, allows us to compute all of this info (+ more) at once, in a data frame. 
```{r, eval=FALSE}
mosaic::favstats(~runs, data=mlb11)
```

While it makes sense to describe a quantitative variable like `runs` in terms of these statistics, what about *categorical* data?  Instead consider the **sample frequency** or **relative frequency distribution**. `tally()` does this by counting the # of times each kind of response was given.
```{r, eval=FALSE}
# see # of teams who hit at least 200 homeruns
mlb11$HighHR <- mlb11$homeruns >= 200
mlb11$HighHR

# counts of records with HighHR
mosaic::tally(~ HighHR, data=mlb11)
```

or instead look at the **relative frequency distribution** by including a new arg
```{r, eval=FALSE}
mosaic::tally(~ HighHR, data=mlb11, format="proportion")
```

In each of the above steps, we make use of a new variable, `HighHR` + use it in as the input for `tally()`.

Next, we make a bar chart of entries in the table by putting the table inside `barchart()`.
```{r, eval=FALSE}
barchart(tally(~HighHR, data=mlb11, margins=FALSE), horizontal=FALSE)
```

We've computed the table then immediately applied the graphical function, `barchart`.

6.  Create a numerical summary for `stolen_bases` and `wins`, and compute the interquartile range for each. 
```{r}
mosaic::favstats(mlb11$stolen_bases)
mosaic::favstats(mlb11$stolen_bases)$max - mosaic::favstats(mlb11$stolen_bases)$min 

mosaic::favstats(mlb11$wins)
mosaic::favstats(mlb11$wins)$max - mosaic::favstats(mlb11$wins)$min
```

7.  Compute the relative frequency distribution for teams with at least 90 `wins`. How many teams reached 90 wins?  
```{r}
mlb11$HighW <- mlb11$wins >= 90
length(mlb11$HighW)
mosaic::tally(~ HighW, data=mlb11, format="proportion")
```

`tally` can be used to tabulate any # of variables you provide.  

8.  What is shown in the following table?  How many teams won >= 90 games + hit >= 200 HR? 
```{r table-smoke-gender, eval=FALSE}
mlb11$HighWins <- mlb11$wins >=90
mosaic::tally(HighWins ~ HighHR, data=mlb11, format="count")
```

Above shows the # of teams that had at least 90 W *and* at least 200 HR in a matrix. Only 3 teams had done so.

9. Describe the center, shape, + spread of `stolen_bases`, using an appropriate plot + the appropriate metrics.
```{r}  
lattice::histogram( ~ stolen_bases, mlb11)
lattice::bwplot( ~ stolen_bases, mlb11)
mosaic::favstats(mlb11$stolen_bases)
```

This is pretty normal, with a median and mean just below 110 SB's, with a spread of about +/- 30 SB's.

**10.  A coach is interested in the link between `stolen_bases` and `runs`. Show the coach a scatter plot, + describe the association. As you make the plot, think carefully about which of these 2 variables is the explanatory variable (+ which is the response).**
```{r}
### Enter code here.
#plot(mlb11$stolen_bases,mlb11$runs)
xyplot(runs ~ stolen_bases, mlb11)
cor(mlb11$stolen_bases,mlb11$runs)
```

There seems to be no association between `stolen_bases` (factor) and `runs` (response), as we can see from the scatterplot and the incredibly low Pearsons' correlation coefficient, $r$.

**11. How can you change the x and y labels on your plots? How can you add a title?**
```{r}
### Enter code here.
library(ggplot2)
ggplot(mlb11, aes(stolen_bases,runs)) +
  geom_point() + 
  xlab("Stolen Bases") + 
  ylab("Runs") + 
  ggtitle("Runs vs. Stolen Bases") + 
  theme_bw()

xyplot(runs ~ stolen_bases, mlb11, xlab='Stolen Bases', ylab='Runs', main='R vs. SB')

cor(mlb11$stolen_bases,mlb11$runs)
```

**12. Using visual evidence, find the variable that you think seems to boast the strongest association to `runs`. Consider any continuous variables between columns 3 (`at_bats`) and 12 (`new_obs`).**

```{r}
#names(mlb11)
corr_mat <- cor(mlb11$runs,mlb11[,-1])
corrplot::corrplot(corr_mat)
corrplot::corrplot(corr_mat,method = "number")

pairs(~runs+bat_avg+new_onbase+new_slug+new_obs
      ,data=mlb11
      ,main="Simple Scatterplot Matrix")
#cor(mlb11[-1])
```

`new_obs` seems to boast the strongest association to `runs`, followed by `new_onbase` and `new_slug`, and `bat_avg`

**13. What is the variance in the number of strikeouts for each team during the 2011 season?**
```{r}
var(mlb11$strikeouts)
```


**14. Make a new variable, `high_BA`, to represent teams that hit for a BA >= 0.270. How many teams fit into this group?**

```{r}
mlb11$high_BA <- mlb11$bat_avg >= .270
sum(mlb11$high_BA)
```

**15. Use `mosaic::favstats` to get the summary statistics w/in each `high_BA` group, + `bwplot()` to make boxplots. For example, what appears to be the link between `homeruns` + `high_BA`? Do teams w/ higher BA tend to hit more HR?**

```{r, }
mosaic::favstats(homeruns ~ high_BA, data = mlb11)
lattice::densityplot(~ homeruns, groups = high_BA, data = mlb11, auto.key = TRUE)
lattice::bwplot(mlb11$homeruns, groups = mlb11$high_BA, data = mlb11)
```

It appears that teams w/ lower BA do indeed to have lower home runs, via the lower median and mean, as well as via the larger AUC of the TRUE curve in the density plots

**16. Repeat the code above, only using `stolen_bases` instead of `homeruns`. Does there appear to be a link between `stolen bases` and `high_BA`?**

```{r, }
mosaic::favstats(stolen_bases ~ high_BA, data = mlb11)
lattice::densityplot(~ stolen_bases, groups = high_BA, data = mlb11, auto.key = TRUE)
```

The centers of the distributions are similary, but the spread for higher BA teams is significantly larger than for lower BA teams

<div id="license">
Portions of this lab were adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel from a lab written by Mark Hansen of UCLA Statistics, a product of OpenIntro that is released under a 
[Creative Commons License](http://creativecommons.org/licenses/by-sa/3.0)
</div>