---
title: "Ch13_RelationalData"
author: "Steve Newns"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(nycflights13)
```

Rare that a data analysis involves only a single table of data = must usually combine multiple tables to answer questions

Multiple tables of data = **relational data** b/c it is the **relations**, not just the individual datasets, that are important.

**Relation** = always defined between a pair of tables + all other relations are built up from this simple idea: *relations of 3+ tables are always a property of the relations between each pair*. 

Sometimes both elements of a pair can be the same table! This is needed if, for example, you have a table of people + each person has a reference to their parents.

To work w/ relational data you need verbs that work w/ pairs of tables --> 3 families of verbs designed to work with relational data:
<ul>
<li> **Mutating joins** = add new variables to 1 data frame from matching observations in another. </li>
<li> **Filtering joins** = filter observations from 1 data frame based on whether or not they match an observation in the other table. </li>
<li> **Set operations** = treat observations as if they were **set elements**. </li>
</ul>

Most common place to find relational data = RDBMS (encompasses almost all modern databases) 

SQL expression in `dplyr` is a little different = a little easier b/c `dplyr` is specialised to do data analysis + makes common data analysis operations easier at the expense of making it more difficult to do other things not commonly needed for data analysis.

## 13.2 nycflights13

`nycflights13` = 4 tibbles related to the `flights` table

`airlines` = look up full carrier name from its abbreviated code:
```{r}
airlines
```

`airports` gives info about each airport, IDed by FAA airport code:

```{r}
airports
```

`planes` gives info about each plane, IDed by its `tailnum`:

```{r}
planes
```

`weather` gives the weather at each NYC airport for each hour:
```{r}
weather
```

For nycflights13:
<ul>
<li> `flights` connects to `planes` via a single variable, `tailnum`. </li> 
<li> `flights` connects to `airlines` through the `carrier`  </li>
<li> `flights` connects to `airports` in 2 ways: via  `origin` + `dest`  </li> 
<li> `flights` connects to `weather` via `origin` (the location), + `year`, `month`, `day` and `hour` (the time). </li> 
</ul>

13.2.1 Exercises

Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?

weather only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

# 13.3 Keys

Variables used to connect each pair of tables = **keys** = a variable/set of variables that uniquely ID's an observation. 

In simple cases, a single variable is sufficient to ID an observation (each plane is uniquely IDed by its `tailnum`). In other cases, multiple variables may be needed (To ID an observation in `weather` you need 5 variables: `year`, `month`, `day`, `hour`, and `origin`.

2 types of keys:

<ul>
<li> **primary key** uniquely ID's an observation in its *own* table (Ex: `planes$tailnum`) </li>
<li> foreign key uniquely ID's an observation in **another* table. (Ex: `flights$tailnum` = FK b/c it appears in `flights` where it matches each flight to a unique plane)</li>
</ul>

A variable can be *both* a PK + a FK (`origin` = part of the `weather` PK + is a FK for `airport` table)

Once you’ve ID'ed PK's in your tables, it’s good practice to verify they do indeed *uniquely* ID each observation. 

1 way to do so = `count()` the PK's + look for entries where n > 1:

```{r}
# check that tailnum uniquely ID's a plane
planes %>% 
  count(tailnum) %>% 
  filter(n > 1)

# check each YMD:H + origin combo uniquely ID's each weather record
weather %>% 
  count(year, month, day, hour, origin) %>% 
  filter(n > 1)
```

Sometimes a table doesn’t have an *explicit* PK = each row is an observation, but no combo of variables reliably ID's it

What’s the PK in `flights`? Might think it would be the date + flight or tail #, but neither of those are unique:

```{r}
# check if YMD + flight # uniquely ID's a flight
flights %>% 
  count(year, month, day, flight) %>% 
  filter(n > 1)

# check if YMD + tail # uniquely ID's a flight
flights %>% 
  count(year, month, day, tailnum) %>% 
  filter(n > 1)
```

When starting to work w/ this data, I had naively assumed each flight # would be only used once per day (would make it much easier to communicate problems w/ a specific flight). Unfortunately that is not the case! 

If a table lacks a PK, it’s sometimes useful to *add one* w/ `mutate()` and `row_number()` = makes it easier to match observations if you’ve done some filtering + want to check back in w/ original data = a **surrogate key**.

A PK + the corresponding FK in another table form a **relation**. (typically **one-to-many** --> Ex: Each flight has 1 plane, but each plane has many flights)

In other data, you’ll occasionally see a **1-to-1** relationship (special case of 1-to-many). You can model **many-to-many** relations w/ a **many-to-1** relation + a **1-to-many relation**. 

For example, in this data there’s a **many-to-many** relationship between airlines + airports: each airline flies to many airports; each airport hosts many airlines.

## 13.3.1 Exercises

Add a surrogate key to flights.

Identify the keys in the following datasets

Lahman::Batting,
babynames::babynames
nasaweather::atmos
fueleconomy::vehicles
ggplot2::diamonds
(You might need to install some packages and read some documentation.)

Draw a diagram illustrating the connections between the Batting, Master, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between Master, Managers, AwardsManagers.

How would you characterise the relationship between the Batting, Pitching, and Fielding tables?

## 13.4 Mutating joins

The first tool we’ll look at for combining a pair of tables is the mutating join. A mutating join allows you to combine variables from two tables. It first matches observations by their keys, then copies across variables from one table to the other.

Like mutate(), the join functions add variables to the right, so if you have a lot of variables already, the new variables won’t get printed out. For these examples, we’ll make it easier to see what’s going on in the examples by creating a narrower dataset:

flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
flights2
#> # A tibble: 336,776 × 8
#>    year month   day  hour origin  dest tailnum carrier
#>   <int> <int> <int> <dbl>  <chr> <chr>   <chr>   <chr>
#> 1  2013     1     1     5    EWR   IAH  N14228      UA
#> 2  2013     1     1     5    LGA   IAH  N24211      UA
#> 3  2013     1     1     5    JFK   MIA  N619AA      AA
#> 4  2013     1     1     5    JFK   BQN  N804JB      B6
#> 5  2013     1     1     6    LGA   ATL  N668DN      DL
#> 6  2013     1     1     5    EWR   ORD  N39463      UA
#> # ... with 3.368e+05 more rows
(Remember, when you’re in RStudio, you can also use View() to avoid this problem.)

Imagine you want to add the full airline name to the flights2 data. You can combine the airlines and flights2 data frames with left_join():

flights2 %>%
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")
#> # A tibble: 336,776 × 7
#>    year month   day  hour tailnum carrier                   name
#>   <int> <int> <int> <dbl>   <chr>   <chr>                  <chr>
#> 1  2013     1     1     5  N14228      UA  United Air Lines Inc.
#> 2  2013     1     1     5  N24211      UA  United Air Lines Inc.
#> 3  2013     1     1     5  N619AA      AA American Airlines Inc.
#> 4  2013     1     1     5  N804JB      B6        JetBlue Airways
#> 5  2013     1     1     6  N668DN      DL   Delta Air Lines Inc.
#> 6  2013     1     1     5  N39463      UA  United Air Lines Inc.
#> # ... with 3.368e+05 more rows
The result of joining airlines to flights2 is an additional variable: name. This is why I call this type of join a mutating join. In this case, you could have got to the same place using mutate() and R’s base subsetting:

flights2 %>%
  select(-origin, -dest) %>% 
  mutate(name = airlines$name[match(carrier, airlines$carrier)])
#> # A tibble: 336,776 × 7
#>    year month   day  hour tailnum carrier                   name
#>   <int> <int> <int> <dbl>   <chr>   <chr>                  <chr>
#> 1  2013     1     1     5  N14228      UA  United Air Lines Inc.
#> 2  2013     1     1     5  N24211      UA  United Air Lines Inc.
#> 3  2013     1     1     5  N619AA      AA American Airlines Inc.
#> 4  2013     1     1     5  N804JB      B6        JetBlue Airways
#> 5  2013     1     1     6  N668DN      DL   Delta Air Lines Inc.
#> 6  2013     1     1     5  N39463      UA  United Air Lines Inc.
#> # ... with 3.368e+05 more rows
But this is hard to generalise when you need to match multiple variables, and takes close reading to figure out the overall intent.

The following sections explain, in detail, how mutating joins work. You’ll start by learning a useful visual representation of joins. We’ll then use that to explain the four mutating join functions: the inner join, and the three outer joins. When working with real data, keys don’t always uniquely identify observations, so next we’ll talk about what happens when there isn’t a unique match. Finally, you’ll learn how to tell dplyr which variables are the keys for a given join.

13.4.1 Understanding joins

To help you learn how joins work, I’m going to use a visual representation:



x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     3, "x3"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2",
     4, "y3"
)
The coloured column represents the “key” variable: these are used to match the rows between the tables. The grey column represents the “value” column that is carried along for the ride. In these examples I’ll show a single key variable, but the idea generalises in a straightforward way to multiple keys and multiple values.

A join is a way of connecting each row in x to zero, one, or more rows in y. The following diagram shows each potential match as an intersection of a pair of lines.



(If you look closely, you might notice that we’ve switched the order of the key and value columns in x. This is to emphasise that joins match based on the key; the value is just carried along for the ride.)

In an actual join, matches will be indicated with dots. The number of dots = the number of matches = the number of rows in the output.



13.4.2 Inner join

The simplest type of join is the inner join. An inner join matches pairs of observations whenever their keys are equal:



(To be precise, this is an inner equijoin because the keys are matched using the equality operator. Since most joins are equijoins we usually drop that specification.)

The output of an inner join is a new data frame that contains the key, the x values, and the y values. We use by to tell dplyr which variable is the key:

x %>% 
  inner_join(y, by = "key")
#> # A tibble: 2 × 3
#>     key val_x val_y
#>   <dbl> <chr> <chr>
#> 1     1    x1    y1
#> 2     2    x2    y2
The most important property of an inner join is that unmatched rows are not included in the result. This means that generally inner joins are usually not appropriate for use in analysis because it’s too easy to lose observations.

13.4.3 Outer joins

An inner join keeps observations that appear in both tables. An outer join keeps observations that appear in at least one of the tables. There are three types of outer joins:

A left join keeps all observations in x.
A right join keeps all observations in y.
A full join keeps all observations in x and y.
These joins work by adding an additional “virtual” observation to each table. This observation has a key that always matches (if no other key matches), and a value filled with NA.

Graphically, that looks like:



The most commonly used join is the left join: you use this whenever you look up additional data from another table, because it preserves the original observations even when there isn’t a match. The left join should be your default join: use it unless you have a strong reason to prefer one of the others.

Another way to depict the different types of joins is with a Venn diagram:



However, this is not a great representation. It might jog your memory about which join preserves the observations in which table, but it suffers from a major limitation: a Venn diagram can’t show what happens when keys don’t uniquely identify an observation.

13.4.4 Duplicate keys

So far all the diagrams have assumed that the keys are unique. But that’s not always the case. This section explains what happens when the keys are not unique. There are two possibilities:

One table has duplicate keys. This is useful when you want to add in additional information as there is typically a one-to-many relationship.



Note that I’ve put the key column in a slightly different position in the output. This reflects that the key is a primary key in y and a foreign key in x.

x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     2, "x3",
     1, "x4"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2"
)
left_join(x, y, by = "key")
#> # A tibble: 4 × 3
#>     key val_x val_y
#>   <dbl> <chr> <chr>
#> 1     1    x1    y1
#> 2     2    x2    y2
#> 3     2    x3    y2
#> 4     1    x4    y1
Both tables have duplicate keys. This is usually an error because in neither table do the keys uniquely identify an observation. When you join duplicated keys, you get all possible combinations, the Cartesian product:



x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     2, "x3",
     3, "x4"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2",
     2, "y3",
     3, "y4"
)
left_join(x, y, by = "key")
#> # A tibble: 6 × 3
#>     key val_x val_y
#>   <dbl> <chr> <chr>
#> 1     1    x1    y1
#> 2     2    x2    y2
#> 3     2    x2    y3
#> 4     2    x3    y2
#> 5     2    x3    y3
#> 6     3    x4    y4
13.4.5 Defining the key columns

So far, the pairs of tables have always been joined by a single variable, and that variable has the same name in both tables. That constraint was encoded by by = "key". You can use other values for by to connect the tables in other ways:

The default, by = NULL, uses all variables that appear in both tables, the so called natural join. For example, the flights and weather tables match on their common variables: year, month, day, hour and origin.

flights2 %>% 
  left_join(weather)
#> Joining, by = c("year", "month", "day", "hour", "origin")
#> # A tibble: 336,776 × 18
#>    year month   day  hour origin  dest tailnum carrier  temp  dewp humid
#>   <dbl> <dbl> <int> <dbl>  <chr> <chr>   <chr>   <chr> <dbl> <dbl> <dbl>
#> 1  2013     1     1     5    EWR   IAH  N14228      UA    NA    NA    NA
#> 2  2013     1     1     5    LGA   IAH  N24211      UA    NA    NA    NA
#> 3  2013     1     1     5    JFK   MIA  N619AA      AA    NA    NA    NA
#> 4  2013     1     1     5    JFK   BQN  N804JB      B6    NA    NA    NA
#> 5  2013     1     1     6    LGA   ATL  N668DN      DL  39.9  26.1  57.3
#> 6  2013     1     1     5    EWR   ORD  N39463      UA    NA    NA    NA
#> # ... with 3.368e+05 more rows, and 7 more variables: wind_dir <dbl>,
#> #   wind_speed <dbl>, wind_gust <dbl>, precip <dbl>, pressure <dbl>,
#> #   visib <dbl>, time_hour <dttm>
A character vector, by = "x". This is like a natural join, but uses only some of the common variables. For example, flights and planes have year variables, but they mean different things so we only want to join by tailnum.

flights2 %>% 
  left_join(planes, by = "tailnum")
#> # A tibble: 336,776 × 16
#>   year.x month   day  hour origin  dest tailnum carrier year.y
#>    <int> <int> <int> <dbl>  <chr> <chr>   <chr>   <chr>  <int>
#> 1   2013     1     1     5    EWR   IAH  N14228      UA   1999
#> 2   2013     1     1     5    LGA   IAH  N24211      UA   1998
#> 3   2013     1     1     5    JFK   MIA  N619AA      AA   1990
#> 4   2013     1     1     5    JFK   BQN  N804JB      B6   2012
#> 5   2013     1     1     6    LGA   ATL  N668DN      DL   1991
#> 6   2013     1     1     5    EWR   ORD  N39463      UA   2012
#> # ... with 3.368e+05 more rows, and 7 more variables: type <chr>,
#> #   manufacturer <chr>, model <chr>, engines <int>, seats <int>,
#> #   speed <int>, engine <chr>
Note that the year variables (which appear in both input data frames, but are not constrained to be equal) are disambiguated in the output with a suffix.

A named character vector: by = c("a" = "b"). This will match variable a in table x to variable b in table y. The variables from x will be used in the output.

For example, if we want to draw a map we need to combine the flights data with the airports data which contains the location (lat and long) of each airport. Each flight has an origin and destination airport, so we need to specify which one we want to join to:

flights2 %>% 
  left_join(airports, c("dest" = "faa"))
#> # A tibble: 336,776 × 15
#>    year month   day  hour origin  dest tailnum carrier
#>   <int> <int> <int> <dbl>  <chr> <chr>   <chr>   <chr>
#> 1  2013     1     1     5    EWR   IAH  N14228      UA
#> 2  2013     1     1     5    LGA   IAH  N24211      UA
#> 3  2013     1     1     5    JFK   MIA  N619AA      AA
#> 4  2013     1     1     5    JFK   BQN  N804JB      B6
#> 5  2013     1     1     6    LGA   ATL  N668DN      DL
#> 6  2013     1     1     5    EWR   ORD  N39463      UA
#> # ... with 3.368e+05 more rows, and 7 more variables: name <chr>,
#> #   lat <dbl>, lon <dbl>, alt <int>, tz <dbl>, dst <chr>, tzone <chr>

flights2 %>% 
  left_join(airports, c("origin" = "faa"))
#> # A tibble: 336,776 × 15
#>    year month   day  hour origin  dest tailnum carrier                name
#>   <int> <int> <int> <dbl>  <chr> <chr>   <chr>   <chr>               <chr>
#> 1  2013     1     1     5    EWR   IAH  N14228      UA Newark Liberty Intl
#> 2  2013     1     1     5    LGA   IAH  N24211      UA          La Guardia
#> 3  2013     1     1     5    JFK   MIA  N619AA      AA John F Kennedy Intl
#> 4  2013     1     1     5    JFK   BQN  N804JB      B6 John F Kennedy Intl
#> 5  2013     1     1     6    LGA   ATL  N668DN      DL          La Guardia
#> 6  2013     1     1     5    EWR   ORD  N39463      UA Newark Liberty Intl
#> # ... with 3.368e+05 more rows, and 6 more variables: lat <dbl>,
#> #   lon <dbl>, alt <int>, tz <dbl>, dst <chr>, tzone <chr>
13.4.6 Exercises

Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
(Don’t worry if you don’t understand what semi_join() does — you’ll learn about it next.)

You might want to use the size or colour of the points to display the average delay for each airport.

Add the location of the origin and destination (i.e. the lat and lon) to flights.

Is there a relationship between the age of a plane and its delays?

What weather conditions make it more likely to see a delay?

What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.

13.4.7 Other implementations

base::merge() can perform all four types of mutating join:

dplyr	merge
inner_join(x, y)	merge(x, y)
left_join(x, y)	merge(x, y, all.x = TRUE)
right_join(x, y)	merge(x, y, all.y = TRUE),
full_join(x, y)	merge(x, y, all.x = TRUE, all.y = TRUE)
The advantages of the specific dplyr verbs is that they more clearly convey the intent of your code: the difference between the joins is really important but concealed in the arguments of merge(). dplyr’s joins are considerably faster and don’t mess with the order of the rows.

SQL is the inspiration for dplyr’s conventions, so the translation is straightforward:

dplyr	SQL
inner_join(x, y, by = "z")	SELECT * FROM x INNER JOIN y USING (z)
left_join(x, y, by = "z")	SELECT * FROM x LEFT OUTER JOIN y USING (z)
right_join(x, y, by = "z")	SELECT * FROM x RIGHT OUTER JOIN y USING (z)
full_join(x, y, by = "z")	SELECT * FROM x FULL OUTER JOIN y USING (z)
Note that “INNER” and “OUTER” are optional, and often omitted.

Joining different variables between the tables, e.g. inner_join(x, y, by = c("a" = "b")) uses a slightly different syntax in SQL: SELECT * FROM x INNER JOIN y ON x.a = y.b. As this syntax suggests, SQL supports a wider range of join types than dplyr because you can connect the tables using constraints other than equality (sometimes called non-equijoins).

13.5 Filtering joins

Filtering joins match observations in the same way as mutating joins, but affect the observations, not the variables. There are two types:

semi_join(x, y) keeps all observations in x that have a match in y.
anti_join(x, y) drops all observations in x that have a match in y.
Semi-joins are useful for matching filtered summary tables back to the original rows. For example, imagine you’ve found the top ten most popular destinations:

top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest
#> # A tibble: 10 × 2
#>    dest     n
#>   <chr> <int>
#> 1   ORD 17283
#> 2   ATL 17215
#> 3   LAX 16174
#> 4   BOS 15508
#> 5   MCO 14082
#> 6   CLT 14064
#> # ... with 4 more rows
Now you want to find each flight that went to one of those destinations. You could construct a filter yourself:

flights %>% 
  filter(dest %in% top_dest$dest)
#> # A tibble: 141,145 × 19
#>    year month   day dep_time sched_dep_time dep_delay arr_time
#>   <int> <int> <int>    <int>          <int>     <dbl>    <int>
#> 1  2013     1     1      542            540         2      923
#> 2  2013     1     1      554            600        -6      812
#> 3  2013     1     1      554            558        -4      740
#> 4  2013     1     1      555            600        -5      913
#> 5  2013     1     1      557            600        -3      838
#> 6  2013     1     1      558            600        -2      753
#> # ... with 1.411e+05 more rows, and 12 more variables:
#> #   sched_arr_time <int>, arr_delay <dbl>, carrier <chr>, flight <int>,
#> #   tailnum <chr>, origin <chr>, dest <chr>, air_time <dbl>,
#> #   distance <dbl>, hour <dbl>, minute <dbl>, time_hour <dttm>
But it’s difficult to extend that approach to multiple variables. For example, imagine that you’d found the 10 days with highest average delays. How would you construct the filter statement that used year, month, and day to match it back to flights?

Instead you can use a semi-join, which connects the two tables like a mutating join, but instead of adding new columns, only keeps the rows in x that have a match in y:

flights %>% 
  semi_join(top_dest)
#> Joining, by = "dest"
#> # A tibble: 141,145 × 19
#>    year month   day dep_time sched_dep_time dep_delay arr_time
#>   <int> <int> <int>    <int>          <int>     <dbl>    <int>
#> 1  2013     1     1      554            558        -4      740
#> 2  2013     1     1      558            600        -2      753
#> 3  2013     1     1      608            600         8      807
#> 4  2013     1     1      629            630        -1      824
#> 5  2013     1     1      656            700        -4      854
#> 6  2013     1     1      709            700         9      852
#> # ... with 1.411e+05 more rows, and 12 more variables:
#> #   sched_arr_time <int>, arr_delay <dbl>, carrier <chr>, flight <int>,
#> #   tailnum <chr>, origin <chr>, dest <chr>, air_time <dbl>,
#> #   distance <dbl>, hour <dbl>, minute <dbl>, time_hour <dttm>
Graphically, a semi-join looks like this:



Only the existence of a match is important; it doesn’t matter which observation is matched. This means that filtering joins never duplicate rows like mutating joins do:



The inverse of a semi-join is an anti-join. An anti-join keeps the rows that don’t have a match:



Anti-joins are useful for diagnosing join mismatches. For example, when connecting flights and planes, you might be interested to know that there are many flights that don’t have a match in planes:

flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(tailnum, sort = TRUE)
#> # A tibble: 722 × 2
#>   tailnum     n
#>     <chr> <int>
#> 1    <NA>  2512
#> 2  N725MQ   575
#> 3  N722MQ   513
#> 4  N723MQ   507
#> 5  N713MQ   483
#> 6  N735MQ   396
#> # ... with 716 more rows
13.5.1 Exercises

What does it mean for a flight to have a missing tailnum? What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)

Filter flights to only show flights with planes that have flown at least 100 flights.

Combine fueleconomy::vehicles and fueleconomy::common to find only the records for the most common models.

Find the 48 hours (over the course of the whole year) that have the worst delays. Cross-reference it with the weather data. Can you see any patterns?

What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?

You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Confirm or reject this hypothesis using the tools you’ve learned above.

13.6 Join problems

The data you’ve been working with in this chapter has been cleaned up so that you’ll have as few problems as possible. Your own data is unlikely to be so nice, so there are a few things that you should do with your own data to make your joins go smoothly.

Start by identifying the variables that form the primary key in each table. You should usually do this based on your understanding of the data, not empirically by looking for a combination of variables that give a unique identifier. If you just look for variables without thinking about what they mean, you might get (un)lucky and find a combination that’s unique in your current data but the relationship might not be true in general.

For example, the altitude and longitude uniquely identify each airport, but they are not good identifiers!

airports %>% count(alt, lon) %>% filter(n > 1)
#> Source: local data frame [0 x 3]
#> Groups: alt [0]
#> 
#> # ... with 3 variables: alt <int>, lon <dbl>, n <int>
Check that none of the variables in the primary key are missing. If a value is missing then it can’t identify an observation!

Check that your foreign keys match primary keys in another table. The best way to do this is with an anti_join(). It’s common for keys not to match because of data entry errors. Fixing these is often a lot of work.

If you do have missing keys, you’ll need to be thoughtful about your use of inner vs. outer joins, carefully considering whether or not you want to drop rows that don’t have a match.

Be aware that simply checking the number of rows before and after the join is not sufficient to ensure that your join has gone smoothly. If you have an inner join with duplicate keys in both tables, you might get unlucky as the number of dropped rows might exactly equal the number of duplicated rows!

13.7 Set operations

The final type of two-table verb are the set operations. Generally, I use these the least frequently, but they are occasionally useful when you want to break a single complex filter into simpler pieces. All these operations work with a complete row, comparing the values of every variable. These expect the x and y inputs to have the same variables, and treat the observations like sets:

intersect(x, y): return only observations in both x and y.
union(x, y): return unique observations in x and y.
setdiff(x, y): return observations in x, but not in y.
Given this simple data:

df1 <- tribble(
  ~x, ~y,
   1,  1,
   2,  1
)
df2 <- tribble(
  ~x, ~y,
   1,  1,
   1,  2
)
The four possibilities are:

intersect(df1, df2)
#> # A tibble: 1 × 2
#>       x     y
#>   <dbl> <dbl>
#> 1     1     1

# Note that we get 3 rows, not 4
union(df1, df2)
#> # A tibble: 3 × 2
#>       x     y
#>   <dbl> <dbl>
#> 1     1     2
#> 2     2     1
#> 3     1     1

setdiff(df1, df2)
#> # A tibble: 1 × 2
#>       x     y
#>   <dbl> <dbl>
#> 1     2     1

setdiff(df2, df1)
#> # A tibble: 1 × 2
#>       x     y
#>   <dbl> <dbl>
#> 1     1     2
