---
title: "Inference for Numerical Data"
author: "Steve Newns"
date: "October 18, 2017"
output: html_document
---

## Getting Started
  
### Load packages

```{r load-packages, message=F, warning=F}
library(statsr)
library(tidyverse)
library(ggplot2)
```

### The data

In 2004, the state of NC released a large data set containing info on births recorded in the state. This data set is useful to researchers studying the relation between habits + practices of expectant mothers + the birth of their children. We will work w/ a random sample of observations from this data set.

```{r load-data}
data(nc)
```

We have observations on 13 different variables, some categorical + some numerical. The meaning of each variable is as follows:

variable         | description
---------------- | ---------------------------------------------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (**premie**) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

1. There are 1,000 cases in this data set, what do the cases represent? 
<ol>
<li> The hospitals where the births took place </li> 
<li> The fathers of the children </li> 
<li> The days of the births </li>
<li> **The births** </li>
</ol>

As a 1st step in the analysis, take a look at the variables in the dataset. 

```{r str}
str(nc)
```

As you review the variable summaries, consider which variables are categorical + which are numerical. For numerical variables, are there outliers? If you aren't sure or want to take a closer look at the data, make a graph.

## EDA

We will first start w/ analyzing weight gained by mothers throughout the pregnancy: `gained`.

Using viz + summary stats, describe the distribution of weight gained by mothers during pregnancy.

```{r summary, warning=F,message=F}
summary(nc$gained)

nc %>%
  ggplot(aes(gained)) + geom_histogram()
```

2. How many mothers are we missing weight gain data from?
  <ol>
  <li> 0 </li>
  <li> 13 </li>
  <li> **27** </li>
  <li> 31 </li>
  </ol>
  
Next, consider the possible relationship between a mother's smoking habit + weight of her baby. Plotting the data is a useful 1st step b/c it helps us quickly visualize trends, ID strong associations, + develop research questions.

Make side-by-side boxplots of `habit` and `weight`. 

```{r}
nc %>%
  filter(!is.na(habit)) %>%
  ggplot(aes(habit,weight)) + geom_boxplot()
```

3. Which of the following is false about the relationship between habit and weight?
<ol>
<li> Median birth weight of babies born to non-smoker mothers is slightly higher than that of babies born to smoker mothers. </li>
<li> Range of birth weights of babies born to non-smoker mothers is greater than that of babies born to smoker mothers. </li>
<li> **Both distributions are extremely right skewed**. </li>
  <ul> ***Only nonsmoker is </ul>
<li> The IQRs of the distributions are roughly equal. </li>
</ol>

The box plots show how the medians of the 2 distributions compare, but we can also compare means of the distributions by grouping data by `habit` + then calculate the mean `weight` in each group.

```{r by-means}
nc %>%
  filter(!is.na(habit)) %>%
  group_by(habit) %>%
  summarise(mean_weight = mean(weight))
```

There is an observed difference, but *is this difference statistically significant?* 

## Inference - Hypothesis Test

<div id="exercise">
**Exercise**: Are all conditions necessary for inference satisfied? Comment on each. You can compute the group sizes using the same `by` command above but replacing `mean(weight)` with `n()`.
</div>

4.  What are the hypotheses for testing if the average weights of babies born to 
smoking and non-smoking mothers are different?
<ul> **$H_0: \mu_{smoking} = \mu_{non-smoking}$; $H_A: \mu_{smoking} \ne \mu_{non-smoking}$** </ul>

We'll use `inference()`, for conducting hypothesis tests + constructing CI's 

```{r inf-weight-habit-ht, tidy=FALSE}
inference(y = weight, x = habit, data = nc, statistic = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

<ul> `y` = response variable we're interested in </ul>
<ul> `x` = explanatory variable (splits the data into 2 groups, smokers + non-smokers </ul> 
<ul> `statistic` = the sample statistic we're using/population parameter we're estimating (can also work w/ "median" + "proportion" </ul>
<ul> `type` of inference we want: a hypothesis test (`"ht"`) or a CI (`"ci"`) + when performing `ht`, we also need to supply the `null` value, which in this case is `0`, since the null sets the 2 population means equal to each other </ul> 
<ul> `alternative` hypothesis can be `"less"`, `"greater"`, or `"twosided"` </ul> 
<ul> `method` of inference can be `"theoretical"` or `"simulation"` based </ul>

For more info on `inference` see the help file w/ `?inference`.

<div id="exercise">
  **Exercise**: What is the conclusion of the hypothesis test?
  </div>

```{r inf-weight-habit-ci, tidy=FALSE}
inference(y = weight, x = habit, data = nc, statistic = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```  
  
  5.  Change `type` argument to `"ci"` to construct + record a CI for the difference between weights of babies born to nonsmoking vs. smoking mothers, + interpret this interval in context of the data. Note that by 
default you'll get a 95% CI. To change the
confidence level, add a new argument (`conf_level`) which takes on a value
between 0-1. Also note that when doing a CI, arguments like
`null` and `alternative` are not useful, so make sure to remove them.
<ol>
<li> We are 95% confident that babies born to nonsmoker mothers are on average 0.05 to 0.58 pounds lighter at birth than babies born to smoker mothers. </li>
<li> We are 95% confident that the difference in average weights of babies whose moms are smokers and nonsmokers is between 0.05 to 0.58 pounds. </li>
<li> We are 95% confident that the difference in average weights of babies in this sample whose moms are smokers and nonsmokers is between 0.05 to 0.58 pounds. </li>
<li> **We are 95% confident that babies born to nonsmoker mothers are on average 0.05 to 0.58 pounds heavier at birth than babies born to smoker mothers**. </li>
</ol>

By default the function reports an interval for ($\mu_{nonsmoker} - \mu_{smoker}$). We can easily change this order by using the `order` argument:

```{r inf-weight-habit-ci2, tidy=FALSE}
inference(y = weight, x = habit, data = nc, statistic = "mean", type = "ci", 
method = "theoretical", order = c("smoker","nonsmoker"))
```

6. Calculate a 99% CI for average length of pregnancies 
(`weeks`). Note since we're doing inference on a single population 
parameter, there is no explanatory variable, so you can omit the `x` variable 
from the function. Which of the following is the correct interpretation of this 
interval?

```{r inf-weight-habit-ci3, tidy=FALSE}
inference(y = weeks, data = nc, statistic = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", conf_level = .99)
``` 

  <ol>
  <li>(38.1526 , 38.5168)</li>
  <li>(38.0892 , 38.5661)</li>
  <li>(6.9779 , 7.2241)</li>
  <li>**(38.0952 , 38.5742)**</li>
  </ol>

<div id="exercise">
  **Exercise**: Calculate a new CI for the same parameter at the 90% confidence level. Comment on the width of this interval versus the one obtained in the the previous exercise.
</div>
```{r inf-weight-habit-ci4, tidy=FALSE}
inference(y = weeks, data = nc, statistic = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", conf_level = .90)
``` 

<div id="exercise">
  **Exercise**: Conduct a hypothesis test evaluating whether the average weight gained by younger mothers is different than the average weight gained by mature mothers.
</div>
```{r inf-weight-habit-ht2, tidy=FALSE}
inference(y = gained, x = mature, data = nc, statistic = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

**The difference is *not* significant (`p > 0.1`), so the average weight gained by younger mothers is not significantly different than the average weight gained by mature mothers**.

7. Now, a non-inference task: Determine the age cutoff for younger and mature mothers. Use a method of your choice, and explain how your method works.
```{r cutoff-mature}
nc %>%
  group_by(mature) %>%
  summarise(min_age = min(mage),
            max_age = max(mage))
```

Looking at the minimum age for a "mature" mother and the maximum age for a "younger" mother, we can see that the age cutoff between the 2 factor levels is the age of 34. Any mother older than 34 years of age would be considered mature, and any mother younger than (*and including*) the age of 34 would be considered younger.

<div id="exercise">
  **Exercise**: Pick a pair of variables: 1 numerical (response) and 1 categorical (explanatory). Come up w/ a research question evaluating the relationship between these variables. Formulate the question in a way that it can be answered using a hypothesis test and/or a CI. Answer your question using the `inference` function, report the statistical results, and also provide an explanation in 
plain language. Be sure to check all assumptions,state your $\alpha$ level, and conclude 
in context. (Note: Picking your own variables, coming up with a research question, and analyzing the data to answer this question is basically what you'll need to do for your project as well.)

#### Is the a significant difference between weight gained by mothers who gave birth to "underweight" babies and mothers who gave birth to "normal"-weight babies?

1st, we visualize the distributions of low birth weight babies and non-low birth weight babies with some side-by-side box plots.

</div>
```{r}
nc %>%
  filter(!is.na(lowbirthweight)) %>%
  filter(!is.na(gained)) %>%
  ggplot(aes(lowbirthweight,gained)) + geom_boxplot()
```

It looks like weight gained by mothers who gave birth to low-birth weight babies are bit more varied, but no too much. Also, the distribution for weight gained by mothers who gave birth to not-low-birth weight babies is not too skewed, but has some outliers, while the distribution of mothers who gave birth to low-birth weight babies looks a little right-skewed, but with no outliers.

Next, we can note that the median weight gained by mothers who gave birth to not-low-birth weight babies is a bit higher, which is something we can say we'd expect, as not-low birth weight babies would, of course, weigh more than low birth weight babies.

But is this difference significantly different?

```{r}
inference(y = gained, x = lowbirthweight, data = nc, statistic = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

We can see that our result is indeed significant (`p < 0.01`), so we can construct a confidence interval.

```{r}
inference(y = gained, x = lowbirthweight, data = nc, statistic = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", conf_level = .95, order = c("nowlow","low"))
```

**We are 95% confident that weight gained by mothers who gave birth to not-low birth weight is on average 1.80 to 7.72 pounds heavier at birth than weight gained by mothers who gave birth to low birth weight babies**

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.
</div>