---
title: "01b_USDA"
output: html_document
---

This dataset contains all foods in USDA database in 100g amounts.
```{r}
usda <- read.csv("data/USDA.csv")
str(usda)
```

So, we have 7058 records/foods with their unique ID's, some descriptions, and some info about calories, macros, sodium, vitamins (in IU's), etc.

Find the food with the highest sodium value.
```{r}
usda %>% filter(Sodium == max(Sodium, na.rm = T)) %>%
  select(Description)
```

Not suprisingly, the food with the most sodium per 100g is in fact Table salt, which has a max level of sodium of 38,758mg per 100g. Let's look at other foods wtih more than 10,000 mg of sodium
```{r}
(high_sodium <- usda %>% filter(Sodium > 10000) %>%
  select(Description))
```

Not too suprisingly, the most high sodium "foods" are soups, sauces, and leavening agents. But, caviar usually mentioned as a high-sodium food. Where is it?
```{r}
usda %>% filter(tolower(Description) == "caviar") %>%
  select(Description, Sodium)
```

There's only 1,500 mg in caviar, which seems small to 10,000 mg, or to 38,000 mg, but its not a fair comparison. We should compare this value for cavia to the mean and standard deviation of sodium across entire data set.
```{r}
usda %>%
  summarise(meanSodium = mean(Sodium, na.rm = T), stdDev = sd(Sodium, na.rm = T))
```


So, the mean sodium value is 322 mg with a standard deviation of 1045 mg, meaning that the highest value from these summary statistics is ~1400 mg total, which is less than caviar. Therefore, we can say caviar is pretty rich in sodium compared to all other foods in dataset.

Now create scatterplot of Protein vs. Fat
```{r}
plot(usda$Protein, usda$TotalFat, xlab = "Protein", ylab = "Total Fat",
     main = "Protein vs. Fat in USDA Food Database", col = "red")
```

It appears that foods that are high in protein are lower in fat, foods that are high in fat are low in protein. Now look at a histogram of Vitamin C.
```{r}
hist(usda$VitaminC, xlab = "Amount of Vitamin C (mg)", col = "blue")
```

Even though the max value is ~2000, almost all foods are actually below 250mg (about ~6000 of them). Limit the x-axis to get a more detailed view.
```{r}
hist(usda$VitaminC, xlim = c(0,100), xlab = "Amount of Vitamin C (mg)", col = "blue")
```

We only see 1 big cell, so we need to break this blob up into smaller pieces by creating divisions/breaks that are 1 mg in length
```{r}
hist(usda$VitaminC, xlim = c(0,100), xlab = "Amount of Vitamin C (mg)",
     breaks = 100, col = "blue") # break up x axis of 100 values into 1 mg bins
```

So, we expected 100 bins, but only got 5 bins, each of length 20 mg/ It turns out R took the *original* interval (0-2000) and divided *that* by 100, giving us 100 bins of size 20 mg. We want bins of length = 1 mg, so we need to break out our range into 2000 bins.
```{r}
hist(usda$VitaminC, xlim = c(0,100), xlab = "Amount of Vitamin C (mg)",
     breaks = 2000, col = "blue") # break up x axis of 100 values into 1 mg bins
```

Now we see that ~5000 foods have less than 1mg of Vitamin C in them

Now create boxplot of sugar
```{r}
boxplot(usda$Sugar, ylab = "Sugar Level (g)", main = "Boxplot of Sugar Levels")
```

The median value seems quite low, at around 5g, and we have many outliers, with multiple foods being around 100g sugar per 100g of food. Let's add a new binary variable that = 1 if sodium is above average and = 0 if below average.
```{r}
library(magrittr)
usda %<>% mutate(above_avg_sodium = if_else(Sodium > mean(Sodium, na.rm = T),1,0))
summary(usda$above_avg_sodium)
```
```{r}
usda %<>% mutate(above_avg_sodium = if_else(Sodium > mean(Sodium, na.rm = T),1,0))
summary(usda$above_avg_sodium)
```
Unfortunately, we've got 84 `NA`'s. Remove them and then make the same variables for protein, fat, and carbs.
```{r}
usda %<>% mutate(highSodium = if_else(Sodium > mean(Sodium, na.rm = T),1,0),
              highProtein = if_else(Protein > mean(Protein, na.rm = T),1,0),
              highFat = if_else(TotalFat > mean(TotalFat, na.rm = T),1,0),
              highCarbs = if_else(Carbohydrate > mean(Carbohydrate, na.rm = T),1,0))
summary(usda[,c("highSodium","highProtein","highFat","highCarbs")])
```

Now, lets look for relationships. 1st, count how many foods have an above average sodium level.
```{r}
table(usda$highSodium)
```
There are 2090 foods that have an above average sodium level. See how many have above average sodium *and* fat.
```{r}
table(usda$highSodium,usda$highFat)
```

There are 712 foods that are very bad for you, apparantly. Let's compute average iron and then sort by high and low protein.
```{r}
tapply(usda$Iron, usda$highProtein, mean, na.rm = T)
```

Looks like the average value of iron for high protein foods is 3.2mg, and for low protein it's 2.6mg. Now, do the same for max Vitamin C for high and low carb foods.
```{r}
tapply(usda$VitaminC, usda$highCarbs, max, na.rm = T)
```

Looks like the max vitamin C value for low carb foods is 1678 mg, and for high carb foods it's 2400 mg.

So, are high carb foods high in vitamin C?
```{r}
tapply(usda$VitaminC, usda$highCarbs, summary, na.rm = TRUE)
```

We see that low carb foods have a lower average and median vitamin C values, so it seems like it's a *general* trend that lower carb foods have lower values of Vitamin C.