---
title: "ISLR_Ch3_LinearRegression"
author: "Steve Newns"
date: "April 20, 2018"
output: html_document
---
# 3.6.2 Simple Linear Regression
```{r}
library(MASS)
library(ISLR)
```

`MASS` library contains `Boston` data set, which records `medv` (median house value) for 506 neighborhoods around Boston. We will **seek to predict `medv` using 13 predictors such as `rm` (avg #  of rooms per house), `age` (avg age of houses), + `lstat` (% of households with low socioeconomic status)**

```{r}
#fix(Boston)
names(Boston)
```

We will start by using `lm()` to fit a simple linear regression model, with `medv` as response + `lstat` as predictor. The basic
```{r}
simpleLM <- lm(medv ~ lstat, data = Boston)
```

Typing `simpleLM` gives some basic info about the model, but  `summary(simpleLM)` gives more detailed information like p-values + standard errors for the coefficients, as well as the R2 statistic and F-score
```{r}
simpleLM
summary(simpleLM)
```

Can use `names()` to find out what other pieces of info are stored in the fit model. 
```{r}
names(simpleLM)
```

Can extract quantities by name (`lm.fit$coefficients`) but it is safer to use **extractor functions** like `coef()` to access them.
```{r}
coef(simpleLM)
```

To obtain a CI for the coefficient estimates, use `confint()`
```{r}
confint(simpleLM)
```

`predict()` can be used to produce CI's *and* **prediction intervals** for the prediction of `medv` for a given value of `lstat`.

  * **CI's** tell you about how well you have determined the parameter of interest
    * Repeat an experiment many times + calculate a CI from each sample, we'd expect ~95% of those intervals to include the true value of the population parameter (have 95% confidence the interval contains the true parameter value)
    * Key point = CI's tell you about the likely location of the true population parameter.

  * **Prediction intervals** tell you *where you can expect to see the next DP sampled*. 
    * Assume the data really are randomly sampled from a Gaussian distribution. 
    * Collect a sample of data + calculate a PI
    * Then sample 1 more value from the population
    * Repeat many times + expect that next value to lie within such a prediction interval in 95% of samples.
    * Key point = A PI tells you *about the distribution of values*, NOT the uncertainty in determining the true population parameter
    * PI's must account for both the **uncertainty** in knowing the value of the true population parameter, plus **scatter/noise** So
    * **A PI is always wider than a confidence interval (less confident)**
```{r}
# create a column (vector) of 5,10,15 as new lstat values to use to predict medv's valie + CI with the model
predict(object = simpleLM, newdata = data.frame(lstat=c(5,10,15)), interval="confidence")
predict(object = simpleLM, newdata = data.frame(lstat=c(5,10,15)), interval="prediction")
```

95% CI associated with a `lstat` value = 10 is `(24.47, 25.63)`, and the 95 % **prediction interval** = `(12.828, 37.28)`. As expected, the *CI and PI's are centered around the same point* = a predicted value of `25.05` for `medv` when `lstat` = 10), but
the latter are substantially wider.

Now plot `medv` + `lstat` along w/ the **least squares regression line** using `plot()` and `abline()`
```{r}
plot(Boston$lstat,Boston$medv)
abline(simpleLM, col="red", lwd=3)
```

There is some evidence for **non-linearity** in this relationship.

`abline()` can be used to draw *any* line, not just the least squares regression line. To draw a line with intercept `a` and slope `b`, type `abline(a,b)`. 

```{r}
par(mfrow=c(2,3))
plot(Boston$lstat,Boston$medv)
abline(simpleLM, lwd=3)
plot(Boston$lstat,Boston$medv)
abline(simpleLM, lwd=3, col="red")
plot(Boston$lstat, Boston$medv, col="red")
plot(Boston$lstat, Boston$medv, pch=20) # change DP shape
plot(Boston$lstat, Boston$medv, pch="+") # change DP shape
plot(1:20,1:20,pch =1:20)
```
Next we examine some diagnostic plots, several of which were discussed
in Section 3.3.3. Four diagnostic plots are automatically produced by applying
the plot() function directly to the output from lm(). In general, this
command will produce one plot at a time, and hitting Enter will generate
the next plot. However, it is often convenient to view all four plots together.
We can achieve this by using the par() function, which tells R to split the par() display screen into separate panels so that multiple plots can be viewed simultaneously.
For example, par(mfrow=c(2,2)) divides the plotting region
into a 2 × 2 grid of panels.
> par
3.6 Lab: Linear Regression 113
> plot(predict (lm.fit), residuals (lm.fit))
> plot(predict (lm.fit), rstudent (lm.fit))
On the basis of the residual plots, there is some evidence of non-linearity.
Leverage statistics can be computed for any number of predictors using the
hatvalues() function. hatvalues()
> plot(hatvalues (lm.fit))
> which.max(hatvalues (lm.fit))
375
The which.max() function identifies the index of the largest element of a which.max() vector. In this case, it tells us which observation has the largest leverage
statistic.
3.6.3 Multiple Linear Regression
In order to fit a multiple linear regression model using least squares, we
again use the lm() function. The syntax lm(y???x1+x2+x3) is used to fit a
model with three predictors, x1, x2, and x3. The summary() function now
outputs the regression coefficients for all the predictors.
> lm.fit=lm(medv???lstat+age ,data=Boston )
> summary (lm.fit)
Call:
lm(formula = medv ??? lstat + age , data = Boston )
Residuals :
Min 1Q Median 3Q Max
-15.98 -3.98 -1.28 1.97 23.16
Coefficients:
Estimate Std. Error t value Pr(>|t|)
(Intercept ) 33.2228 0.7308 45.46 <2e-16 ***
lstat -1.0321 0.0482 -21.42 <2e-16 ***
age 0.0345 0.0122 2.83 0.0049 **
---
Signif. codes: 0 *** 0.001 ** 0.01 * 0.05 . 0.1 1
Residual standard error: 6.17 on 503 degrees of freedom
Multiple R-squared: 0.551, Adjusted R-squared: 0.549
F-statistic : 309 on 2 and 503 DF, p-value: <2e-16
The Boston data set contains 13 variables, and so it would be cumbersome
to have to type all of these in order to perform a regression using all of the
predictors. Instead, we can use the following short-hand:
> lm.fit=lm(medv???.,data=Boston)
> summary (lm.fit)
Call:
lm(formula = medv ??? ., data = Boston)
114 3. Linear Regression
Residuals :
Min 1Q Median 3Q Max
-15.594 -2.730 -0.518 1.777 26.199
Coefficients:
Estimate Std. Error t value Pr(>|t|)
(Intercept ) 3.646e+01 5.103e+00 7.144 3.28e-12 ***
crim -1.080e-01 3.286e-02 -3.287 0.001087 **
zn 4.642e-02 1.373e-02 3.382 0.000778 ***
indus 2.056e-02 6.150e-02 0.334 0.738288
chas 2.687e+00 8.616e-01 3.118 0.001925 **
nox -1.777e+01 3.820e+00 -4.651 4.25e-06 ***
rm 3.810e+00 4.179e-01 9.116 < 2e-16 ***
age 6.922e-04 1.321e-02 0.052 0.958229
dis -1.476e+00 1.995e-01 -7.398 6.01e-13 ***
rad 3.060e-01 6.635e-02 4.613 5.07e-06 ***
tax -1.233e-02 3.761e-03 -3.280 0.001112 **
ptratio -9.527e-01 1.308e-01 -7.283 1.31e-12 ***
black 9.312e-03 2.686e-03 3.467 0.000573 ***
lstat -5.248e-01 5.072e-02 -10.347 < 2e-16 ***
---
Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
Residual standard error: 4.745 on 492 degrees of freedom
Multiple R-Squared : 0.7406, Adjusted R-squared : 0.7338
F-statistic : 108.1 on 13 and 492 DF, p-value: < 2.2e-16
We can access the individual components of a summary object by name
(type ?summary.lm to see what is available). Hence summary(lm.fit)$r.sq
gives us the R2, and summary(lm.fit)$sigma gives us the RSE. The vif() vif() function, part of the car package, can be used to compute variance inflation
factors. Most VIF's are low to moderate for this data. The car package is
not part of the base R installation so it must be downloaded the first time
you use it via the install.packages option in R.
> library (car)
> vif(lm.fit)
crim zn indus chas nox rm age
1.79 2.30 3.99 1.07 4.39 1.93 3.10
dis rad tax ptratio black lstat
3.96 7.48 9.01 1.80 1.35 2.94
What if we would like to perform a regression using all of the variables but
one? For example, in the above regression output, age has a high p-value.
So we may wish to run a regression excluding this predictor. The following
syntax results in a regression using all predictors except age.
> lm.fit1=lm(medv???.-age ,data=Boston )
> summary (lm.fit1)
...
Alternatively, the update() function can be used. update()
3.6 Lab: Linear Regression 115
> lm.fit1=update(lm.fit , ???.-age)
3.6.4 Interaction Terms
It is easy to include interaction terms in a linear model using the lm() function.
The syntax lstat:black tells R to include an interaction term between
lstat and black. The syntax lstat*age simultaneously includes lstat, age,
and the interaction term lstat×age as predictors; it is a shorthand for
lstat+age+lstat:age.
> summary (lm(medv???lstat*age ,data=Boston))
Call:
lm(formula = medv ??? lstat * age , data = Boston )
Residuals :
Min 1Q Median 3Q Max
-15.81 -4.04 -1.33 2.08 27.55
Coefficients:
Estimate Std. Error t value Pr(>|t|)
(Intercept ) 36.088536 1.469835 24.55 < 2e-16 ***
lstat -1.392117 0.167456 -8.31 8.8e-16 ***
age -0.000721 0.019879 -0.04 0.971
lstat:age 0.004156 0.001852 2.24 0.025 *
---
Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
Residual standard error: 6.15 on 502 degrees of freedom
Multiple R-squared: 0.556, Adjusted R-squared: 0.553
F-statistic : 209 on 3 and 502 DF, p-value: <2e-16
3.6.5 Non-linear Transformations of the Predictors
The lm() function can also accommodate non-linear transformations of the
predictors. For instance, given a predictor X, we can create a predictor X2
using I(X^2). The function I() is needed since the ^ has a special meaning I() in a formula; wrapping as we do allows the standard usage in R, which is
to raise X to the power 2. We now perform a regression of medv onto lstat
and lstat2.
> lm.fit2=lm(medv???lstat+I(lstat^2))
> summary (lm.fit2)
Call:
lm(formula = medv ??? lstat + I(lstat ^2))
Residuals :
Min 1Q Median 3Q Max
-15.28 -3.83 -0.53 2.31 25.41
116 3. Linear Regression
Coefficients:
Estimate Std. Error t value Pr(>|t|)
(Intercept ) 42.86201 0.87208 49.1 <2e-16 ***
lstat -2.33282 0.12380 -18.8 <2e-16 ***
I(lstat ^2) 0.04355 0.00375 11.6 <2e-16 ***
---
Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
Residual standard error: 5.52 on 503 degrees of freedom
Multiple R-squared: 0.641, Adjusted R-squared: 0.639
F-statistic : 449 on 2 and 503 DF, p-value: <2e-16
The near-zero p-value associated with the quadratic term suggests that
it leads to an improved model. We use the anova() function to further anova() quantify the extent to which the quadratic fit is superior to the linear fit.
> lm.fit=lm(medv???lstat)
> anova(lm.fit ,lm.fit2)
Analysis of Variance Table
Model 1: medv ??? lstat
Model 2: medv ??? lstat + I(lstat^2)
Res.Df RSS Df Sum of Sq F Pr(>F)
1 504 19472
2 503 15347 1 4125 135 <2e-16 ***
---
Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
Here Model 1 represents the linear submodel containing only one predictor,
lstat, while Model 2 corresponds to the larger quadratic model that has two
predictors, lstat and lstat2. The anova() function performs a hypothesis
test comparing the two models. The null hypothesis is that the two models
fit the data equally well, and the alternative hypothesis is that the full
model is superior. Here the F-statistic is 135 and the associated p-value is
virtually zero. This provides very clear evidence that the model containing
the predictors lstat and lstat2 is far superior to the model that only
contains the predictor lstat. This is not surprising, since earlier we saw
evidence for non-linearity in the relationship between medv and lstat. If we
type
> par(mfrow=c(2,2))
> plot(lm.fit2)
then w
3.6 Lab: Linear Regression 117
> lm.fit5=lm(medv???poly(lstat ,5))
> summary (lm.fit5)
Call:
lm(formula = medv ??? poly(lstat , 5))
Residuals :
Min 1Q Median 3Q Max
-13.543 -3.104 -0.705 2.084 27.115
Coefficients:
Estimate Std. Error t value Pr(>|t|)
(Intercept ) 22.533 0.232 97.20 < 2e-16 ***
poly(lstat , 5)1 -152.460 5.215 -29.24 < 2e-16 ***
poly(lstat , 5)2 64.227 5.215 12.32 < 2e-16 ***
poly(lstat , 5)3 -27.051 5.215 -5.19 3.1e-07 ***
poly(lstat , 5)4 25.452 5.215 4.88 1.4e-06 ***
poly(lstat , 5)5 -19.252 5.215 -3.69 0.00025 ***
---
Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
Residual standard error: 5.21 on 500 degrees of freedom
Multiple R-squared: 0.682, Adjusted R-squared: 0.679
F-statistic : 214 on 5 and 500 DF, p-value: <2e-16
This suggests that including additional polynomial terms, up to fifth order,
leads to an improvement in the model fit! However, further investigation of
the data reveals that no polynomial terms beyond fifth order have signifi-
cant p-values in a regression fit.
Of course, we are in no way restricted to using polynomial transformations
of the predictors. Here we try a log transformation.
> summary (lm(medv???log(rm),data=Boston))
...
3.6.6 Qualitative Predictors
We will now examine the Carseats data, which is part of the ISLR library.
We will attempt to predict Sales (child car seat sales) in 400 locations
based on a number of predictors.
> fix(Carseats )
> names(Carseats )
[1] "Sales" "CompPrice " "Income" "Advertising "
[5] "Population " "Price" "ShelveLoc " "Age"
[9] "Education " "Urban" "US"
The Carseats data includes qualitative predictors such as Shelveloc, an indicator
of the quality of the shelving location-that is, the space within
a store in which the car seat is displayed-at each location. The predictor
Shelveloc takes on three possible values, Bad, Medium, and Good.
118 3. Linear Regression
Given a qualitative variable such as Shelveloc, R generates dummy variables
automatically. Below we fit a multiple regression model that includes some
interaction terms.
> lm.fit=lm(Sales???.+Income :Advertising +Price:Age ,data=Carseats )
> summary (lm.fit)
Call:
lm(formula = Sales ??? . + Income:Advertising + Price:Age , data =
Carseats )
Residuals :
Min 1Q Median 3Q Max
-2.921 -0.750 0.018 0.675 3.341
Coefficients:
Estimate Std. Error t value Pr(>|t|)
(Intercept ) 6.575565 1.008747 6.52 2.2e-10 ***
CompPrice 0.092937 0.004118 22.57 < 2e-16 ***
Income 0.010894 0.002604 4.18 3.6e-05 ***
Advertising 0.070246 0.022609 3.11 0.00203 **
Population 0.000159 0.000368 0.43 0.66533
Price -0.100806 0.007440 -13.55 < 2e-16 ***
ShelveLocGood 4.848676 0.152838 31.72 < 2e-16 ***
ShelveLocMedium 1.953262 0.125768 15.53 < 2e-16 ***
Age -0.057947 0.015951 -3.63 0.00032 ***
Education -0.020852 0.019613 -1.06 0.28836
UrbanYes 0.140160 0.112402 1.25 0.21317
USYes -0.157557 0.148923 -1.06 0.29073
Income:Advertising 0.000751 0.000278 2.70 0.00729 **
Price:Age 0.000107 0.000133 0.80 0.42381
---
Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
Residual standard error: 1.01 on 386 degrees of freedom
Multiple R-squared: 0.876, Adjusted R-squared: 0.872
F-statistic : 210 on 13 and 386 DF, p-value: <2e-16
The contrasts() function returns the coding that R uses for the dummy contrasts() variables.
> attach(Carseats )
> contrasts (ShelveLoc )
Good Medium
Bad 0 0
Good 1 0
Medium 0 1
Use ?contrasts to learn about other contrasts, and how to set them.
R has created a ShelveLocGood dummy variable that takes on a value of
1 if the shelving location is good, and 0 otherwise. It has also created a
ShelveLocMedium dummy variable that equals 1 if the shelving location is
medium, and 0 otherwise. A bad shelving location corresponds to a zero
for each of the two dummy variables. The fact that the coefficient for
3.6 Lab: Linear Regression 119
ShelveLocGood in the regression output is positive indicates that a good
shelving location is associated with high sales (relative to a bad location).
And ShelveLocMedium has a smaller positive coefficient, indicating that a
medium shelving location leads to higher sales than a bad shelving location
but lower sales than a good shelving location.
3.6.7 Writing Functions
As we have seen, R comes with many useful functions, and still more functions
are available by way of R libraries. However, we will often be interested
in performing an operation for which no function is available. In this
setting, we may want to write our own function. For instance, below we
provide a simple function that reads in the ISLR and MASS libraries, called
LoadLibraries(). Before we have created the function, R returns an error if
we try to call it.
> LoadLibraries
Error: object 'LoadLibraries ' not found
> LoadLibraries()
Error: could not find function " LoadLibraries"
We now create the function. Note that the + symbols are printed by R and
should not be typed in. The { symbol informs R that multiple commands
are about to be input. Hitting Enter after typing { will cause R to print the
+ symbol. We can then input as many commands as we wish, hitting Enter
after each one. Finally the } symbol informs R that no further commands
will be entered.
> LoadLibraries= function (){
+ library(ISLR)
+ library(MASS)
+ print("The libraries have been loaded .")
+ }
Now if we type in LoadLibraries, R will tell us what is in the function.
> LoadLibraries
function (){
library(ISLR)
library(MASS)
print("The libraries have been loaded .")
}
If we call the function, the libraries are loaded in and the print statement
is output.
> LoadLibraries()
[1] "The libraries have been loaded ."
120 3. Linear Regression
3.7 Exercises
Conceptual
1. Describe the null hypotheses to which the p-values given in Table 3.4
correspond. Explain what conclusions you can draw based on these
p-values. Your explanation should be phrased in terms of sales, TV,
radio, and newspaper, rather than in terms of the coefficients of the
linear model.
2. Carefully explain the differences between the KNN classifier and KNN
regression methods.
3. Suppose we have a data set with five predictors, X1 = GPA, X2 = IQ,
X3 = Gender (1 for Female and 0 for Male), X4 = Interaction between
GPA and IQ, and X5 = Interaction between GPA and Gender. The
response is starting salary after graduation (in thousands of dollars).
Suppose we use least squares to fit the model, and get ??^0 = 50, ??^1 =
20, ??^2 = 0.07, ??^3 = 35, ??^4 = 0.01, ??^5 = ???10.
(a) Which answer is correct, and why?
i. For a fixed value of IQ and GPA, males earn more on average
than females.
ii. For a fixed value of IQ and GPA, females earn more on
average than males.
iii. For a fixed value of IQ and GPA, males earn more on average
than females provided that the GPA is high enough.
iv. For a fixed value of IQ and GPA, females earn more on
average than males provided that the GPA is high enough.
(b) Predict the salary of a female with IQ of 110 and a GPA of 4.0.
(c) True or false: Since the coefficient for the GPA/IQ interaction
term is very small, there is very little evidence of an interaction
effect. Justify your answer.
4. I collect a set of data (n = 100 observations) containing a single
predictor and a quantitative response. I then fit a linear regression
model to the data, as well as a separate cubic regression, i.e. Y =
??0 + ??1X + ??2X2 + ??3X3 + .
(a) Suppose that the true relationship between X and Y is linear,
i.e. Y = ??0 + ??1X + . Consider the training residual sum of
squares (RSS) for the linear regression, and also the training
RSS for the cubic regression. Would we expect one to be lower
than the other, would we expect them to be the same, or is there
not enough information to tell? Justify your answer.
3.7 Exercises 121
(b) Answer (a) using test rather than training RSS.
(c) Suppose that the true relationship between X and Y is not linear,
but we don't know how far it is from linear. Consider the training
RSS for the linear regression, and also the training RSS for the
cubic regression. Would we expect one to be lower than the
other, would we expect them to be the same, or is there not
enough information to tell? Justify your answer.
(d) Answer (c) using test rather than training RSS.
5. Consider the fitted values that result from performing linear regression
without an intercept. In this setting, the ith fitted value takes
the form
y^i = xi??, ^
where
??^ =
n
i=1
xiyi

/
n
i=1
x2
i

. (3.38)
Show that we can write
y^i = n
i=1
aiyi .
What is ai?
Note: We interpret this result by saying that the fitted values from
linear regression are linear combinations of the response values.
6. Using (3.4), argue that in the case of simple linear regression, the
least squares line always passes through the point (¯x, y¯).
7. It is claimed in the text that in the case of simple linear regression
of Y onto X, the R2 statistic (3.17) is equal to the square of the
correlation between X and Y (3.18). Prove that this is the case. For
simplicity, you may assume that ¯x = ¯y = 0.
Applied
8. This question involves the use of simple linear regression on the Auto
data set.
(a) Use the lm() function to perform a simple linear regression with
mpg as the response and horsepower as the predictor. Use the
summary() function to print the results. Comment on the output.
For example:
122 3. Linear Regression
i. Is there a relationship between the predictor and the response?
ii. How strong is the relationship between the predictor and
the response?
iii. Is the relationship between the predictor and the response
positive or negative?
iv. What is the predicted mpg associated with a horsepower of
98? What are the associated 95 % confidence and prediction
intervals?
(b) Plot the response and the predictor. Use the abline() function
to display the least squares regression line.
(c) Use the plot() function to produce diagnostic plots of the least
squares regression fit. Comment on any problems you see with
the fit.
9. This question involves the use of multiple linear regression on the
Auto data set.
(a) Produce a scatterplot matrix which includes all of the variables
in the data set.
(b) Compute the matrix of correlations between the variables using
the function cor(). You will need to exclude the name variable, cor() which is qualitative.
(c) Use the lm() function to perform a multiple linear regression
with mpg as the response and all other variables except name as
the predictors. Use the summary() function to print the results.
Comment on the output. For instance:
i. Is there a relationship between the predictors and the response?
ii. Which predictors appear to have a statistically significant
relationship to the response?
iii. What does the coefficient for the year variable suggest?
(d) Use the plot() function to produce diagnostic plots of the linear
regression fit. Comment on any problems you see with the fit.
Do the residual plots suggest any unusually large outliers? Does
the leverage plot identify any observations with unusually high
leverage?
(e) Use the * and : symbols to fit linear regression models with
interaction effects. Do any interactions appear to be statistically
significant?
(f) Try a few different transformations of the variables, such as
log(X), ???
X, X2. Comment on your findings.
3.7 Exercises 123
10. This question should be answered using the Carseats data set.
(a) Fit a multiple regression model to predict Sales using Price,
Urban, and US.
(b) Provide an interpretation of each coefficient in the model. Be
careful-some of the variables in the model are qualitative!
(c) Write out the model in equation form, being careful to handle
the qualitative variables properly.
(d) For which of the predictors can you reject the null hypothesis
H0 : ??j = 0?
(e) On the basis of your response to the previous question, fit a
smaller model that only uses the predictors for which there is
evidence of association with the outcome.
(f) How well do the models in (a) and (e) fit the data?
(g) Using the model from (e), obtain 95 % confidence intervals for
the coefficient(s).
(h) Is there evidence of outliers or high leverage observations in the
model from (e)?
11. In this problem we will investigate the t-statistic for the null hypothesis
H0 : ?? = 0 in simple linear regression without an intercept. To
begin, we generate a predictor x and a response y as follows.
> set.seed(1)
> x=rnorm(100)
> y=2*x+rnorm (100)
(a) Perform a simple linear regression of y onto x, without an intercept.
Report the coefficient estimate ??^, the standard error of
this coefficient estimate, and the t-statistic and p-value associated
with the null hypothesis H0 : ?? = 0. Comment on these
results. (You can perform regression without an intercept using
the command lm(y???x+0).)
(b) Now perform a simple linear regression of x onto y without an
intercept, and report the coefficient estimate, its standard error,
and the corresponding t-statistic and p-values associated with
the null hypothesis H0 : ?? = 0. Comment on these results.
(c) What is the relationship between the results obtained in (a) and
(b)?
(d) For the regression of Y onto X without an intercept, the tstatistic
for H0 : ?? = 0 takes the form ??/^ SE(??^), where ??^ is
given by (3.38), and where
SE(??^) = !
n
i=1(yi ??? xi??^)2
(n ??? 1)
n
i=1 x2
i
.
124 3. Linear Regression
(These formulas are slightly different from those given in Sections
3.1.1 and 3.1.2, since here we are performing regression
without an intercept.) Show algebraically, and confirm numerically
in R, that the t-statistic can be written as
(
???n ??? 1)
n
i=1 xiyi
(

n
i=1 x2
i )(
n
i=1 y2
i ) ??? (

n
i=1 xiyi )2 .
(e) Using the results from (d), argue that the t-statistic for the regression
of y onto x is the same as the t-statistic for the regression
of x onto y.
(f) In R, show that when regression is performed with an intercept,
the t-statistic for H0 : ??1 = 0 is the same for the regression of y
onto x as it is for the regression of x onto y.
12. This problem involves simple linear regression without an intercept.
(a) Recall that the coefficient estimate ??^ for the linear regression of
Y onto X without an intercept is given by (3.38). Under what
circumstance is the coefficient estimate for the regression of X
onto Y the same as the coefficient estimate for the regression of
Y onto X?
(b) Generate an example in R with n = 100 observations in which
the coefficient estimate for the regression of X onto Y is different
from the coefficient estimate for the regression of Y onto X.
(c) Generate an example in R with n = 100 observations in which
the coefficient estimate for the regression of X onto Y is the
same as the coefficient estimate for the regression of Y onto X.
13. In this exercise you will create some simulated data and will fit simple
linear regression models to it. Make sure to use set.seed(1) prior to
starting part (a) to ensure consistent results.
(a) Using the rnorm() function, create a vector, x, containing 100
observations drawn from a N(0, 1) distribution. This represents
a feature, X.
(b) Using the rnorm() function, create a vector, eps, containing 100
observations drawn from a N(0, 0.25) distribution i.e. a normal
distribution with mean zero and variance 0.25.
(c) Using x and eps, generate a vector y according to the model
Y = ???1+0.5X + . (3.39)
What is the length of the vector y? What are the values of ??0
and ??1 in this linear model?
3.7 Exercises 125
(d) Create a scatterplot displaying the relationship between x and
y. Comment on what you observe.
(e) Fit a least squares linear model to predict y using x. Comment
on the model obtained. How do ??^0 and ??^1 compare to ??0 and
??1?
(f) Display the least squares line on the scatterplot obtained in (d).
Draw the population regression line on the plot, in a different
color. Use the legend() command to create an appropriate legend.
(g) Now fit a polynomial regression model that predicts y using x
and x2. Is there evidence that the quadratic term improves the
model fit? Explain your answer.
(h) Repeat (a)-(f) after modifying the data generation process in
such a way that there is less noise in the data. The model (3.39)
should remain the same. You can do this by decreasing the variance
of the normal distribution used to generate the error term
 in (b). Describe your results.
(i) Repeat (a)-(f) after modifying the data generation process in
such a way that there is more noise in the data. The model
(3.39) should remain the same. You can do this by increasing
the variance of the normal distribution used to generate the
error term  in (b). Describe your results.
(j) What are the confidence intervals for ??0 and ??1 based on the
original data set, the noisier data set, and the less noisy data
set? Comment on your results.
14. This problem focuses on the collinearity problem.
(a) Perform the following commands in R:
> set.seed(1)
> x1=runif (100)
> x2=0.5*x1+rnorm (100)/10
> y=2+2*x1+0.3*x2+rnorm (100)
The last line corresponds to creating a linear model in which y is
a function of x1 and x2. Write out the form of the linear model.
What are the regression coefficients?
(b) What is the correlation between x1 and x2? Create a scatterplot
displaying the relationship between the variables.
(c) Using this data, fit a least squares regression to predict y using
x1 and x2. Describe the results obtained. What are ??^0, ??^1, and
??^2? How do these relate to the true ??0, ??1, and ??2? Can you
reject the null hypothesis H0 : ??1 = 0? How about the null
hypothesis H0 : ??2 = 0?
126 3. Linear Regression
(d) Now fit a least squares regression to predict y using only x1.
Comment on your results. Can you reject the null hypothesis
H0 : ??1 = 0?
(e) Now fit a least squares regression to predict y using only x2.
Comment on your results. Can you reject the null hypothesis
H0 : ??1 = 0?
(f) Do the results obtained in (c)-(e) contradict each other? Explain
your answer.
(g) Now suppose we obtain one additional observation, which was
unfortunately mismeasured.
> x1=c(x1, 0.1)
> x2=c(x2, 0.8)
> y=c(y,6)
Re-fit the linear models from (c) to (e) using this new data. What
effect does this new observation have on the each of the models?
In each model, is this observation an outlier? A high-leverage
point? Both? Explain your answers.
15. This problem involves the Boston data set, which we saw in the lab
for this chapter. We will now try to predict per capita crime rate
using the other variables in this data set. In other words, per capita
crime rate is the response, and the other variables are the predictors.
(a) For each predictor, fit a simple linear regression model to predict
the response. Describe your results. In which of the models is
there a statistically significant association between the predictor
and the response? Create some plots to back up your assertions.
(b) Fit a multiple regression model to predict the response using
all of the predictors. Describe your results. For which predictors
can we reject the null hypothesis H0 : ??j = 0?
(c) How do your results from (a) compare to your results from (b)?
Create a plot displaying the univariate regression coefficients
from (a) on the x-axis, and the multiple regression coefficients
from (b) on the y-axis. That is, each predictor is displayed as a
single point in the plot. Its coefficient in a simple linear regression
model is shown on the x-axis, and its coefficient estimate
in the multiple linear regression model is shown on the y-axis.
(d) Is there evidence of non-linear association between any of the
predictors and the response? To answer this question, for each
predictor X, fit a model of the form
Y = ??0 + ??1X + ??2X2 + ??3X3 + .