---
title: "ISLR_Ch3_LinearRegression"
author: "Steve Newns"
date: "April 20, 2018"
output: html_document
---
# 3.6.2 Simple Linear Regression
```{r}
library(MASS)
library(ISLR)
```

`MASS` library contains `Boston` data set, which records `medv` (median house value) for 506 neighborhoods around Boston. We will **seek to predict `medv` using 13 predictors such as `rm` (avg #  of rooms per house), `age` (avg age of houses), + `lstat` (% of households with low socioeconomic status)**

```{r}
#fix(Boston)
names(Boston)
```

We will start by using `lm()` to fit a simple linear regression model, with `medv` as response + `lstat` as predictor. The basic
```{r}
simpleLM <- lm(medv ~ lstat, data = Boston)
```

Typing `simpleLM` gives some basic info about the model, but  `summary(simpleLM)` gives more detailed information like p-values + standard errors for the coefficients, as well as the R2 statistic and F-score
```{r}
simpleLM
summary(simpleLM)
```

Can use `names()` to find out what other pieces of info are stored in the fit model. 
```{r}
names(simpleLM)
```

Can extract quantities by name (`lm.fit$coefficients`) but it is safer to use **extractor functions** like `coef()` to access them.
```{r}
coef(simpleLM)
```

To obtain a CI for the coefficient estimates, use `confint()`
```{r}
confint(simpleLM)
```

`predict()` can be used to produce CI's *and* **prediction intervals** for the prediction of `medv` for a given value of `lstat`.

  * **CI's** tell you about how well you have determined the parameter of interest
    * Repeat an experiment many times + calculate a CI from each sample, we'd expect ~95% of those intervals to include the true value of the population parameter (have 95% confidence the interval contains the true parameter value)
    * Key point = CI's tell you about the likely location of the true population parameter.

  * **Prediction intervals** tell you *where you can expect to see the next DP sampled*. 
    * Assume the data really are randomly sampled from a Gaussian distribution. 
    * Collect a sample of data + calculate a PI
    * Then sample 1 more value from the population
    * Repeat many times + expect that next value to lie within such a prediction interval in 95% of samples.
    * Key point = A PI tells you *about the distribution of values*, NOT the uncertainty in determining the true population parameter
    * PI's must account for both the **uncertainty** in knowing the value of the true population parameter, plus **scatter/noise** So
    * **A PI is always wider than a confidence interval (less confident)**
    
  * CI's give a range for `E[y???x]`, PI's give a range *for `y` itself*.
  * Naturally, best guess for y = `E[y???x]`, so both intervals are centered around the same value, `x??^`, but with *different standard errors*
  * We *guess** the expected value of `E[y???x]` more precisely than we estimate `y` itself** so **estimating y requires including the variance that comes from the true error term**
    * Ex: Could get perfect estimates of ?? (coefficients), so an estimate of `E[y???x]` would be perfect, but we still wouldn't be sure what `y` itself was b/c *there is a true error term we need to consider (irreducible)*. 
    * Our confidence "interval" would just be a single DP, as we estimate E[y???x] exactly right, but our PI would be wider b/c we take the true error term into account.
```{r}
# create a column (vector) of 5,10,15 as new lstat values to use to predict medv's valie + CI with the model
print("CI")
predict(object = simpleLM, newdata = data.frame(lstat=c(5,10,15)), interval="confidence")
print("PI")
predict(object = simpleLM, newdata = data.frame(lstat=c(5,10,15)), interval="prediction")
```

95% CI associated with a `lstat` value = 10 is `(24.47, 25.63)`, and the 95 % **prediction interval** = `(12.828, 37.28)`. As expected, the *CI and PI's are centered around the same point* = a predicted value of `25.05` for `medv` when `lstat` = 10, but
the latter are substantially wider.

Now plot `medv` + `lstat` along w/ the **least squares regression line** using `plot()` and `abline()`
```{r}
plot(Boston$lstat,Boston$medv)
abline(simpleLM, col="red", lwd=3)
```

There is some evidence for **non-linearity** in this relationship, as seen by the upper-left area of the plot

`abline()` can be used to draw *any* line, not just the least squares regression line. To draw a line with intercept `a` and slope `b`, type `abline(a,b)`. 

```{r}
par(mfrow=c(2,3))
plot(Boston$lstat,Boston$medv)
abline(simpleLM, lwd=3)
plot(Boston$lstat,Boston$medv)
abline(simpleLM, lwd=3, col="red")
plot(Boston$lstat, Boston$medv, col="red")
plot(Boston$lstat, Boston$medv, pch=20) # change DP shape
plot(Boston$lstat, Boston$medv, pch="+") # change DP shape
plot(1:20,1:20,pch =1:20)
```

Next we examine some **diagnostic plots**. 4 are automatically produced by applying `plot()` directly to the output from a linear model.
```{r}
par(mfrow=c(2,2)) # #see all plots together
plot(simpleLM) # residuals v fitted, QQ, Scale-Location, residuals v leverage
```

There looks to be a curving pattern as we move down the x-axis to the right,almost like a fish hook, indicating non-linearity between `lstat` and `medv` such that our residuals are not **identically and independently distributed (IID)**

**Normal Q-Q plot** = empirical quantiles vs. theoretical Standard Normal quantiles. For our model, our fitted quantiles are "too large", indicating a slight-right skew.

  * quantiles from a N(??,??2) population should be linearly related to Standard Normal quantiles
    * Let xp = pth quantile from a N(??,??2), then probability of *non-standard normal* X being below a real value x is P(X <= x) = p, so clearly, for the *standardized* normal Z, P(Z <= (xp - ??)/??)) = p, and therefore xp = ?? + zp*??

**"Residuals vs Fitted" and "Scale/Spread-Location"** = essentially the same, showing if there is a trend to the residuals (if they're spread equally along the ranges of predictors). This again shows the fish-hook pattern indicating non-IID residuals and non-linearity in the relationship.

**Residuals vs. Leverage** = tells us which DP's have the greatest influence on the regression (i.e. high leverage points).

  * Not all outliers are influential in linear regression so results wouldn't be much different whether included or not.
  * Watch for outlying values at the upper right or lower right corners = spots where cases can be influential against a regression line. 
  * Look for cases outside of a **Cook's distance** (dashed line) , meaning they have high **Cook's distance scores** + are influential to the regression results, as the results will be altered if we exclude those cases.
    * Cook's distance = function of the leverage and standardized residual associated with each data poin
  * Looks' like we have several high leverage points that influence our regression as noted by being quite far from the dotted Cook's distance line and being noted with annotations (215,413,375) 

Alternatively, can compute residuals from a linear regression fit using `residuals()`, or `rstudent()` to return **studentized residuals**, and we can use this to plot the residuals against the fitted values (1st + 3rd plots above)

  * **studentized residuals** = alternative criterion for IDing outliers
    * basic idea = delete observations ` at a time, each time refitting the regression model on the remaining `n-1` observations
    * Then, compare the observed response values to their fitted values based on the models with the `i`th observation deleted. 
    * This produces **deleted residuals**, and standardizing the deleted residuals produces studentized residuals.
    *  deleted residual `d_i = y_i - y_i_hat` = observed response for ith variable minus predicted response based on the estimated model with the ith observation deleted
    
      * data point i being influential implies the DP "pulls" the estimated regression line towards itself, which means the observed response would be close to the predicted response. 
      * But, if removed, the estimated regression line would "bounce back" away from the observed response, thereby **resulting in a large deleted residual** (i.e. a DP w/ a large deleted residual suggests the DP is influential)
      
    * "Largeness" of deleted residuals depend on units of measurement, just as ordinary residuals do
    * Solve this by **standardizing** = dividing each deleted residual by an estimate of its standard deviation.
    * studentized residual, `t_i = d_i/sd(d_i) = e_i/sqrt(MSE_i(1-h_ii)` = same as ordinary residual `e_i` divided by a factor that includes the MSE *based on estimated model **w/ ith observation deleted**, `MSE_i`*, and the **leverage, `h_ii`**
    * TLDR: studentized residual = just a deleted residual divided by its estimated standard deviation
      
      * In general, studentized residuals = more effective than standardized residuals. 
      * If an observation has a studentized residual > 3 (in absolute value), can call it an outlier.
```{r}
par(mfrow=c(2,1))
plot(x=predict(simpleLM),y=residuals(simpleLM))
plot(x=predict(simpleLM),y=rstudent(simpleLM))
```

On the basis of the residual plots, there is some evidence of non-linearity. **Leverage statistics** can be computed for any # of predictors using `hatvalues()`.
```{r}
#hatvalues(simpleLM)
plot(hatvalues(simpleLM))
which.max(hatvalues(simpleLM))
```
We see the index of the observation has the largest leverage statistic as 365

### 3.6.3 Multiple Linear Regression

To fit a multiple linear regression model using **least squares**, again use `lm()` w/ `n` predictors
```{r}
multLR <- lm(medv ~ lstat + age, Boston)
summary(multLR)
```

```{r}
allMultLR <- lm(medv ~ ., Boston)
summary(allMultLR)
```

Can access individual components of a **summary object** by name (`?summary.lm`). Hence `summary(lm.fit)$r.sq` gives us R2 + `summary(lm.fit)$sigma` gives us the **RSE**. `vif()`, part of the `car` package (functions to apply to a fitted regression model) can be used to compute **variance inflation** factors. Most VIF's are low to moderate for this data.

* **VIF** = ratio of the variance of ??_j when fitting the full model divided by the variance of ??_j if fit on its own
  * smallest possible = 1 (indicates complete absence of collinearity)
  * Typically in practice, there's a small amount of collinearity among predictors
  * Rule of thumb = VIF > 5 or 10 indicates a problematic amount of collinearity. 
  * `VIF(??_j ) = 1/(1 ??? R2_(X_j | X_-j)` where `R2_(X_j | X_-j)` = R2 from a regression of `X_j` onto all other predictors, `X_-j`)
  * If `R2_(X_j | X_-j)` = close to 1, collinearity is present, so the VIF will be large.
```{R}
#install.packages("car",dependencies=TRUE)
#update.packages(checkBuilt = TRUE)
library(car)
vif(allMultLR)
```

What if we'd like to perform a regression using all variables but one?
```{r}
multLR_noAge <- lm(medv ~ . - age, Boston) # largest p-value in full model
summary(multLR_noAge)
```

Alternatively, `update()` can be used.
```{r}
multLR_noAge <- update(allMultLR, ~ . - age)
summary(multLR_noAge)
```

### 3.6.4 Interaction Terms
It is easy to include **interaction terms** in a linear model, such as w/ syntax `lstat:black` = interaction term between `lstat` +  `black`. The syntax `lstat*age` simultaneously includes `lstat`, `age`, *and the interaction term `lstat × age`* as predictors, as it is a shorthand for `lstat+age+lstat:age`
```{r}
multLR_interaction <- lm(medv ~ lstat*age, Boston)
summary(multLR_interaction)
```

### 3.6.5 Non-linear Transformations of Predictors

`lm()` function can also accommodate non-linear transformations of predictors. Given a predictor `X`, can create a predictor `X2`
using `I(X^2)` (`I()` is needed since `^` has a special meaning in a formula)
```{r}
multLR_transform <- lm(medv ~ lstat + I(lstat^2), Boston)
summary(multLR_transform)
```

The near-0 p-value associated with the quadratic term suggests it leads to an improved model. **Use `?anova()` to further quantify the extent to which the quadratic fit is superior to the linear fit.**

* `anova()` = computes analysis of variance (or deviance) tables for 1+ fitted model objects, tested in order specified whether the model terms are significant
```{r}
anova(lm(medv~lstat,Boston),multLR_transform)
```

`anova()` performs a **hypothesis test** comparing the 2 models w/ a null H0 = the 2 models fit the data equally well, + the alternative H1 = the full model (1st given) is superior. 

Here the **F-statistic** = 135 + the associated p-value = virtually 0. This info provides very clear evidence that the quadratic model is far superior (not surprising, since earlier we saw evidence for non-linearity in the relationship between `medv` + `lstat`
```{r}
par(mfrow=c(2,2))
plot(multLR_transform)
```

There looks to be no pattern for our normal nor studentizes residuals, indicating linearity in our fit model such that our residuals *are* identically and independently distributed (IID)

**Normal Q-Q plot** = For our new model, our fitted quantiles are still "too large" in the upper-right, indicating a slight-right skew.

**Residuals vs. Leverage** = Looks like we still have several high leverage points that influence our regression as noted by being quite far from the dotted Cook's distance line and being noted with annotations

In order to create a cubic fit, we can include a predictor of the form `I(X^3)`. However, this approach can start to get cumbersome for higher-order polynomials. A better approach involves using `poly()` to create the polynomial within `lm()`
```{r}
# produce a fifth-order polynomial fit
fifthOrder <- lm(medv ~ poly(lstat,5), Boston)
summary(fifthOrder)
```

This suggests that including additional polynomial terms, up to 5th order, leads to an improvement in the model fit! However, further investigation of data reveals that no polynomial terms beyond 5th order have significant p-values in a regression fit. Of course, we are in no way restricted to only using polynomial transformations of the predictors. Here we try a **log transformation.**
```{r}
summary(lm(medv ~ log(rm), Boston))
```
         
### 3.6.6 Qualitative Predictors
`Carseats` = attempt to predict `Sales` (child car seat sales) in 400 locations based on a number of predictors.
```{r}
library(ISLR)
library(tidyverse)
glimpse(Carseats)
```

Data includes qualitative predictors (Shelveloc = indicator of the quality of the shelving location [space within a store in which the car seat is displayed-at each location] taking 3 possible values = Bad, Medium, and Good)

Given a qualitative variable such as Shelveloc, R generates dummy variables *automatically*
```{r}
carSeatModel1 <- lm(Sales ~ . + Income:Advertising + Price:Age, Carseats)
summary(carSeatModel1)
```

`contrasts()` returns the coding R uses for the dummy `contrasts()` variables.
```{r}
contrasts(Carseats$ShelveLoc)
```

R has created a `ShelveLocGood` dummy variable that = 1 1 if shelving location = good, + 0 otherwise and `ShelveLocMedium` dummy variable = 1 if shelving location = medium, + 0 otherwise. A "bad"" shelving location = 0 for each of the other 2 dummy variables. The fact that the coefficient for `ShelveLocGood`= positive indicates a good shelving location is associated w/ high sales (***relative to a bad location***) + `ShelveLocMedium` = a smaller positive coefficient, indicating a medium shelving location leads to higher sales than a bad shelving location but lower sales than a good shelving location.

### 3.6.7 Writing Functions

Write a simple function that reads in the ISLR and MASS libraries
```{r}
LoadLibraries <- function() {
  library(ISLR)
  library(MASS)
  print("Libraries have been loaded")
}
LoadLibraries()
```

## 3.7 Exercises
### Conceptual
**1. Describe the null hypotheses to which the p-values given in Table 3.4 (hyperplane) correspond. Explain what conclusions you can draw based on these p-values. Your explanation should be phrased in terms of sales, TV, radio, and newspaper, rather than in terms of the coefficients of the linear model.**

* H01 : There is no relationship between TV and Sales (B1 = 0)
* H02 : There is no relationship between Radio and Sales (B2 = 0)
* H03 : There is no relationship between Newspaper and Sales (B3 = 0)
* If our p-values (probability of observing any value equal>= |t| larger in absolute value, assuming Bx = 0) for any of these predictors are below our threshold (here, it's .05), then we have the confidence to reject the null hypothesis that there is no relationship between the specified predictor and Sales the values we've observed are not due to chance.

**2. Carefully explain the differences between the KNN classifier and KNN regression methods.**

* 

**3. Suppose we have a data set with 5 predictors, X1 = GPA, X2 = IQ, X3 = Gender (1 = Female, 0 = Male), X4 = Interaction between GPA and IQ, and X5 = Interaction between GPA and Gender. The response = starting salary after graduation (in thousands of $). Suppose we use least squares to fit the model, and get B^ = 50, B^1 = 20, B^2 = 0.07, B^3 = 35, B^4 = 0.01, B^5 = -10**

* **(a) Which answer is correct, and why?** = 
    * i. For a fixed value of IQ and GPA, males earn more on average than females.
    * ii. For a fixed value of IQ and GPA, females earn more on average than males.
    * **iii. For a fixed value of IQ and GPA, males earn more on average than females provided that the GPA is high enough. **
    * Least Squares model for females = 50 + 20X1 + .07X2 + 35 + .01X4 -10X5 == 85 + 20(GPA) + .07(IQ) + .01(GPA:IQ) - 10(GP:Gender) == 85 + 10(GPA) + .07(IQ) + .01(GPA:IQ)
    * Least Squares model for males = 50 + 20X1 + .07X2 + .01X4 == 50 + 20(GPA) + .07(IQ) + .01(GPA:IQ)
    * Therefore, starting salary for males is greater if GPA is large enough such that GPA >= 3.5 (simplifying, adding the 2 models gives 35 = 10GPA)
    * iv. For a fixed value of IQ and GPA, females earn more on average than males provided that the GPA is high enough.
* **(b) Predict the salary of a female with IQ of 110 and a GPA of 4.0.**
    * 85 + 10(GPA) + .07(IQ) + .01(GPA:IQ) == 85 + 10(4) + .07(110) + .01((110\*4)) = 85 + 40 + 7.7 + 4.4 = $137.1K/yr
* **(c) T/F: Since the coefficient for the GPA/IQ interaction term is very small, there is very little evidence of an interaction effect. Justify your answer.**
    * False. The actual value of the interaction term's coefficient does not matter in deciding if there's evidence of an interaction effect is present. What we must do is check the p-value for the coefficient which tests the H0: B^4 = 0 and associate the results with the t-statistic and F score
      * **F-statistic** = ((TSS-RSS)/p)/(RSS/(n-p-1)) = for no relationship between the response and predictors, expect F-statistic to take on a value close to 1
      * TSS = sum of (y_i - y_bar)^2, RSS = sum of (y_i - y_i_hat)^2

**4. I collect a set of data (n = 100 observations) containing a single predictor and a quantitative response. I then fit a linear regression model to the data, as well as a separate cubic regression, i.e. Y = B0 + B1X + B2X^2 + B3X^3 + e.**

* **(a) Suppose the true relationship between X + Y is linear, i.e. Y = B0 + B1X + e. Consider the *training* residual sum of squares (RSS) for the *linear* regression, and also the training RSS for the *cubic* regression. Would we expect one to be lower than the other, would we expect them to be the same, or is there not enough information to tell? Justify your answer.**
    * We cannot expect the cubic regression to be better, as it *is still a linear model*. Therefore, we do not have enough information to tell which training RSS would be lower. We *could* estimate that by adding the extra cubic term, our least squares fit *could* be a bit closer to the true regression line, and thereby giving the cubic model a lower training RSS, but it may not.
* **(b) Answer (a) using test rather than training RSS.* **
    * We still do not have enough to make confident estimates, but a conservative one would be that the cubic model *could* overfit the data (move towards non-linear fits) and therefore its test RSS would be higher.
* **(c) Suppose the true relationship between X + Y is *not* linear, but we *don't know how far it is from linear*. Consider the training RSS for the linear regression + the training RSS for the cubic regression. Would we expect one to be lower than the other, would we expect them to be the same, or is there not enough information to tell? Justify your answer.* **
    * Even though we do not know how far the relationship is from linear, the fact that we have added in a higher-order polynomial in the cubic regression would give a more flexible model that would better fit a non-linear relationship, thereby giving a lower training RSS
* **(d) Answer (c) using test rather than training RSS.* **
    * We don't have enough info say which test RSS would be lower, since we don't know how far it is from linear. If the true non-linear relationship is *closer to linear than cubic*, linear regression test RSS could be lower, or vice versa (see: Bias-Variance trade-off)

**5. Consider the fitted values that result from performing linear regression *without an intercept*. In this setting, the ith fitted value takes the form y^_i = x_iB, where B = sum_i[x_i*y_i]/sum_k[x_k^2]. Show that we can write y^_i = sum_j[a_j*y_j], and what is a_j?**
  * **Note: We interpret this result by saying that the fitted values from linear regression are *linear combinations* of the response values.**
  * Plugging in sum_i[x_i\*y_i]/sum_k[x_k^2] for B in y^_i = x_iB, we get y^_i = x_i(sum_j[x_j\*y_j]/sum_k[x_k^2]), flipping around terms, we get sum_j[(x_i\*x_j)/sum_k[x_k^2]]\*y_i so a_j = (x_i\*x_j)/sum_k[x_k^2]]

**6. Using (3.4), argue that in the case of simple linear regression, the least squares line always passes through the point (x¯, y¯).**

**7. It is claimed in the text that in the case of simple linear regression of Y onto X, the R2 statistic (3.17) is equal to the square of the correlation between X and Y (3.18). Prove that this is the case. For simplicity, you may assume that ¯x = ¯y = 0.**

### Applied
8. This question involves the use of simple linear regression on the Auto data set.

**(a) Use the lm() function to perform a simple linear regression with mpg as the response and horsepower as the predictor. Use the summary() function to print the results. Comment on the output. For example:**
```{r}
library(ISLR)
data(Auto)

lm1 <- lm(mpg~horsepower,Auto)
summary(lm1)
```
  * **i. Is there a relationship between the predictor and the response?**
    * The low p-value and \*\*\* significance level indicates that we can reject the null H0: B1 = 0 and say we there is a negative relationship between predictor `horsepower` and response `mpg` 
  * **ii. How strong is the relationship between the predictor and the response?**
      * The low p-value and \*\*\* significance level indicates a relationship, and the moderate R2 value indicated a moderately strong relationship, which we can look at with a scatterplot and correlation calculation
```{r}
library(ggplot2)
ggplot(Auto,aes(horsepower,mpg)) + 
  geom_point(colour="red")
cor(Auto$horsepower,Auto$mpg)
cor(Auto$horsepower,Auto$mpg)**2
```
  * **iii. Is the relationship between the predictor and the response positive or negative?**
    * It's clearly negative
  * **iv. What is the predicted mpg associated with a horsepower of 98? What are the associated 95 % confidence + prediction intervals?**
    * Using our regression formula `39.935861 - 0.157845(98)`, we get:
```{r}
print("predicted mpg for hp = 98")
39.935861 - 0.157845*(98)
print("CI for predicted mpg for hp = 98")
predict(object = lm1, newdata = data.frame(horsepower=98), interval="confidence")
print("PI for predicted mpg for hp = 98")
predict(object = lm1, newdata = data.frame(horsepower=98), interval="prediction")
```
**(b) Plot the response and the predictor. Use the abline() function to display the least squares regression line.**
```{r}
#coef(lm1)
ggplot(Auto,aes(horsepower,mpg)) + 
  geom_point(colour="red") + 
  geom_abline(intercept = 39.9358610, slope = -0.1578447 )
```
**(c) Use the plot() function to produce diagnostic plots of the least squares regression fit. Comment on any problems you see with the fit.**
```{r}
par(mfrow=c(2,2))
plot(lm1)
```
  * We can see a pattern in our residuals in both the residuals and Scale-Location plot, indicating that our residuals are not identically and independently distributed (IID) and our relationship between the predictor and response is non-linear.
  * In our Q-Q plot, we have the fitted quantiles are raising a bit too high at the higher end of theoretical quantiles, indicating a right skew.
  * For our Residuals vs Leverage, we have what looks like 2-possible high levarage points (indicated by annotations of 117 and 94), and some outliers (> 2 or < -2 on the y-axis), which means these points are influencing our regression of `mpg` onto `horsepower`

**9. This question involves the use of multiple linear regression on the Auto data set.**
**(a) Produce a scatterplot matrix which includes all of the variables in the data set.**
```{r}
pairs(~ mpg+displacement+horsepower+weight+acceleration, Auto)
```
**(b) Compute the matrix of correlations between the variables using (exclude `name`, as its qualitative)**
```{r}
cor(select(Auto,-name))
```

**(c) Use the lm() function to perform a multiple linear regression with `mpg` as the response and all other variables except name as
the predictors. Use the summary() function to print the results. Comment on the output**
```{R}
lm2 <- lm(mpg~.-name,Auto)
summary(lm2)
```
  * **i. Is there a relationship between the predictors and the response?**
    * There appears to be a relationship between `mpg` and *some* predictors = `weight`,year,origin, and some `displacement`
  * **ii. Which predictors appear to have a statistically significant relationship to the response?**
    * Very sig = `weight`,year,origin, Moderately = `displacement`
  * **iii. What does the coefficient for the year variable suggest?**
    * It suggests that for each year increase, we have a .75 unit increase in `mpg`
* **(d) Use the plot() function to produce diagnostic plots of the linear regression fit. Comment on any problems you see with the fit. Do the residual plots suggest any unusually large outliers? Does the leverage plot identify any observations with unusually high leverage?**
```{r}
par(mfrow=c(2,2))
plot(lm2)
```
  * There may be a *slight* pattern in Residuals vs. Fitted, but it is not apparant in Scale-Location, indicating that we are probably able to say that our residuals are identically and independently distributed (IID) and our relationship between the predictor and response is linear.
  * In our Q-Q plot, we have the fitted quantiles are raising a bit too high at the higher end of theoretical quantiles, indicating a right skew.
  * For our Residuals vs Leverage, we have a couple outliers (> 2 or < -2 on the y-axis) and a high levarage point (14), which means these points are influencing our regression of `mpg` onto `horsepower`

* **(e) Use the \* and : symbols to fit linear regression models with interaction effects. Do any interactions appear to be statistically significant?**
```{r}
lm3 <- lm(mpg~.-name+cylinders*(.-name),Auto)
lm4 <- lm(mpg~.-name+displacement*(.-name),Auto)
lm5 <- lm(mpg~.-name+horsepower*(.-name),Auto)
lm6 <- lm(mpg~.-name+weight*(.-name),Auto)
lm7 <- lm(mpg~.-name+acceleration*(.-name),Auto)
lm8 <- lm(mpg~.-name+year*(.-name),Auto)
lm9 <- lm(mpg~.-name+origin*(.-name),Auto)
#lm3 <- lm(mpg~.+cylinders*displacement-name,Auto)
#lm3 <- lm(mpg~.+cylinders*displacement-name,Auto)
#lm3 <- lm(mpg~.+cylinders*displacement-name,Auto)
#lm3 <- lm(mpg~.+cylinders*displacement-name,Auto)
#lm3 <- lm(mpg~.+cylinders*displacement-name,Auto)
#lm3 <- lm(mpg~.+cylinders*displacement-name,Auto)
#lm3 <- lm(mpg~.+cylinders*displacement-name,Auto)
summary(lm3)
summary(lm4)
summary(lm5)
summary(lm6)
summary(lm7)
summary(lm8)
summary(lm9)
```
* **(f) Try a few different transformations of the variables, such as log(X), ??? X, X2. Comment on your findings.**
  * cylinders:year  displacement:weight   displacement:year  horsepower:year  weight:year 
  * cylinders:acceleration    weight:acceleration   acceleration:year   acceleration:origin
  
**10. This question should be answered using the Carseats data set.**
**(a) Fit a multiple regression model to predict Sales using Price, Urban, and US.**
```{r}
data("Carseats")
lm10 <- lm(Sales~Price+Urban+US,Carseats)
summary(lm10)
```
**(b) Provide an interpretation of each coefficient in the model. Be careful-some of the variables in the model are qualitative!**
**(c) Write out the model in equation form, being careful to handle the qualitative variables properly.**
  * `model = 13.043 - .054\*Price - .024\*Urban + 1.2\*US`
**(d) For which of the predictors can you reject the null hypothesis H0 : ??j = 0?**
  * `Price` and `USyes`
**(e) On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the outcome.**
```{r}
lm11 <- lm(Sales~Price+US,Carseats)
summary(lm11)
```
**(f) How well do the models in (a) and (e) fit the data?**
  * 2nd model fits barely better (R2 = .2354 vs. .2335)
**(g) Using the model from (e), obtain 95 % confidence intervals for the coefficient(s).**
```{r}
confint(lm11)
```
**(h) Is there evidence of outliers or high leverage observations in the model from (e)?**
```{r}
par(mfrow=c(2,2))
plot(lm11)
```
  * 1 high leverage point

**11. In this problem we will investigate the t-statistic for the null hypothesis H0 : ?? = 0 in simple linear regression without an intercept. To begin, we generate a predictor x and a response y as follows.**
```{r}
set.seed(1)
x <- rnorm(100)
y <- 2*x+rnorm (100)
```
**(a) Perform a simple linear regression of y onto x, without an intercept. Report the coefficient estimate ??^, the standard error of this coefficient estimate, and the t-statistic and p-value associated with the null hypothesis H0 : ?? = 0. Comment on these results. (You can perform regression without an intercept using the command lm(y???x+0).)**
**(b) Now perform a simple linear regression of x onto y without an intercept, and report the coefficient estimate, its standard error, and the corresponding t-statistic and p-values associated with the null hypothesis H0 : ?? = 0. Comment on these results.**
**(c) What is the relationship between the results obtained in (a) and (b)?**
**(d) For the regression of Y onto X without an intercept, the tstatistic for H0 : ?? = 0 takes the form ??/^ SE(??^), where ??^ is given by (3.38), and where (These formulas are slightly different from those given in Sections 3.1.1 and 3.1.2, since here we are performing regression without an intercept.) Show algebraically, and confirm numerically in R, that the t-statistic can be written as**
**(e) Using the results from (d), argue that the t-statistic for the regression of y onto x is the same as the t-statistic for the regression of x onto y.**
**(f) In R, show that when regression is performed with an intercept, the t-statistic for H0 : ??1 = 0 is the same for the regression of y onto x as it is for the regression of x onto y.**

12. This problem involves simple linear regression without an intercept.
* (a) Recall that the coefficient estimate ??^ for the linear regression of Y onto X without an intercept is given by (3.38). Under what circumstance is the coefficient estimate for the regression of X onto Y the same as the coefficient estimate for the regression of Y onto X?
* (b) Generate an example in R with n = 100 observations in which the coefficient estimate for the regression of X onto Y is different from the coefficient estimate for the regression of Y onto X.
* (c) Generate an example in R with n = 100 observations in which the coefficient estimate for the regression of X onto Y is the same as the coefficient estimate for the regression of Y onto X.

13. In this exercise you will create some simulated data and will fit simple linear regression models to it. Make sure to use set.seed(1) prior to starting part (a) to ensure consistent results.
* (a) Using the rnorm() function, create a vector, x, containing 100 observations drawn from a N(0, 1) distribution. This represents a feature, X.
* (b) Using the rnorm() function, create a vector, eps, containing 100 observations drawn from a N(0, 0.25) distribution i.e. a normal distribution with mean zero and variance 0.25.
* (c) Using x and eps, generate a vector y according to the model Y = ???1+0.5X + . (3.39) What is the length of the vector y? What are the values of ??0 and ??1 in this linear model?
* (d) Create a scatterplot displaying the relationship between x and y. Comment on what you observe.
* (e) Fit a least squares linear model to predict y using x. Comment on the model obtained. How do ??^0 and ??^1 compare to ??0 and ??1?
* (f) Display the least squares line on the scatterplot obtained in (d). Draw the population regression line on the plot, in a different color. Use the legend() command to create an appropriate legend.
* (g) Now fit a polynomial regression model that predicts y using x and x2. Is there evidence that the quadratic term improves the model fit? Explain your answer.
* (h) Repeat (a)-(f) after modifying the data generation process in such a way that there is less noise in the data. The model (3.39) should remain the same. You can do this by decreasing the variance of the normal distribution used to generate the error term in (b). Describe your results.
* (i) Repeat (a)-(f) after modifying the data generation process in such a way that there is more noise in the data. The model (3.39) should remain the same. You can do this by increasing the variance of the normal distribution used to generate the error term  in (b). Describe your results.
* (j) What are the confidence intervals for ??0 and ??1 based on the original data set, the noisier data set, and the less noisy data set? Comment on your results.

14. This problem focuses on the collinearity problem.
* (a) Perform the following commands in R:
```{r}
set.seed(1)
x1 <- runif (100)
x2 <- 0.5*x1+rnorm (100)/10
y <- 2+2*x1+0.3*x2+rnorm (100)
```
The last line corresponds to creating a linear model in which y is a function of x1 and x2. Write out the form of the linear model.  What are the regression coefficients?
* (b) What is the correlation between x1 and x2? Create a scatterplot displaying the relationship between the variables.
* (c) Using this data, fit a least squares regression to predict y using x1 and x2. Describe the results obtained. What are ??^0, ??^1, and ??^2? How do these relate to the true ??0, ??1, and ??2? Can you reject the null hypothesis H0 : ??1 = 0? How about the null hypothesis H0 : ??2 = 0?
* (d) Now fit a least squares regression to predict y using only x1. Comment on your results. Can you reject the null hypothesis H0 : ??1 = 0?
* (e) Now fit a least squares regression to predict y using only x2. Comment on your results. Can you reject the null hypothesis H0 : ??1 = 0?
* (f) Do the results obtained in (c)-(e) contradict each other? Explain your answer.
* (g) Now suppose we obtain one additional observation, which was unfortunately mismeasured.
```{r}
x1 <- c(x1, 0.1)
x2 <- c(x2, 0.8)
y <- c(y,6)
```
Re-fit the linear models from (c) to (e) using this new data. What effect does this new observation have on the each of the models? In each model, is this observation an outlier? A high-leverage point? Both? Explain your answers.

15. This problem involves the Boston data set, which we saw in the lab for this chapter. We will now try to predict per capita crime rate using the other variables in this data set. In other words, per capita crime rate is the response, and the other variables are the predictors.
* (a) For each predictor, fit a simple linear regression model to predict the response. Describe your results. In which of the models is there a statistically significant association between the predictor and the response? Create some plots to back up your assertions.
* (b) Fit a multiple regression model to predict the response using all of the predictors. Describe your results. For which predictors can we reject the null hypothesis H0 : ??j = 0?
* (c) How do your results from (a) compare to your results from (b)? Create a plot displaying the univariate regression coefficients from (a) on the x-axis, and the multiple regression coefficients from (b) on the y-axis. That is, each predictor is displayed as a single point in the plot. Its coefficient in a simple linear regression model is shown on the x-axis, and its coefficient estimate in the multiple linear regression model is shown on the y-axis.
* (d) Is there evidence of non-linear association between any of the predictors and the response? To answer this question, for each predictor X, fit a model of the form Y = ??0 + ??1X + ??2X2 + ??3X3 + .