---
title: "ISLR_Ch3_LinearRegression_exercises"
author: "Steve Newns"
date: "April 20, 2018"
output: html_document
---
```{r,message=F,warning=F}
library(tidyverse)
```
## 3.7 Exercises
### Conceptual
**1. Describe the null hypotheses to which the p-values given in Table 3.4 (hyperplane) correspond. Explain what conclusions you can draw based on these p-values. Your explanation should be phrased in terms of sales, TV, radio, and newspaper, rather than in terms of the coefficients of the linear model.**

* $H_{01}$ : There is no relationship between TV and Sales ($\beta_1$ = 0)
* $H_{02}$ : There is no relationship between Radio and Sales ($\beta_2$ = 0)
* $H_{03}$ : There is no relationship between Newspaper and Sales ($\beta_3$ = 0)
* If our **p-values** (*probability of observing any value >= $|t|$ in absolute value, assuming $\beta_x$ = 0*) for any of these predictors are below our threshold (here, it's .05), then we have the confidence to reject the null hypothesis $H_0$ that there is no relationship between the specified predictor and `Sales` such that the values we've observed are not due to chance.

**2. Carefully explain the differences between the KNN classifier and KNN regression methods.**

* **KNN classifier** = ID's the $k$ closest neighbors to a new point, $x_0$, and then estimates the **conditional probability $P(Y=j|X=x_0)$** for class $j$ as the fraction of DPs in the neighborhood where the response $y$ = class $j$, and then classifying the new $x_0$ DP to the class $j$ with the highest conditional probability (most neighbors in the neighborhood)
* **KNN regression** = ID's the $k$ closest neighbors to a new point, $x_0$, and then estimates $f(x_0)$ as the *average* of all training responses in that neighborhood = $\hat{f}(x_0)$
  * takes an abstract point as input + treats as a *real* observation + estimates the output $f(x_0)$ it would predict if real
  * becomes a function $\hat{f}(x_0)$ that outputs a prediction for any input
  * **TLDR: it's a *quantitative estimate by averaging the result of the $k$ nearest neighbors***

**3. Suppose we have a data set with 5 predictors, $X_1$ = GPA, $X_2$ = IQ, $X_3$ = Gender (1 = Female, 0 = Male), $X_4$ = Interaction between GPA and IQ, and $X_5$ = Interaction between GPA and Gender. The response $y$ = starting salary after graduation (in thousands of $). Suppose we use least squares to fit the model, and get $\hat{\beta_0}$ = 50, $\hat{\beta_1}$ = 20, $\hat{\beta_2}$ = 0.07, $\hat{\beta_3}$ = 35, $\hat{\beta_4}$ = 0.01, $\hat{\beta_5}$ = -10**

**(a) Which answer is correct, and why?**
  
  * For a fixed value of IQ and GPA, males earn more on average than females.
  * For a fixed value of IQ and GPA, females earn more on average than males.
  * **For a fixed value of IQ and GPA, males earn more on average than females provided that the GPA is high enough. **
    * Least Squares model for females = $50 + 20X_1 + .07X_2 + 35 + .01X_3 -10X_5$ == 85 + 20(GPA) + .07(IQ) + .01(GPA:IQ) - 10(GP:Gender) == 85 + 10(GPA) + .07(IQ) + .01(GPA:IQ)
    * Least Squares model for males = $50 + 20X_2 + .07X_2 + .01X_3$ == 50 + 20(GPA) + .07(IQ) + .01(GPA:IQ)
    * Therefore, starting salary for males is greater if GPA is large enough such that GPA >= 3.5 (simplifying, adding the 2 models gives 35 = 10GPA)
  * For a fixed value of IQ and GPA, females earn more on average than males provided that the GPA is high enough.

**(b) Predict the salary of a female with IQ of 110 and a GPA of 4.0.**
  
  * 85 + 10(GPA) + .07(IQ) + .01(GPA:IQ) == 85 + 10(4) + .07(110) + .01((110\*4)) = 85 + 40 + 7.7 + 4.4 = $137.1K/yr

**(c) T/F: Since the coefficient for the GPA/IQ interaction term is very small, there is very little evidence of an interaction effect. Justify your answer.**
  
  * False. The actual value of the interaction term's coefficient does not matter in deciding if there's evidence of an interaction effect is present. What we must do is check the p-value for the coefficient which tests the H0: B^4 = 0 and associate the results with the t-statistic and F score
    * **F-statistic** = $\frac {\frac {TSS-RSS} {p}} {\frac {RSS} {n-p-1}}$ = for no relationship between the response and predictors, expect F-statistic to take on a value close to 1
    * TSS = $\sum_{i=1}^{n}(y_i - \bar{y})^2$, RSS = $\sum_{i=1}^{n}(y_i - \hat{y_i})^2$

**4. I collect a set of data (n = 100 observations) containing a single predictor and a quantitative response. I then fit a linear regression model to the data, as well as a separate cubic regression, i.e. Y = B0 + B1X + B2X^2 + B3X^3 + e.**

* **(a) Suppose the true relationship between X + Y is linear, i.e. Y = B0 + B1X + e. Consider the *training* residual sum of squares (RSS) for the *linear* regression, and also the training RSS for the *cubic* regression. Would we expect one to be lower than the other, would we expect them to be the same, or is there not enough information to tell? Justify your answer.**
    * We cannot expect the cubic regression to be better, as it *is still a linear model*. Therefore, we do not have enough information to tell which training RSS would be lower. We *could* estimate that by adding the extra cubic term, our least squares fit *could* be a bit closer to the true regression line, and thereby giving the cubic model a lower training RSS, but it may not.
* **(b) Answer (a) using test rather than training RSS.* **
    * We still do not have enough to make confident estimates, but a conservative one would be that the cubic model *could* overfit the data (move towards non-linear fits) and therefore its test RSS would be higher.
* **(c) Suppose the true relationship between X + Y is *not* linear, but we *don't know how far it is from linear*. Consider the training RSS for the linear regression + the training RSS for the cubic regression. Would we expect one to be lower than the other, would we expect them to be the same, or is there not enough information to tell? Justify your answer.* **
    * Even though we do not know how far the relationship is from linear, the fact that we have added in a higher-order polynomial in the cubic regression would give a more flexible model that would better fit a non-linear relationship, thereby giving a lower training RSS
* **(d) Answer (c) using test rather than training RSS.* **
    * We don't have enough info say which test RSS would be lower, since we don't know how far it is from linear. If the true non-linear relationship is *closer to linear than cubic*, linear regression test RSS could be lower, or vice versa (see: Bias-Variance trade-off)

**5. Consider the fitted values that result from performing linear regression *without an intercept*. In this setting, the ith fitted value takes the form y^_i = x_iB, where B = sum_i[x_i*y_i]/sum_k[x_k^2]. Show that we can write y^_i = sum_j[a_j*y_j], and what is a_j?**

  * **Note: We interpret this result by saying that the fitted values from linear regression are *linear combinations* of the response values.**
  * Plugging in sum_i[x_i\*y_i]/sum_k[x_k^2] for B in y^_i = x_iB, we get y^_i = x_i(sum_j[x_j\*y_j]/sum_k[x_k^2]), flipping around terms, we get sum_j[(x_i\*x_j)/sum_k[x_k^2]]\*y_i so a_j = (x_i\*x_j)/sum_k[x_k^2]]

**6. Using (3.4), argue that in the case of simple linear regression, the least squares line always passes through the point (x?, y?).**

  * least square line equation:  y = B0^ + B1^\*x, so if we substitute x_bar for x, we get y = B0^ + B1^\*x_bar
  * (3.4) hows us that the intercept = (0, x_bar - B1\*y_bar)
  * Plug in f(x_bar) into the linear equation which gives us x = x_bar, y = B0 + B1*x_bar
  * Plugging in optimal values of B0 into the equation gives y = (y_bar - B1*x_bar + B1*x_bar, which simplifies to y = y_bar when x = x_bar
  
**7. It is claimed in the text that in the case of simple linear regression of Y onto X, the R2 statistic (3.17) is equal to the square of the correlation between X and Y (3.18). Prove that this is the case. For simplicity, you may assume that ?x = ?y = 0.**
  
  * We know R2 = 1 - RSS/TSS = 1 - (sum_i((y_i - y_i^)^2)).
  * Having y_i^ = Bi^x_i, write R2 = 1 - (sum_i((y_i - sum_j(x_j\*y_j)/sum_j(x_j^2*x_i))^2)/sum_j(y_i^2))

### Applied
8. This question involves the use of simple linear regression on the Auto data set.

**(a) Use the lm() function to perform a simple linear regression with mpg as the response and horsepower as the predictor. Use the summary() function to print the results. Comment on the output. For example:**
```{r}
library(ISLR)
data(Auto)

lm1 <- lm(mpg~horsepower,Auto)
summary(lm1)
```
  * **i. Is there a relationship between the predictor and the response?**
    * The low p-value and \*\*\* significance level indicates that we can reject the null H0: B1 = 0 and say we there is a negative relationship between predictor `horsepower` and response `mpg` 
  * **ii. How strong is the relationship between the predictor and the response?**
      * The low p-value and \*\*\* significance level indicates a relationship, and the moderate R2 value indicated a moderately strong relationship, which we can look at with a scatterplot and correlation calculation
```{r}
library(ggplot2)
ggplot(Auto,aes(horsepower,mpg)) + 
  geom_point(colour="red")
cor(Auto$horsepower,Auto$mpg)
cor(Auto$horsepower,Auto$mpg)**2
```
  * **iii. Is the relationship between the predictor and the response positive or negative?**
    * It's clearly negative
  * **iv. What is the predicted mpg associated with a horsepower of 98? What are the associated 95 % confidence + prediction intervals?**
    * Using our regression formula `39.935861 - 0.157845(98)`, we get:
```{r}
print("predicted mpg for hp = 98")
39.935861 - 0.157845*(98)
print("CI for predicted mpg for hp = 98")
predict(object = lm1, newdata = data.frame(horsepower=98), interval="confidence")
print("PI for predicted mpg for hp = 98")
predict(object = lm1, newdata = data.frame(horsepower=98), interval="prediction")
```
**(b) Plot the response and the predictor. Use the abline() function to display the least squares regression line.**
```{r}
#coef(lm1)
ggplot(Auto,aes(horsepower,mpg)) + 
  geom_point(colour="red") + 
  geom_abline(intercept = 39.9358610, slope = -0.1578447 )
```
**(c) Use the plot() function to produce diagnostic plots of the least squares regression fit. Comment on any problems you see with the fit.**
```{r}
par(mfrow=c(2,2))
plot(lm1)
```
  * We can see a pattern in our residuals in both the residuals and Scale-Location plot, indicating that our residuals are not identically and independently distributed (IID) and our relationship between the predictor and response is non-linear.
  * In our Q-Q plot, we have the fitted quantiles are raising a bit too high at the higher end of theoretical quantiles, indicating a right skew.
  * For our Residuals vs Leverage, we have what looks like 2-possible high levarage points (indicated by annotations of 117 and 94), and some outliers (> 2 or < -2 on the y-axis), which means these points are influencing our regression of `mpg` onto `horsepower`

**9. This question involves the use of multiple linear regression on the Auto data set.**
**(a) Produce a scatterplot matrix which includes all of the variables in the data set.**
```{r}
pairs(~ mpg+displacement+horsepower+weight+acceleration, Auto)
```
**(b) Compute the matrix of correlations between the variables using (exclude `name`, as its qualitative)**
```{r}
cor(dplyr::select(Auto,-name))
```

**(c) Use the lm() function to perform a multiple linear regression with `mpg` as the response and all other variables except name as
the predictors. Use the summary() function to print the results. Comment on the output**
```{R}
lm2 <- lm(mpg~.-name,Auto)
summary(lm2)
```
  * **i. Is there a relationship between the predictors and the response?**
    * There appears to be a relationship between `mpg` and *some* predictors = `weight`,year,origin, and some `displacement`
  * **ii. Which predictors appear to have a statistically significant relationship to the response?**
    * Very sig = `weight`,year,origin, Moderately = `displacement`
  * **iii. What does the coefficient for the year variable suggest?**
    * It suggests that for each year increase, we have a .75 unit increase in `mpg`
* **(d) Use the plot() function to produce diagnostic plots of the linear regression fit. Comment on any problems you see with the fit. Do the residual plots suggest any unusually large outliers? Does the leverage plot identify any observations with unusually high leverage?**
```{r}
par(mfrow=c(2,2))
plot(lm2)
```
  * There may be a *slight* pattern in Residuals vs. Fitted, but it is not apparant in Scale-Location, indicating that we are probably able to say that our residuals are identically and independently distributed (IID) and our relationship between the predictor and response is linear.
  * In our Q-Q plot, we have the fitted quantiles are raising a bit too high at the higher end of theoretical quantiles, indicating a right skew.
  * For our Residuals vs Leverage, we have a couple outliers (> 2 or < -2 on the y-axis) and a high levarage point (14), which means these points are influencing our regression of `mpg` onto `horsepower`

* **(e) Use the \* and : symbols to fit linear regression models with interaction effects. Do any interactions appear to be statistically significant?**
```{r}
lm3 <- lm(mpg~.-name+cylinders*(.-name),Auto)
lm4 <- lm(mpg~.-name+displacement*(.-name),Auto)
lm5 <- lm(mpg~.-name+horsepower*(.-name),Auto)
lm6 <- lm(mpg~.-name+weight*(.-name),Auto)
lm7 <- lm(mpg~.-name+acceleration*(.-name),Auto)
lm8 <- lm(mpg~.-name+year*(.-name),Auto)
lm9 <- lm(mpg~.-name+origin*(.-name),Auto)
summary(lm3)
summary(lm4)
summary(lm5)
summary(lm6)
summary(lm7)
summary(lm8)
summary(lm9)
```
  * cylinders:year  displacement:weight   displacement:year  horsepower:year  weight:year 
  * cylinders:acceleration    weight:acceleration   acceleration:year   acceleration:origin

* **(f) Try a few different transformations of the variables, such as log(X), sqrt(X), X^2. Comment on your findings.**
```{r}
summary(lm(mpg ~ sqrt(displacement) + . - name - cylinders - origin - acceleration, Auto))
summary(lm(mpg ~ log(displacement) + . - name - cylinders - acceleration - origin, Auto))
summary(lm(mpg ~ I(displacement^2) + . - displacement - name - acceleration  - cylinders, Auto)) # now origin matters, lower R2
```

**10. This question should be answered using the Carseats data set.**
**(a) Fit a multiple regression model to predict Sales using Price, Urban, and US.**
```{r}
data("Carseats")
lm10 <- lm(Sales~Price+Urban+US,Carseats)
summary(lm10)
```
**(b) Provide an interpretation of each coefficient in the model. Be careful-some of the variables in the model are qualitative!**
  * If the store is in an urban location, we'd expect the number of sales to decrease by ~22 units, if the store is in the US, we'd expect sales to increase by 1.2K, and for each increase in price by 1 unit, we expect sales to decrease by 5.44 = 5 units.
  * Not, if a store is in the US *and* in an urban location, then we'd expect sales to increase by 1178 units (sum)
**(c) Write out the model in equation form, being careful to handle the qualitative variables properly.**
  * `model = 13.043 - .054\*Price - .024\*Urban + 1.2\*US`
**(d) For which of the predictors can you reject the null hypothesis H0 : Bj = 0?**
  * `Price` and `USyes`
**(e) On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the outcome.**
```{r}
lm11 <- lm(Sales~Price+US,Carseats)
summary(lm11)
```
**(f) How well do the models in (a) and (e) fit the data?**
  * 2nd model fits barely better (`R2 = .2354` vs. `.2335`)
**(g) Using the model from (e), obtain 95 % confidence intervals for the coefficient(s).**
```{r}
confint(lm11)
```
**(h) Is there evidence of outliers or high leverage observations in the model from (e)?**
```{r}
par(mfrow=c(2,2))
plot(lm11)
```
  * 1 high leverage point

**11. In this problem we will investigate the t-statistic for the null hypothesis H0: B = 0 in simple linear regression without an intercept. To begin, generate a predictor x and a response y as follows.**
```{r}
set.seed(1)
x <- rnorm(100)
y <- 2*x+rnorm(100)
```
**(a) Perform a simple linear regression of y onto x, w/out an intercept. Report coefficient estimate B^, the Std. Error of this coefficient estimate, + the t-statistic + p-value associated w/ the null. Comment on these results. **
```{r}
# regression w/out intercept
summary(lm(y~x+0))
```
  * We have `B^ = 1.9939`, `Std. Error = .1065`, `t = 18.73` (large = good for deviation), and `p-value = 2e-16` (very significant)
  * We have a p-value (probability of observing any value equal>= |t| larger in absolute value, assuming B = 0) that is very low, which means that we can reject the null that there is no relationship between predictor + response (reject that H0: B = 0) and the difference are not due to chance.
**(b) Now perform a simple linear regression of x onto y without an intercept, and report the coefficient estimate, its standard error, and the corresponding t-statistic and p-values associated with the null hypothesis H0 : ?? = 0. Comment on these results.**
```{r}
summary(lm(x~y+0))
```
  * We have `B^ = .3911`, `Std. Error = 02089`, `t = 18.73` (same as above), and `p-value = 2e-16` (very significant)
  * We have a p-value (probability of observing any value equal>= |t| larger in absolute value, assuming B = 0) that is very low, which means that we can reject the null that there is no relationship between predictor + response (reject that H0: B = 0) and the difference are not due to chance.
**(c) What is the relationship between the results obtained in (a) and (b)?**
 * While we have different coefficient estimates, we've obtained the same t-stastic and same p-value, meaning that the evidence to reject the null was the same in both models
**(d) For the regression of Y onto X without an intercept, the t-statistic for H0 : B = 0 takes the form B^ / SE(B^), where B^ = given by (3.38), and where Std. Error have its own formula. (slightly different from Sections 3.1.1 + 3.1.2, since here we're performing regression w/out an intercept.) Show algebraically, and+ confirm numerically in R, that the t-statistic can be written as**
```{r}
b_hat_x <- sum(x*y)/(sum(x^2))
(stderr_b_hat_x <- sqrt(sum((y-x*b_hat_x)^2)/((length(x)-1)*sum(x^2))))

b_hat_y <- sum(y*x)/(sum(y^2))
(stderr_b_hat_y <- sqrt(sum((x-y*b_hat_y)^2)/((length(y)-1)*sum(y^2))))
```
**(e) Using the results from (d), argue that the t-statistic for the regression of y onto x is the same as the t-statistic for the regression of x onto y.**
  * Since `t = B / Std. Error`, and we have different B values for each, we can use the different values for SE(B) to get the same t-statistic
```{r}
(t_x <- b_hat_x/stderr_b_hat_x)
(t_y <- b_hat_y/stderr_b_hat_y)
```
**(f) In R, show that when regression is performed with an intercept, the t-statistic for H0 : B1 = 0 is the same for the regression of y onto x as it is for the regression of x onto y.**
```{r}
x_on_y <- lm(x~y)
summary(x_on_y)
x_on_y$coefficients[1]/.04266
x_on_y$coefficients[2]/.02099
```

```{r}
y_on_x <- lm(y~x)
summary(y_on_x)
y_on_x$coefficients[1]/.09699
y_on_x$coefficients[2]/.10773
```

**12. This problem involves simple linear regression without an intercept.**
**(a) Recall the coefficient estimate B^ for linear regression of Y onto X w/out an intercept is given by (3.38). Under what circumstance is the coefficient estimate for the regression of X onto Y the same as the coefficient estimate for the regression of Y onto X?**
  * Looking @ the formulas for both B^ values, the numerator is the same `(sum_i(x_i*y_i))`, but with different denominators, `sum_i(y_i^2)` vs `sum_i(x_i^2)`. So, when those 2 values are equal, then we get the same B^ for each regression.
**(b) Generate an example in R with n = 100 observations in which the coefficient estimate for the regression of X onto Y is *different* from the coefficient estimate for the regression of Y onto X.**
```{r}
set.seed(1)
x <- rnorm(100)
y <- rnorm(100)
summary(lm(y~x+0))$coefficients
summary(lm(x~y+0))$coefficients
```
**(c) Generate an example in R with n = 100 observations in which the coefficient estimate for the regression of X onto Y is the same as the coefficient estimate for the regression of Y onto X.**
```{r}
set.seed(1)
x <- rnorm(100)
y <- rev(x)
summary(lm(y~x+0))$coefficients
summary(lm(x~y+0))$coefficients
```

**13. In this exercise you will create some simulated data + will fit simple linear regression models to it. Make sure to use set.seed(1) prior to starting part (a) to ensure consistent results.**
**(a) Using the rnorm() function, create a vector, x, containing 100 observations drawn from a N(0, 1) distribution. This represents a feature, X.**
```{r}
set.seed(1)
x <- rnorm(100) # N(0,1) by default
```
**(b) Using the rnorm() function, create a vector, eps, containing 100 observations drawn from a N(0, 0.25) distribution i.e. a normal distribution with mean zero and variance 0.25.**
```{r}
eps <- rnorm(100,sd=sqrt(.25))
```
**(c) Using x and eps, generate a vector y according to the model `Y = -1 +0.5X + eps. (3.39) What is the length of the vector y? What are the values of B0 and B1 in this linear model?**
```{r}
y <- -1 + .5*x + eps
head(y)
length(y) 
```
  * We are given the values for B0 = -1 and B1 = .5 
**(d) Create a scatterplot displaying the relationship between x and y. Comment on what you observe.**
```{r}
ggplot(aes(x,y),data=data.frame(x,y)) + 
  geom_point(color="red") + 
  theme_bw()
#plot(x,y)
```
  * We have a mostly linear model, with some noise introduced by our `eps` vector of irreducible error.
**(e) Fit a least squares linear model to predict y using x. Comment on the model obtained. How do B^0 and B^1 compare to B0 and B1?**
```{r}
lm_eps <- lm(y~x)
summary(lm_eps)
```
  * They are virtually identical, with a lower B^0 and *slightly* lower B^1, which can be rounded to B1
**(f) Display the least squares line on the scatterplot obtained in (d). Draw the population regression line on the plot, in a different color. Use the legend() command to create an appropriate legend.**
```{r}
ggplot(aes(x,y),data=data.frame(x,y)) + 
  geom_point(color="red") + 
  theme_bw() + 
  geom_abline(intercept = -1.01885, slope = 0.49947)
```
**(g) Now fit a polynomial regression model that predicts y using x and x2. Is there evidence that the quadratic term improves the model fit? Explain your answer.**
```{r}
lm_poly <- lm(y~poly(x,2))
summary(lm_poly)
```
  * R2 has increased (technically), but the second-order polynomial is not significant, meaning there's no evidence that the quadratic term improves model fit.
  
**(h) Repeat (a)-(f) w/ less noise in the data (reduce variance in epsilon generation). The model (3.39) should remain the same. Describe your results.**
```{r}
set.seed(1)
x1 <- rnorm(100) # N(0,1) by default
eps1 <- rnorm(100,sd=sqrt(.05))
y1 <- -1 + .5*x1 + eps1
lm_eps1 <- lm(y1~x1)
summary(lm_eps1)

ggplot(aes(x,y),data=data.frame(x,y)) + 
  geom_point(color="red") + 
  theme_bw() + 
  geom_abline(intercept = -1.01885, slope = 0.49947)
ggplot(aes(x1,y1),data=data.frame(x1,y1)) + 
  geom_point(color="red") + 
  theme_bw() + 
  geom_abline(intercept = -1.00843, slope = 0.49976)
```
**(i) Repeat (a)-(f) after modifying the data generation process in such a way that there is more noise in the data. The model (3.39) should remain the same. You can do this by increasing the variance of the normal distribution used to generate the error term  in (b). Describe your results.**
```{r}
set.seed(1)
x2 <- rnorm(100) # N(0,1) by default
eps2 <- rnorm(100,sd=sqrt(.5))
y2 <- -1 + .5*x2 + eps2
lm_eps2 <- lm(y2~x2)
summary(lm_eps2)

par(mfrow=c(2,2))
ggplot(aes(x,y),data=data.frame(x,y)) + 
  geom_point(color="red") + 
  theme_bw() + 
  geom_abline(intercept = -1.01885, slope = 0.49947)
ggplot(aes(x1,y1),data=data.frame(x1,y1)) + 
  geom_point(color="red") + 
  theme_bw() + 
  geom_abline(intercept = -1.00843, slope = 0.49976)
ggplot(aes(x2,y2),data=data.frame(x2,y2)) + 
  geom_point(color="red") + 
  theme_bw() + 
  geom_abline(intercept = -1.02665, slope = 0.49925)
```
**(j) What are the confidence intervals for B0 and B1 based on the original, noisier, + less noisy data sets? Comment on your results.**
```{r}
confint(lm_eps)
confint(lm_eps1)
confint(lm_eps2)
```

**14. This problem focuses on the collinearity problem.**
**(a) Perform the following commands in R:**
```{r}
set.seed(1)
x1 <- runif(100)
x2 <- 0.5*x1 + rnorm(100)/10
y <- 2 + 2*x1 + 0.3*x2 + rnorm(100)
```
**The last line corresponds to creating a linear model in which y = a function of x1 + x2. Write out the form of the linear model. What are the regression coefficients?**
  * Model = `y = B0 + B1x1 + B2x2 + eps` where B0 = B1 = 2, B2 = .3, and epsilon is 100 random numbers between 0-1 from the normal distribution
**(b) What is the correlation between x1 and x2? Create a scatterplot displaying the relationship between the variables.**
```{r}
cor(x1,x2)
ggplot(aes(x1,x2),data=data.frame(x1,x2)) + 
  geom_point(color="red") + 
  theme_bw() 
```
  * There's a pretty strong positive correlation between both predictors
**(c) Using this data, fit a least squares regression to predict y using x1 + x2. Describe the results obtained. What are B^0, B^1, and B^2? How do these relate to the true B0, B1, and B2? Can you reject the null H0: B1 = 0? How about the null hypothesis H0: B2 = 0?**
```{r}
lm_14 <- lm(y~x1+x2)
summary(lm_14)
```
  * Our B^0 has slightly increased from 2 to 2.1305, while our B^1 has decreased by over .5 from 2 down to 1.4396, while B^3 has increased from .3 to 1.0097. We cannot reject the null H0: B2 = 0, but we *may* be able to reject the null H0: B1 = 0, depending on our threshold for doing so (is 1 * enough?)
**(d) Now fit a least squares regression to predict y using only x1. Comment on your results. Can you reject the null hypothesis H0 : B1 = 0?**
```{r}
lm_14_2 <- lm(y~x1)
summary(lm_14_2)
```
**(e) Now fit a least squares regression to predict y using only x2. Comment on your results. Can you reject the null H0 : B1 = 0?**
  * It seems that now we can reject the null H0 : B1 = 0
```{r}
lm_14_3 <- lm(y~x2)
summary(lm_14_3)
```
  * It seems that now we can reject the null H0 : B1 = 0
**(f) Do the results obtained in (c)-(e) contradict each other? Explain your answer.**
  * No, each predictor can only influence the response in isolation, with no other effects from the other predictors.
  * **We know this because r_x1x2 = .835, i.e .x1 and x2 are collinear**
  * Therefore, we cannot (or it is difficult to) distinguish the individual effects of each predictor in isolation that we can see in the simple linear models with 1 predictor
  
**(g) Now suppose we obtain one additional observation, which was unfortunately mismeasured.**
```{r}
x1 <- c(x1, 0.1)
x2 <- c(x2, 0.8)
y <- c(y,6)
```
**Re-fit the linear models from (c) to (e) using this new data. What effect does this new observation have on the each of the models? In each model, is this observation an outlier? A high-leverage point? Both? Explain your answers.**
```{r}
lm_14 <- lm(y~x1+x2)
summary(lm_14)
par(mfrow=c(2,2))
plot(lm_14,main="both")

lm_14_2 <- lm(y~x1)
summary(lm_14_2)
par(mfrow=c(2,2))
plot(lm_14_2,main="x1")

lm_14_3 <- lm(y~x2)
summary(lm_14_3)
par(mfrow=c(2,2))
plot(lm_14_3,main="x2")
```
  * It now appears that we can see some individual effect from x2 in the multiple linear regression model (as opposed to x1 in the original multiple regression), while both individual effects are still clearly visible in the simple models.
  * R2 has increased in the multiple model and the x2 model, but decreased in the x1 model
  * Looking at the x1 plots, we see that observation 101 is indeed an outlier (Residuals vs. leverage), and that it's fitted quantiles is "too large", indicating an overestimation.
  * We can also see it clearly annotatedin the residuals vs. fitted and scale-location plots as an outlier (>2 or <-2)
  * It is a high-leverage point for the x2 model, as well as in the multiple regression.

**15. Using Boston data set, try to predict per capita crime rate using the other variables in this data set. In other words, per capita crime rate is the response, and the other variables are the predictors.**
**(a) For each predictor, fit a simple linear regression model to predict the response. Describe your results. In which of the models is there a statistically significant association between the predictor and the response? Create some plots to back up your assertions.**
```{r}
library(broom)
library(MASS)
names <- names(Boston)
names <- names[2:length(names)]

df_singles <- as.data.frame(data_frame(formula = paste('crim ~', names), 
           model = map(formula, lm, data = Boston))) %>%
  mutate(tidied = map(model, tidy)) %>%
  unnest(tidied) %>%
  mutate(significant = if_else(p.value < .001,"***",
                                 if_else(p.value < .01,"**",
                                         if_else(p.value < .05,"*","no"))))
(df_singles <- df_singles %>% filter(!term=="(Intercept)"))
```
  * Each variable, in isolation, has a statistically significant association with the response `crim` *except* for being on the Charles River (`chas`).
**(b) Fit a multiple regression model to predict the response using all of the predictors. Describe your results. For which predictors can we reject the null hypothesis H0: Bj = 0?**
```{r}
fullBostonCrim <- lm(crim~.,Boston)
summary(fullBostonCrim)
```
  * The variables for which we can reject the null H0: Bj = 0 are `rad`, `dis`, less so `medv`, and less so `zn`, `black`. So, looking at the coefficients, we can see that an increase of 1 unit in the index of accessibility to radial highways would increase per capita crime rate by .588 units, an increase of 1 unit in the weighted mean of distances to five Boston employment centres would decrease per capita crime rate by .987 units, and an increase in in median home value of $1K would mean a decrease of .198 units for crime rate.
**(c) How do results from (a) compare to results from (b)? Create a plot displaying univariate regression coefficients from (a) on x-axis, + multiple regression coefficients from (b) on y-axis (i.e. each predictor is displayed as a single DP**
  * The multiple linear regression has eliminated signicant associations indicated by the simple linear regression models for indus, nox, rm, age, tax, ptratio, and lstat, and decreased the significance of zn, black, and medv, and for dis and rad (but these last 3 ar still significant)
```{r}
simpEstimate <- fullBostonCrim$coefficients[2:length(fullBostonCrim$coefficients)]
coeffs <- df_singles %>%
  filter(!term=="(Intercept)") %>% 
  dplyr::select(term,estimate) %>% 
  mutate(multEstimate = estimate
         ,simpEstimate = simpEstimate) %>%
  dplyr::select(-estimate)

ggplot(coeffs,aes(simpEstimate,multEstimate)) + 
  geom_point()
```
**(d) Is there evidence of non-linear association between any of the predictors + the response? To answer this question, for each predictor X, fit a model of the form `Y = B0 + B1X + B2X^2 + B3X^3 + epsilon`**
```{r}
df <- as.data.frame(data_frame(formula = paste('crim ~ ', names, " + I(", names, "^2)", " + I(", names, "^3)", sep=""),
                         model = map(formula, lm, data = Boston))) %>%
  mutate(tidied = map(model, tidy)) %>%
  unnest(tidied) %>%
  mutate(significant = if_else(p.value < .001,"***",
                                 if_else(p.value < .01,"**",
                                         if_else(p.value < .05,"*","no"))))
df %>% filter(!term=="(Intercept)")
```
* It appears that indus, nox, dis, medv, less so, ptratio, and (barely) age have a non-linear relationship with crim