---
title: "ISLR_Ch3_LinearRegression"
author: "Steve Newns"
date: "April 20, 2018"
output: html_document
---
# 3.6.2 Simple Linear Regression
```{r}
library(MASS)
library(ISLR)
```

`MASS` library contains `Boston` data set, which records `medv` (median house value) for 506 neighborhoods around Boston. We will **seek to predict `medv` using 13 predictors such as `rm` (avg #  of rooms per house), `age` (avg age of houses), + `lstat` (% of households with low socioeconomic status)**

```{r}
#fix(Boston)
names(Boston)
```

We will start by using `lm()` to fit a simple linear regression model, with `medv` as response + `lstat` as predictor. The basic
```{r}
simpleLM <- lm(medv ~ lstat, data = Boston)
```

Typing `simpleLM` gives some basic info about the model, but  `summary(simpleLM)` gives more detailed information like p-values + standard errors for the coefficients, as well as the R2 statistic and F-score
```{r}
simpleLM
summary(simpleLM)
```

Can use `names()` to find out what other pieces of info are stored in the fit model. 
```{r}
names(simpleLM)
```

Can extract quantities by name (`lm.fit$coefficients`) but it is safer to use **extractor functions** like `coef()` to access them.
```{r}
coef(simpleLM)
```

To obtain a CI for the coefficient estimates, use `confint()`
```{r}
confint(simpleLM)
```

`predict()` can be used to produce CI's *and* **prediction intervals** for the prediction of `medv` for a given value of `lstat`.

  * **CI's** tell you about how well you have determined the parameter of interest
    * Repeat an experiment many times + calculate a CI from each sample, we'd expect ~95% of those intervals to include the true value of the population parameter (have 95% confidence the interval contains the true parameter value)
    * Key point = CI's tell you about the likely location of the true population parameter.

  * **Prediction intervals** tell you *where you can expect to see the next DP sampled*. 
    * Assume the data really are randomly sampled from a Gaussian distribution. 
    * Collect a sample of data + calculate a PI
    * Then sample 1 more value from the population
    * Repeat many times + expect that next value to lie within such a prediction interval in 95% of samples.
    * Key point = A PI tells you *about the distribution of values*, NOT the uncertainty in determining the true population parameter
    * PI's must account for both the **uncertainty** in knowing the value of the true population parameter, plus **scatter/noise** So
    * **A PI is always wider than a confidence interval (less confident)**
    
  * CI's give a range for `E[y???x]`, PI's give a range *for `y` itself*.
  * Naturally, best guess for y = `E[y???x]`, so both intervals are centered around the same value, `x??^`, but with *different standard errors*
  * We *guess** the expected value of `E[y???x]` more precisely than we estimate `y` itself** so **estimating y requires including the variance that comes from the true error term**
    * Ex: Could get perfect estimates of ?? (coefficients), so an estimate of `E[y???x]` would be perfect, but we still wouldn't be sure what `y` itself was b/c *there is a true error term we need to consider (irreducible)*. 
    * Our confidence "interval" would just be a single DP, as we estimate E[y???x] exactly right, but our PI would be wider b/c we take the true error term into account.
```{r}
# create a column (vector) of 5,10,15 as new lstat values to use to predict medv's valie + CI with the model
print("CI")
predict(object = simpleLM, newdata = data.frame(lstat=c(5,10,15)), interval="confidence")
print("PI")
predict(object = simpleLM, newdata = data.frame(lstat=c(5,10,15)), interval="prediction")
```

95% CI associated with a `lstat` value = 10 is `(24.47, 25.63)`, and the 95 % **prediction interval** = `(12.828, 37.28)`. As expected, the *CI and PI's are centered around the same point* = a predicted value of `25.05` for `medv` when `lstat` = 10, but
the latter are substantially wider.

Now plot `medv` + `lstat` along w/ the **least squares regression line** using `plot()` and `abline()`
```{r}
plot(Boston$lstat,Boston$medv)
abline(simpleLM, col="red", lwd=3)
```

There is some evidence for **non-linearity** in this relationship, as seen by the upper-left area of the plot

`abline()` can be used to draw *any* line, not just the least squares regression line. To draw a line with intercept `a` and slope `b`, type `abline(a,b)`. 

```{r}
par(mfrow=c(2,3))
plot(Boston$lstat,Boston$medv)
abline(simpleLM, lwd=3)
plot(Boston$lstat,Boston$medv)
abline(simpleLM, lwd=3, col="red")
plot(Boston$lstat, Boston$medv, col="red")
plot(Boston$lstat, Boston$medv, pch=20) # change DP shape
plot(Boston$lstat, Boston$medv, pch="+") # change DP shape
plot(1:20,1:20,pch =1:20)
```

Next we examine some **diagnostic plots**. 4 are automatically produced by applying `plot()` directly to the output from a linear model.
```{r}
par(mfrow=c(2,2)) # #see all plots together
plot(simpleLM) # residuals v fitted, QQ, Scale-Location, residuals v leverage
```

There looks to be a curving pattern as we move down the x-axis to the right,almost like a fish hook, indicating non-linearity between `lstat` and `medv` such that our residuals are not **identically and independently distributed (IID)**

**Normal Q-Q plot** = empirical quantiles vs. theoretical Standard Normal quantiles. For our model, our fitted quantiles are "too large", indicating a slight-right skew.

  * quantiles from a N(??,??2) population should be linearly related to Standard Normal quantiles
    * Let xp = pth quantile from a N(??,??2), then probability of *non-standard normal* X being below a real value x is P(X <= x) = p, so clearly, for the *standardized* normal Z, P(Z <= (xp - ??)/??)) = p, and therefore xp = ?? + zp*??

**"Residuals vs Fitted" and "Scale/Spread-Location"** = essentially the same, showing if there is a trend to the residuals (if they're spread equally along the ranges of predictors). This again shows the fish-hook pattern indicating non-IID residuals and non-linearity in the relationship.

**Residuals vs. Leverage** = tells us which DP's have the greatest influence on the regression (i.e. high leverage points).

  * Not all outliers are influential in linear regression so results wouldn't be much different whether included or not.
  * Watch for outlying values at the upper right or lower right corners = spots where cases can be influential against a regression line. 
  * Look for cases outside of a **Cook's distance** (dashed line) , meaning they have high **Cook's distance scores** + are influential to the regression results, as the results will be altered if we exclude those cases.
    * Cook's distance = function of the leverage and standardized residual associated with each data poin
  * Looks' like we have several high leverage points that influence our regression as noted by being quite far from the dotted Cook's distance line and being noted with annotations (215,413,375) 

Alternatively, can compute residuals from a linear regression fit using `residuals()`, or `rstudent()` to return **studentized residuals**, and we can use this to plot the residuals against the fitted values (1st + 3rd plots above)

  * **studentized residuals** = alternative criterion for IDing outliers
    * basic idea = delete observations ` at a time, each time refitting the regression model on the remaining `n-1` observations
    * Then, compare the observed response values to their fitted values based on the models with the `i`th observation deleted. 
    * This produces **deleted residuals**, and standardizing the deleted residuals produces studentized residuals.
    *  deleted residual `d_i = y_i - y_i_hat` = observed response for ith variable minus predicted response based on the estimated model with the ith observation deleted
    
      * data point i being influential implies the DP "pulls" the estimated regression line towards itself, which means the observed response would be close to the predicted response. 
      * But, if removed, the estimated regression line would "bounce back" away from the observed response, thereby **resulting in a large deleted residual** (i.e. a DP w/ a large deleted residual suggests the DP is influential)
      
    * "Largeness" of deleted residuals depend on units of measurement, just as ordinary residuals do
    * Solve this by **standardizing** = dividing each deleted residual by an estimate of its standard deviation.
    * studentized residual, `t_i = d_i/sd(d_i) = e_i/sqrt(MSE_i(1-h_ii)` = same as ordinary residual `e_i` divided by a factor that includes the MSE *based on estimated model **w/ ith observation deleted**, `MSE_i`*, and the **leverage, `h_ii`**
    * TLDR: studentized residual = just a deleted residual divided by its estimated standard deviation
      
      * In general, studentized residuals = more effective than standardized residuals. 
      * If an observation has a studentized residual > 3 (in absolute value), can call it an outlier.
```{r}
par(mfrow=c(2,1))
plot(x=predict(simpleLM),y=residuals(simpleLM))
plot(x=predict(simpleLM),y=rstudent(simpleLM))
```

On the basis of the residual plots, there is some evidence of non-linearity. **Leverage statistics** can be computed for any # of predictors using `hatvalues()`.
```{r}
#hatvalues(simpleLM)
plot(hatvalues(simpleLM))
which.max(hatvalues(simpleLM))
```
We see the index of the observation has the largest leverage statistic as 365

### 3.6.3 Multiple Linear Regression

To fit a multiple linear regression model using **least squares**, again use `lm()` w/ `n` predictors
```{r}
multLR <- lm(medv ~ lstat + age, Boston)
summary(multLR)
```

```{r}
allMultLR <- lm(medv ~ ., Boston)
summary(allMultLR)
```

Can access individual components of a **summary object** by name (`?summary.lm`). Hence `summary(lm.fit)$r.sq` gives us R2 + `summary(lm.fit)$sigma` gives us the **RSE**. `vif()`, part of the `car` package (functions to apply to a fitted regression model) can be used to compute **variance inflation** factors. Most VIF's are low to moderate for this data.

* **VIF** = ratio of the variance of ??_j when fitting the full model divided by the variance of ??_j if fit on its own
  * smallest possible = 1 (indicates complete absence of collinearity)
  * Typically in practice, there's a small amount of collinearity among predictors
  * Rule of thumb = VIF > 5 or 10 indicates a problematic amount of collinearity. 
  * `VIF(??_j ) = 1/(1 ??? R2_(X_j | X_-j)` where `R2_(X_j | X_-j)` = R2 from a regression of `X_j` onto all other predictors, `X_-j`)
  * If `R2_(X_j | X_-j)` = close to 1, collinearity is present, so the VIF will be large.
```{R}
#install.packages("car",dependencies=TRUE)
#update.packages(checkBuilt = TRUE)
library(car)
vif(allMultLR)
```

What if we'd like to perform a regression using all variables but one?
```{r}
multLR_noAge <- lm(medv ~ . - age, Boston) # largest p-value in full model
summary(multLR_noAge)
```

Alternatively, `update()` can be used.
```{r}
multLR_noAge <- update(allMultLR, ~ . - age)
summary(multLR_noAge)
```

### 3.6.4 Interaction Terms
It is easy to include **interaction terms** in a linear model, such as w/ syntax `lstat:black` = interaction term between `lstat` +  `black`. The syntax `lstat*age` simultaneously includes `lstat`, `age`, *and the interaction term `lstat × age`* as predictors, as it is a shorthand for `lstat+age+lstat:age`
```{r}
multLR_interaction <- lm(medv ~ lstat*age, Boston)
summary(multLR_interaction)
```

### 3.6.5 Non-linear Transformations of Predictors

`lm()` function can also accommodate non-linear transformations of predictors. Given a predictor `X`, can create a predictor `X2`
using `I(X^2)` (`I()` is needed since `^` has a special meaning in a formula)
```{r}
multLR_transform <- lm(medv ~ lstat + I(lstat^2), Boston)
summary(multLR_transform)
```

The near-0 p-value associated with the quadratic term suggests it leads to an improved model. **Use `?anova()` to further quantify the extent to which the quadratic fit is superior to the linear fit.**

* `anova()` = computes analysis of variance (or deviance) tables for 1+ fitted model objects, tested in order specified whether the model terms are significant
```{r}
anova(lm(medv~lstat,Boston),multLR_transform)
```

`anova()` performs a **hypothesis test** comparing the 2 models w/ a null H0 = the 2 models fit the data equally well, + the alternative H1 = the full model (1st given) is superior. 

Here the **F-statistic** = 135 + the associated p-value = virtually 0. This info provides very clear evidence that the quadratic model is far superior (not surprising, since earlier we saw evidence for non-linearity in the relationship between `medv` + `lstat`
```{r}
par(mfrow=c(2,2))
plot(multLR_transform)
```

There looks to be no pattern for our normal nor studentizes residuals, indicating linearity in our fit model such that our residuals *are* identically and independently distributed (IID)

**Normal Q-Q plot** = For our new model, our fitted quantiles are still "too large" in the upper-right, indicating a slight-right skew.

**Residuals vs. Leverage** = Looks like we still have several high leverage points that influence our regression as noted by being quite far from the dotted Cook's distance line and being noted with annotations

In order to create a cubic fit, we can include a predictor of the form `I(X^3)`. However, this approach can start to get cumbersome for higher-order polynomials. A better approach involves using `poly()` to create the polynomial within `lm()`
```{r}
# produce a fifth-order polynomial fit
fifthOrder <- lm(medv ~ poly(lstat,5), Boston)
summary(fifthOrder)
```

This suggests that including additional polynomial terms, up to 5th order, leads to an improvement in the model fit! However, further investigation of data reveals that no polynomial terms beyond 5th order have significant p-values in a regression fit. Of course, we are in no way restricted to only using polynomial transformations of the predictors. Here we try a **log transformation.**
```{r}
summary(lm(medv ~ log(rm), Boston))
```
         
### 3.6.6 Qualitative Predictors
We will now examine the Carseats data, which is part of the ISLR library.
We will attempt to predict Sales (child car seat sales) in 400 locations
based on a number of predictors.
> fix(Carseats )
> names(Carseats )
[1] "Sales" "CompPrice " "Income" "Advertising "
[5] "Population " "Price" "ShelveLoc " "Age"
[9] "Education " "Urban" "US"
The Carseats data includes qualitative predictors such as Shelveloc, an indicator
of the quality of the shelving location-that is, the space within
a store in which the car seat is displayed-at each location. The predictor
Shelveloc takes on three possible values, Bad, Medium, and Good.
118 3. Linear Regression
Given a qualitative variable such as Shelveloc, R generates dummy variables
automatically. Below we fit a multiple regression model that includes some
interaction terms.
> lm.fit=lm(Sales???.+Income :Advertising +Price:Age ,data=Carseats )
> summary (lm.fit)
Call:
lm(formula = Sales ??? . + Income:Advertising + Price:Age , data =
Carseats )
Residuals :
Min 1Q Median 3Q Max
-2.921 -0.750 0.018 0.675 3.341
Coefficients:
Estimate Std. Error t value Pr(>|t|)
(Intercept ) 6.575565 1.008747 6.52 2.2e-10 ***
CompPrice 0.092937 0.004118 22.57 < 2e-16 ***
Income 0.010894 0.002604 4.18 3.6e-05 ***
Advertising 0.070246 0.022609 3.11 0.00203 **
Population 0.000159 0.000368 0.43 0.66533
Price -0.100806 0.007440 -13.55 < 2e-16 ***
ShelveLocGood 4.848676 0.152838 31.72 < 2e-16 ***
ShelveLocMedium 1.953262 0.125768 15.53 < 2e-16 ***
Age -0.057947 0.015951 -3.63 0.00032 ***
Education -0.020852 0.019613 -1.06 0.28836
UrbanYes 0.140160 0.112402 1.25 0.21317
USYes -0.157557 0.148923 -1.06 0.29073
Income:Advertising 0.000751 0.000278 2.70 0.00729 **
Price:Age 0.000107 0.000133 0.80 0.42381
---
Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
Residual standard error: 1.01 on 386 degrees of freedom
Multiple R-squared: 0.876, Adjusted R-squared: 0.872
F-statistic : 210 on 13 and 386 DF, p-value: <2e-16
The contrasts() function returns the coding that R uses for the dummy contrasts() variables.
> attach(Carseats )
> contrasts (ShelveLoc )
Good Medium
Bad 0 0
Good 1 0
Medium 0 1
Use ?contrasts to learn about other contrasts, and how to set them.
R has created a ShelveLocGood dummy variable that takes on a value of
1 if the shelving location is good, and 0 otherwise. It has also created a
ShelveLocMedium dummy variable that equals 1 if the shelving location is
medium, and 0 otherwise. A bad shelving location corresponds to a zero
for each of the two dummy variables. The fact that the coefficient for
3.6 Lab: Linear Regression 119
ShelveLocGood in the regression output is positive indicates that a good
shelving location is associated with high sales (relative to a bad location).
And ShelveLocMedium has a smaller positive coefficient, indicating that a
medium shelving location leads to higher sales than a bad shelving location
but lower sales than a good shelving location.
3.6.7 Writing Functions
As we have seen, R comes with many useful functions, and still more functions
are available by way of R libraries. However, we will often be interested
in performing an operation for which no function is available. In this
setting, we may want to write our own function. For instance, below we
provide a simple function that reads in the ISLR and MASS libraries, called
LoadLibraries(). Before we have created the function, R returns an error if
we try to call it.
> LoadLibraries
Error: object 'LoadLibraries ' not found
> LoadLibraries()
Error: could not find function " LoadLibraries"
We now create the function. Note that the + symbols are printed by R and
should not be typed in. The { symbol informs R that multiple commands
are about to be input. Hitting Enter after typing { will cause R to print the
+ symbol. We can then input as many commands as we wish, hitting Enter
after each one. Finally the } symbol informs R that no further commands
will be entered.
> LoadLibraries= function (){
+ library(ISLR)
+ library(MASS)
+ print("The libraries have been loaded .")
+ }
Now if we type in LoadLibraries, R will tell us what is in the function.
> LoadLibraries
function (){
library(ISLR)
library(MASS)
print("The libraries have been loaded .")
}
If we call the function, the libraries are loaded in and the print statement
is output.
> LoadLibraries()
[1] "The libraries have been loaded ."
120 3. Linear Regression
3.7 Exercises
Conceptual
1. Describe the null hypotheses to which the p-values given in Table 3.4
correspond. Explain what conclusions you can draw based on these
p-values. Your explanation should be phrased in terms of sales, TV,
radio, and newspaper, rather than in terms of the coefficients of the
linear model.
2. Carefully explain the differences between the KNN classifier and KNN
regression methods.
3. Suppose we have a data set with five predictors, X1 = GPA, X2 = IQ,
X3 = Gender (1 for Female and 0 for Male), X4 = Interaction between
GPA and IQ, and X5 = Interaction between GPA and Gender. The
response is starting salary after graduation (in thousands of dollars).
Suppose we use least squares to fit the model, and get ??^0 = 50, ??^1 =
20, ??^2 = 0.07, ??^3 = 35, ??^4 = 0.01, ??^5 = ???10.
(a) Which answer is correct, and why?
i. For a fixed value of IQ and GPA, males earn more on average
than females.
ii. For a fixed value of IQ and GPA, females earn more on
average than males.
iii. For a fixed value of IQ and GPA, males earn more on average
than females provided that the GPA is high enough.
iv. For a fixed value of IQ and GPA, females earn more on
average than males provided that the GPA is high enough.
(b) Predict the salary of a female with IQ of 110 and a GPA of 4.0.
(c) True or false: Since the coefficient for the GPA/IQ interaction
term is very small, there is very little evidence of an interaction
effect. Justify your answer.
4. I collect a set of data (n = 100 observations) containing a single
predictor and a quantitative response. I then fit a linear regression
model to the data, as well as a separate cubic regression, i.e. Y =
??0 + ??1X + ??2X2 + ??3X3 + .
(a) Suppose that the true relationship between X and Y is linear,
i.e. Y = ??0 + ??1X + . Consider the training residual sum of
squares (RSS) for the linear regression, and also the training
RSS for the cubic regression. Would we expect one to be lower
than the other, would we expect them to be the same, or is there
not enough information to tell? Justify your answer.
3.7 Exercises 121
(b) Answer (a) using test rather than training RSS.
(c) Suppose that the true relationship between X and Y is not linear,
but we don't know how far it is from linear. Consider the training
RSS for the linear regression, and also the training RSS for the
cubic regression. Would we expect one to be lower than the
other, would we expect them to be the same, or is there not
enough information to tell? Justify your answer.
(d) Answer (c) using test rather than training RSS.
5. Consider the fitted values that result from performing linear regression
without an intercept. In this setting, the ith fitted value takes
the form
y^i = xi??, ^
where
??^ =
n
i=1
xiyi

/
n
i=1
x2
i

. (3.38)
Show that we can write
y^i = n
i=1
aiyi .
What is ai?
Note: We interpret this result by saying that the fitted values from
linear regression are linear combinations of the response values.
6. Using (3.4), argue that in the case of simple linear regression, the
least squares line always passes through the point (¯x, y¯).
7. It is claimed in the text that in the case of simple linear regression
of Y onto X, the R2 statistic (3.17) is equal to the square of the
correlation between X and Y (3.18). Prove that this is the case. For
simplicity, you may assume that ¯x = ¯y = 0.
Applied
8. This question involves the use of simple linear regression on the Auto
data set.
(a) Use the lm() function to perform a simple linear regression with
mpg as the response and horsepower as the predictor. Use the
summary() function to print the results. Comment on the output.
For example:
122 3. Linear Regression
i. Is there a relationship between the predictor and the response?
ii. How strong is the relationship between the predictor and
the response?
iii. Is the relationship between the predictor and the response
positive or negative?
iv. What is the predicted mpg associated with a horsepower of
98? What are the associated 95 % confidence and prediction
intervals?
(b) Plot the response and the predictor. Use the abline() function
to display the least squares regression line.
(c) Use the plot() function to produce diagnostic plots of the least
squares regression fit. Comment on any problems you see with
the fit.
9. This question involves the use of multiple linear regression on the
Auto data set.
(a) Produce a scatterplot matrix which includes all of the variables
in the data set.
(b) Compute the matrix of correlations between the variables using
the function cor(). You will need to exclude the name variable, cor() which is qualitative.
(c) Use the lm() function to perform a multiple linear regression
with mpg as the response and all other variables except name as
the predictors. Use the summary() function to print the results.
Comment on the output. For instance:
i. Is there a relationship between the predictors and the response?
ii. Which predictors appear to have a statistically significant
relationship to the response?
iii. What does the coefficient for the year variable suggest?
(d) Use the plot() function to produce diagnostic plots of the linear
regression fit. Comment on any problems you see with the fit.
Do the residual plots suggest any unusually large outliers? Does
the leverage plot identify any observations with unusually high
leverage?
(e) Use the * and : symbols to fit linear regression models with
interaction effects. Do any interactions appear to be statistically
significant?
(f) Try a few different transformations of the variables, such as
log(X), ???
X, X2. Comment on your findings.
3.7 Exercises 123
10. This question should be answered using the Carseats data set.
(a) Fit a multiple regression model to predict Sales using Price,
Urban, and US.
(b) Provide an interpretation of each coefficient in the model. Be
careful-some of the variables in the model are qualitative!
(c) Write out the model in equation form, being careful to handle
the qualitative variables properly.
(d) For which of the predictors can you reject the null hypothesis
H0 : ??j = 0?
(e) On the basis of your response to the previous question, fit a
smaller model that only uses the predictors for which there is
evidence of association with the outcome.
(f) How well do the models in (a) and (e) fit the data?
(g) Using the model from (e), obtain 95 % confidence intervals for
the coefficient(s).
(h) Is there evidence of outliers or high leverage observations in the
model from (e)?
11. In this problem we will investigate the t-statistic for the null hypothesis
H0 : ?? = 0 in simple linear regression without an intercept. To
begin, we generate a predictor x and a response y as follows.
> set.seed(1)
> x=rnorm(100)
> y=2*x+rnorm (100)
(a) Perform a simple linear regression of y onto x, without an intercept.
Report the coefficient estimate ??^, the standard error of
this coefficient estimate, and the t-statistic and p-value associated
with the null hypothesis H0 : ?? = 0. Comment on these
results. (You can perform regression without an intercept using
the command lm(y???x+0).)
(b) Now perform a simple linear regression of x onto y without an
intercept, and report the coefficient estimate, its standard error,
and the corresponding t-statistic and p-values associated with
the null hypothesis H0 : ?? = 0. Comment on these results.
(c) What is the relationship between the results obtained in (a) and
(b)?
(d) For the regression of Y onto X without an intercept, the tstatistic
for H0 : ?? = 0 takes the form ??/^ SE(??^), where ??^ is
given by (3.38), and where
SE(??^) = !
n
i=1(yi ??? xi??^)2
(n ??? 1)
n
i=1 x2
i
.
124 3. Linear Regression
(These formulas are slightly different from those given in Sections
3.1.1 and 3.1.2, since here we are performing regression
without an intercept.) Show algebraically, and confirm numerically
in R, that the t-statistic can be written as
(
???n ??? 1)
n
i=1 xiyi
(

n
i=1 x2
i )(
n
i=1 y2
i ) ??? (

n
i=1 xiyi )2 .
(e) Using the results from (d), argue that the t-statistic for the regression
of y onto x is the same as the t-statistic for the regression
of x onto y.
(f) In R, show that when regression is performed with an intercept,
the t-statistic for H0 : ??1 = 0 is the same for the regression of y
onto x as it is for the regression of x onto y.
12. This problem involves simple linear regression without an intercept.
(a) Recall that the coefficient estimate ??^ for the linear regression of
Y onto X without an intercept is given by (3.38). Under what
circumstance is the coefficient estimate for the regression of X
onto Y the same as the coefficient estimate for the regression of
Y onto X?
(b) Generate an example in R with n = 100 observations in which
the coefficient estimate for the regression of X onto Y is different
from the coefficient estimate for the regression of Y onto X.
(c) Generate an example in R with n = 100 observations in which
the coefficient estimate for the regression of X onto Y is the
same as the coefficient estimate for the regression of Y onto X.
13. In this exercise you will create some simulated data and will fit simple
linear regression models to it. Make sure to use set.seed(1) prior to
starting part (a) to ensure consistent results.
(a) Using the rnorm() function, create a vector, x, containing 100
observations drawn from a N(0, 1) distribution. This represents
a feature, X.
(b) Using the rnorm() function, create a vector, eps, containing 100
observations drawn from a N(0, 0.25) distribution i.e. a normal
distribution with mean zero and variance 0.25.
(c) Using x and eps, generate a vector y according to the model
Y = ???1+0.5X + . (3.39)
What is the length of the vector y? What are the values of ??0
and ??1 in this linear model?
3.7 Exercises 125
(d) Create a scatterplot displaying the relationship between x and
y. Comment on what you observe.
(e) Fit a least squares linear model to predict y using x. Comment
on the model obtained. How do ??^0 and ??^1 compare to ??0 and
??1?
(f) Display the least squares line on the scatterplot obtained in (d).
Draw the population regression line on the plot, in a different
color. Use the legend() command to create an appropriate legend.
(g) Now fit a polynomial regression model that predicts y using x
and x2. Is there evidence that the quadratic term improves the
model fit? Explain your answer.
(h) Repeat (a)-(f) after modifying the data generation process in
such a way that there is less noise in the data. The model (3.39)
should remain the same. You can do this by decreasing the variance
of the normal distribution used to generate the error term
 in (b). Describe your results.
(i) Repeat (a)-(f) after modifying the data generation process in
such a way that there is more noise in the data. The model
(3.39) should remain the same. You can do this by increasing
the variance of the normal distribution used to generate the
error term  in (b). Describe your results.
(j) What are the confidence intervals for ??0 and ??1 based on the
original data set, the noisier data set, and the less noisy data
set? Comment on your results.
14. This problem focuses on the collinearity problem.
(a) Perform the following commands in R:
> set.seed(1)
> x1=runif (100)
> x2=0.5*x1+rnorm (100)/10
> y=2+2*x1+0.3*x2+rnorm (100)
The last line corresponds to creating a linear model in which y is
a function of x1 and x2. Write out the form of the linear model.
What are the regression coefficients?
(b) What is the correlation between x1 and x2? Create a scatterplot
displaying the relationship between the variables.
(c) Using this data, fit a least squares regression to predict y using
x1 and x2. Describe the results obtained. What are ??^0, ??^1, and
??^2? How do these relate to the true ??0, ??1, and ??2? Can you
reject the null hypothesis H0 : ??1 = 0? How about the null
hypothesis H0 : ??2 = 0?
126 3. Linear Regression
(d) Now fit a least squares regression to predict y using only x1.
Comment on your results. Can you reject the null hypothesis
H0 : ??1 = 0?
(e) Now fit a least squares regression to predict y using only x2.
Comment on your results. Can you reject the null hypothesis
H0 : ??1 = 0?
(f) Do the results obtained in (c)-(e) contradict each other? Explain
your answer.
(g) Now suppose we obtain one additional observation, which was
unfortunately mismeasured.
> x1=c(x1, 0.1)
> x2=c(x2, 0.8)
> y=c(y,6)
Re-fit the linear models from (c) to (e) using this new data. What
effect does this new observation have on the each of the models?
In each model, is this observation an outlier? A high-leverage
point? Both? Explain your answers.
15. This problem involves the Boston data set, which we saw in the lab
for this chapter. We will now try to predict per capita crime rate
using the other variables in this data set. In other words, per capita
crime rate is the response, and the other variables are the predictors.
(a) For each predictor, fit a simple linear regression model to predict
the response. Describe your results. In which of the models is
there a statistically significant association between the predictor
and the response? Create some plots to back up your assertions.
(b) Fit a multiple regression model to predict the response using
all of the predictors. Describe your results. For which predictors
can we reject the null hypothesis H0 : ??j = 0?
(c) How do your results from (a) compare to your results from (b)?
Create a plot displaying the univariate regression coefficients
from (a) on the x-axis, and the multiple regression coefficients
from (b) on the y-axis. That is, each predictor is displayed as a
single point in the plot. Its coefficient in a simple linear regression
model is shown on the x-axis, and its coefficient estimate
in the multiple linear regression model is shown on the y-axis.
(d) Is there evidence of non-linear association between any of the
predictors and the response? To answer this question, for each
predictor X, fit a model of the form
Y = ??0 + ??1X + ??2X2 + ??3X3 + .