---
title: "Lecture2 - baseball stats"
---
```{r}
library(Lahman)
library(mosaic)
library(tidyverse)

data(Teams)
glimpse(Teams)
```

```{r}
# get teams after 1970 with new variables
teams1 <- Teams %>%
  filter(yearID >= 1970) %>%
  mutate(X1B = H - X2B - X3B - HR, # derive single from hits minus doubles + triples + HR
         TB = X1B + 2*X2B + 3*X3B + 4*HR, # calculate total bases
         RC = ((H + BB)*TB)/(AB + BB)) # calculate runs created
teams1 %>%
  ggplot(aes(x=RC,y=R)) + 
  geom_point(colour = "black", fill = "white", alpha=.3) +
  ggtitle("Runs by Runs Created, 1970-2015")
#xyplot(R~RC, data=teams1)
```

There's a very strong, positive relationship between Runs and RC, which we can clarify via a correlation calculation to get `r`, and get `R2` as well.

```{r}
(r <- cor(teams1$R,teams1$RC))
(r2 <- cor(teams1$R,teams1$RC)**2)
```

Can now check association between a team's totals runs and other quantitative variables.
```{r}
(cor_mtrx <- round(cor(select(teams1, R, RC, AB, H, HR, SO, SB, CS, RA, DP)),3))
```

Looks like H and HR are the only ones with stronger positive relationships, while RA and AB have a moderate positive relationship, with CS having the only negative relationship, but it's quite weak. This suggests that a team's total runs are influenced heavily by more RC, H, and HR, and less so by RA and AB. We can also view this as a correlation plot
```{r}
library(corrplot)
corrplot(cor_mtrx,method="number")
#corrplot(cor_mtrx) = gives circles, not #'s of r values
```

What about runs created against more popular but advanced metrics?
```{r}
teams1 <- teams1 %>%
  mutate(OBP = (H+BB)/(AB+BB),
         SLG = (X1B+2*X2B+3*X3B+4*HR)/AB,
         OPS = OBP+SLG)

cor_mtrx2 <- round(cor(select(teams1, R, RC, OBP, SLG, OPS)),3)
corrplot(cor_mtrx2,method="number")
```

All 4 variables have strong positive relationships with runs, with OBP having the lowest, but we can ignore it as OPS contains effects of both OBP and SLG.

**What we've shown**:

* 1 - Is runs created important to success?
  * Yes. Strong link to team runs
	
* 2 - How well does the metric measure a player's contribution?
  * Pretty well. 
  * Other advanced formulas exist
  * Adjustments possible

* 3 - Is the metric repeatable? 
  * Let's find out
  
### Repeatability
```{r}
teams2 <- teams1 %>%
  arrange(franchID, yearID) %>%
  group_by(franchID) %>%
  mutate(next_yr_R = dplyr::lead(R)) # lead() = find "next"/"previous" values in a vector

cor_mtrx3 <- round(cor(select(ungroup(teams2), next_yr_R, R, RC, OBP, SLG, OPS), use="pairwise.complete.obs"),3)
corrplot(cor_mtrx3,method="number")
```

All have weak positive relationship.